\documentclass[11pt,a4paper]{article}

% ============================================================================
% PACKAGES
% ============================================================================
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{mathtools}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{hyperref}
\usepackage{cleveref}
\usepackage{xcolor}
\usepackage{listings}
\usepackage{tikz}
\usetikzlibrary{arrows,shapes,positioning,calc,fit,backgrounds}
\usepackage{geometry}
\geometry{margin=1in}
\usepackage{natbib}
\usepackage{enumitem}
\usepackage{subcaption}
\usepackage{multirow}
\setlength{\emergencystretch}{3em}

% ============================================================================
% THEOREM ENVIRONMENTS
% ============================================================================
\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{remark}{Remark}[section]
\newtheorem{example}{Example}[section]

% Make algorithmic line anchors globally unique under hyperref.
\makeatletter
\renewcommand{\theALG@line}{\thealgorithm.\arabic{ALG@line}}
\makeatother

% ============================================================================
% CUSTOM COMMANDS
% ============================================================================
\newcommand{\dataset}{\mathcal{D}}
\newcommand{\query}{\mathbf{q}}
\newcommand{\score}{\text{score}}
\newcommand{\embed}{\phi}
\newcommand{\graph}{\mathcal{G}}
\newcommand{\nodes}{\mathcal{V}}
\newcommand{\edges}{\mathcal{E}}
\newcommand{\rel}{\text{rel}}
\newcommand{\confidence}{\gamma}
\newcommand{\evidence}{\mathcal{E}\text{vid}}
\newcommand{\fingerprint}{\mathbf{f}}
\newcommand{\compatibility}{\kappa}
\newcommand{\RRF}{\text{RRF}}
\newcommand{\BM}{\text{BM25}}
\newcommand{\NDCG}{\text{NDCG}}
\newcommand{\MRR}{\text{MRR}}
\newcommand{\sigmoid}{\sigma}

% Code listing style
\lstset{
    basicstyle=\ttfamily\small,
    keywordstyle=\color{blue},
    commentstyle=\color{gray},
    stringstyle=\color{orange},
    breaklines=true,
    frame=single,
    numbers=left,
    numberstyle=\tiny\color{gray}
}

% ============================================================================
% TITLE
% ============================================================================
\title{
    \textbf{Neural Search: Experiment-Aware Retrieval, Graph Reasoning, and Provenance for Reusable Neuroscience Datasets}\\[0.5em]
    \large A Scientific and Mathematical Foundation for Structured Scientific Discovery
}

\author{
    Neural Search Development Team\\
    \texttt{neural-search@research.org}
}

\date{June 2026 -- updated 30 June 2026}

\begin{document}

\maketitle

% ============================================================================
% ABSTRACT
% ============================================================================
\begin{abstract}
Neural Search is an experiment-aware discovery framework for neural, behavioral, clinical, molecular, connectomics, morphology, imaging, and computational neuroscience datasets. Unlike generic document retrieval or chunk-based RAG systems, Neural Search treats datasets as structured scientific objects with typed metadata, ontological grounding, provenance tracking, dense field embeddings, and graph-based relational context. The current live corpus artifact contains 7,171 normalized records across DANDI, OpenNeuro, NeuroVault, Zenodo, Figshare, NeuroMorpho, Allen, GIN, Brain Image Library, BlueBrain, IBL, CRCNS, OSF, Harvard Dataverse, Buzsaki Lab, SPARK, and smaller sources. Corpus expansion is implemented as a tiered ingestion and enrichment pipeline: high-volume open repositories are admitted through raw/processed-data, neuroscience-signal, license, and persistent-identifier gates, while curated neuroscience sources and DANDI/NWB enrichers add modality, species, task, paper-link, and brain-region evidence with explicit provenance. A full-corpus BGE-large-en-v1.5 dense field embedding cache is now materialized as \texttt{real\_all.dense.field\_embeddings.agg.ids.json} and \texttt{real\_all.dense.field\_embeddings.agg.mat.npy}, covering all 7,171 corpus records at 1,024 dimensions; the ablation ladder loads this cache before falling back to FAISS or JSONL parsing, reducing embedding-load latency from seconds to milliseconds. The older 625-record turbovec index remains a separate approximate-nearest-neighbor engineering artifact and should be regenerated against the full embedding cache before final ANN claims. We also implement a layered Field-State memory system: a lightweight claims/gaps/Obsidian review layer for scientific claim tracking, and a \textbf{Field-State Memory Graph}---a relational graph substrate that stores datasets, papers, concepts, modalities, species, brain regions, tasks, affordances, neuro-judge evidence packets, user feedback signals, and versioned snapshots in a single queryable, exportable, incrementally-updatable graph. The full-corpus knowledge graph (\texttt{neural\_search\_graph.real\_corpus.json}, rebuilt 30~June~2026) contains 7,438 typed nodes and 81,518 typed edges across 7,171 dataset nodes, 118 brain-region nodes, 85 task nodes, 46 modality nodes, and 18 species nodes. Dataset nodes now use real corpus identifiers such as \texttt{node:dataset:dandi:000785} and carry aliases such as \texttt{dandi:000785}, \texttt{000785}, and \texttt{dataset:dandi:000785}, matching the retrieval layer's graph-node resolver. This fixes the previous zero-overlap failure mode in which retrieval IDs such as \texttt{dandi:000785} could not resolve to demo-era graph nodes. In addition to 22,997 base dataset-to-concept edges, the graph now contains 58,521 explicit cross-dataset relationship edges: \texttt{same\_region\_cross\_modality} (48,212) and \texttt{same\_task\_cross\_species} (10,309). A separate field-state memory graph (\texttt{update\_20260613T053153Z}) tracks 2,200 nodes and 3,788 typed edges spanning evidence packets, feedback signals, and curation artifacts for the 629-record reviewed slice. Provenance invariants keep neuro-judge silver labels, CrossRef/NWB-derived silver enrichments, downstream user feedback, and human gold labels separate. A DuckDB-backed coverage ledger (7,176 datasets; 29,561 entries across six dimensions: brain\_regions, modalities, recording\_scales, species, tasks, behavioral\_events) enables gap-analysis queries, source coverage rates, and dark-pair identification. A coverage gap boost adds an additive inverse-log-frequency rarity signal (capped at $+0.10$) to retrieval scores, preferring underrepresented anatomical regions and modalities. A source-diversity post-processor applies a greedy windowed algorithm (default \texttt{max\_per\_source}=3) to prevent single-repository domination of result lists. A SPECTER2 hybrid retrieval module (config-gated, disabled by default) adds scientific paper embedding similarity as an additive signal. An LLM query expansion fallback (Claude Haiku, LRU-cached, fails safe) activates only when rule-based parsing extracts zero brain regions and zero tasks, filling empty queries with structured ontological terms. Structured dimension-match badges in the frontend surface per-dimension evidence quality (full/partial/inferred/none) for every result card. An incremental update pipeline detects changed records by content hash, produces versioned snapshot manifests, computes memory diffs, preserves human review overlays, and generates provisional learning signals without overwriting canonical labels. A literature ingestion layer has linked papers to corpus datasets and run rule-based typed-finding extraction (27 fields: frequency band, temporal pattern, negation, spatial frame, and more) over a tier-1 finding-extraction run that is still in progress (190K+ findings as of 23~June~2026). Current validation artifacts show preliminary usefulness-score separation between relevant and irrelevant results, a now-nonzero graph-context signal for real corpus IDs, and a canonical LLM-judged qrels snapshot with 317 queries and 13,654 non-error query--dataset labels. The qrels review workflow has also been smoke-tested for human adjudication, but no publication-grade human-labeled retrieval benchmark is complete yet. The next required validation steps are dual-judge agreement, human adjudication of a stratified subset, deterministic exact lookup, hard-negative violation tracking, source-skew analysis, calibration, and content-validated analysis affordances.
\end{abstract}

\begin{table}[htbp]
\centering
\caption{Current Artifact Status}
\label{tab:current-artifact-status}
\small
\begin{tabular}{p{0.26\linewidth}p{0.24\linewidth}p{0.42\linewidth}}
\toprule
Artifact or layer & Current status & Validation caveat \\
\midrule
Normalized dataset corpus & 7,171 rows; 7,121 unique record identifiers & Source-specific metadata precision audits pending. \\
Dense embeddings and turbovec index & Full 7,171-record dense field embedding cache, 1,024 dimensions; older 625-record turbovec ANN index still pending rebuild & Dense retrieval can now cover the real corpus; ANN recall/latency must be regenerated before final claims. \\
Full-corpus knowledge graph & 7,438 nodes and 81,518 edges, including 58,521 cross-dataset relation edges & Real corpus IDs now resolve into graph context; qrels-backed improvement remains preliminary and human validation is pending. \\
Coverage ledger and UI evidence surfaces & 7,176 datasets and 29,561 dimension entries; Atlas and affordance panels implemented & Affordance precision and content-derived validation are pending. \\
Literature ingestion & 255,940 OpenAlex tier~1 neuroscience papers ingested & Literature search is not yet integrated into the unified retrieval path. \\
Paper-dataset linker & Linker run over all 7,171 datasets; 168 DOI-exact links, 225 title-fuzzy-local links, 6,778 not found & ``Run over all datasets'' is not the same as ``all datasets linked''; link precision audit pending. \\
Finding extraction & 122K findings enriched by a 27-field typed pass; earlier extraction snapshot: 16,150 papers processed; 12,050 findings; 74.6\% yield & Live operational estimate and field-coverage audit only; not validated finding quality or final yield. \\
Qrels and benchmark layer & Export/import/adjudication workflow smoke-tested; three adjudicated labels & LLM qrels are silver diagnostics only; publication-grade NDCG/MRR/Recall requires human-adjudicated qrels. \\
\bottomrule
\end{tabular}
\end{table}

\tableofcontents
\newpage

% ============================================================================
% SECTION 1: INTRODUCTION
% ============================================================================
\section{Introduction}
\label{sec:introduction}

\subsection{Motivation and Problem Statement}

The exponential growth of neuroscience datasets across repositories and standards such as DANDI and NWB \citep{dandiArchive2026,ruebel2022nwb}, OpenNeuro and BIDS \citep{markiewicz2021openneuro,gorgolewski2016bids}, NeuroVault, Zenodo, Figshare, NeuroMorpho, Allen Brain Observatory, GIN, OSF, CRCNS, IBL, Brain Image Library, and BlueBrain, paired with the parallel explosion of published findings indexed by services such as OpenAlex \citep{priem2022openalex}, presents a fundamental discovery challenge. Researchers seeking datasets for replication, meta-analysis, or novel analysis face a search problem that traditional document retrieval cannot adequately address. The reproducibility goal is aligned with FAIR data principles \citep{wilkinson2016fair}.

\begin{definition}[Scientific Dataset Search Problem]
Given a query $\query$ expressing experimental intent (tasks, modalities, species, analysis goals, constraints), retrieve a ranked list of datasets $\dataset_1, \dataset_2, \ldots, \dataset_k$ such that each $\dataset_i$ maximally satisfies the experimental requirements while providing sufficient metadata, quality assurance, and analysis affordances for scientific reuse.
\end{definition}

This problem differs fundamentally from web search or document retrieval:

\begin{enumerate}[label=(\roman*)]
    \item \textbf{Structured constraints}: A query like ``mouse Neuropixels decision-making data'' implicitly specifies organism (Mus musculus), recording modality (Neuropixels electrophysiology), and task ontology (decision-making behavioral paradigm).

    \item \textbf{Hard negatives}: Exclusions such as ``NOT EEG'' or ``without fMRI'' represent hard constraints, not soft preferences.

    \item \textbf{Analysis affordances}: Relevance depends not only on what data exists but whether it supports intended analyses (e.g., choice decoding, Q-learning modeling).

    \item \textbf{Provenance requirements}: Scientific reuse demands traceable evidence for metadata claims.
\end{enumerate}

\subsection{Thesis Statement}

The central thesis of Neural Search is that \textbf{neuroscience search should retrieve reusable experimental contexts, not just documents}. We formalize this through:

\begin{enumerate}
    \item \textbf{Dataset as experiment}: The object of search is a normalized dataset record with typed scientific structure.
    \item \textbf{Meaning as typed structure}: Scientific concepts are normalized into ontological categories.
    \item \textbf{Search as explanation}: Every result carries interpretable evidence.
    \item \textbf{Provenance as guardrail}: Assertions carry confidence and extraction lineage.
    \item \textbf{Benchmarks as compass}: Ranking changes are evaluated against reproducible metrics.
\end{enumerate}

\subsection{Neuronpedia Platform Vision}
\label{sec:neuronpedia}

The broader design goal of Neural Search is to serve as a \textit{Neuronpedia for neuroscience}: a universal shared platform for discovering, discussing, and understanding neuroscience datasets across all brain regions and timescales. Analogous to how Neuronpedia surfaces neuron-level features from sparse autoencoders, Neural Search surfaces dataset-level experimental affordances grounded in structured ontologies and a knowledge graph.

The platform enables three complementary discovery modes:
\begin{enumerate}
    \item \textbf{Query-driven discovery:} Researchers enter a natural-language experimental intent, and the hybrid retrieval engine returns datasets ranked by experiment-to-experiment compatibility.
    \item \textbf{Atlas-driven browsing:} A Brain Atlas interface renders all 130+ canonical brain regions as a color-coded coverage heatmap; clicking any region reveals the datasets that record from it.
    \item \textbf{Affordance-driven filtering:} An Analysis Affordance panel on each dataset card shows which of the 21 registered analysis types are structurally supported, with required-feature evidence for each.
\end{enumerate}

This positions Neural Search not merely as a search engine but as a \textit{corpus intelligence layer} that accelerates the path from research question to validated, analysis-ready dataset.

\subsection{Contributions and Current Scope}

This paper distinguishes implemented infrastructure from completed empirical validation. The current contributions are:

\begin{enumerate}
    \item A neuroscience dataset retrieval architecture combining normalized corpus metadata, dense field embeddings, structured constraints, graph context, and intent-aware usefulness signals.
    \item A corpus expansion pipeline that combines high-volume open-source ingestion, source-specific normalizers, conservative inclusion gates, DANDI/NWB metadata enrichers, paper-abstract mining, and regional coverage reports.
    \item A benchmark protocol for graded dataset-query relevance, hard-negative violations, calibration, source-skew analysis, and ablation over sparse, dense, graph, provenance, and usefulness components.
    \item A lightweight Field-State layer that tracks scientific claims, benchmark gaps, high-value next opportunities, heuristic opportunity scores, JSONL artifacts, and Markdown reports.
    \item A local Obsidian-compatible review memory that preserves human notes, review overlays, validation decisions, memory indices, sync logs, and diff reports without turning Obsidian into the primary database.
    \item An evaluation-memory system that exports qrels candidates, imports human labels, creates adjudicated qrels artifacts, computes agreement summaries, and links evidence status back to field-state claims.
\end{enumerate}

Items 1--6 are implemented as repository infrastructure and covered by targeted engineering tests. The empirical claim that Neural Search improves scientific retrieval quality on the expanded corpus remains future validation until the corpus, embedding, and index manifests are reconciled, adjudicated qrels exist at benchmark scale, and run files, metric reports, hard-negative analyses, source-skew diagnostics, and calibration reports are complete.

\subsection{Artifact Reconciliation and Evidence Tiers}
\label{sec:artifact-reconciliation}

The following table records the current source of truth for each major artifact layer as of the \textbf{2026-06-20} repository state. All quantitative claims in this paper must be traced to one of these layers; any claim traced to a stale or quarantined report should be treated as historical until reproduced from the artifacts below.

\begin{table}[htbp]
\centering
\caption{Artifact truth table (2026-06-30). Canonical manifest: \texttt{reports/eval/current\_artifact\_manifest.json}.}
\label{tab:artifact-truth}
\small
\begin{tabular}{p{0.22\linewidth}p{0.35\linewidth}p{0.33\linewidth}}
\toprule
Layer & Canonical artifact & Validation caveat \\
\midrule
Normalized corpus & \texttt{data/corpus/normalized/combined\_corpus.jsonl/full\_corpus\_v09.jsonl} — 7,171 rows, 7,121 unique \texttt{source\_id}s & \texttt{combined\_corpus.jsonl} is a directory, not a file; loaders must reference \texttt{full\_corpus\_v09.jsonl} explicitly \\
Dense index & \texttt{real\_all.dense.field\_embeddings.agg.mat.npy}: 7,171 records $\times$ 1,024 dim; older \texttt{turbovec} ANN index: 625 records $\times$ 1,024 dim & Full dense retrieval cache is current; ANN recall/latency metrics require a turbovec rebuild \\
Full-corpus KG & \texttt{data/graph/neural\_search\_graph.real\_corpus.json} -- 7,438 nodes / 81,518 edges & Real corpus IDs now resolve into graph context; human-qrels-backed improvement unproven \\
Field-state memory graph & \texttt{artifacts/field\_state/memory\_graph\_manifest.json} — 629-record curation slice, 2,200 nodes / 3,788 edges & Curated subset; not full-corpus \\
Gold qrels & \texttt{artifacts/qrels\_gold.jsonl} — 0 rows & Pending human annotation \\
Field-state adjudicated & \texttt{artifacts/field\_state/adjudicated\_qrels.jsonl} — 3 rows & Smoke-test only \\
Silver/bronze qrels & \texttt{artifacts/qrels\_silver.jsonl} (175 rows), \texttt{artifacts/qrels\_bronze.jsonl} (319 rows) & LLM-generated diagnostic signals; not peer-reviewed \\
Finding extraction & 122K findings enriched by a 27-field typed pass; earlier extraction snapshot: 16,150 papers / 12,050 findings & Quality not validated; treat as audit queue \\
\bottomrule
\end{tabular}
\end{table}

Reports that cite 10,404 corpus records (\texttt{reports/eval/corpus\_manifest.json}), 675 annotated pairs (\texttt{reports/eval/neuro\_judge\_retrieval\_metrics.md}), or an index of 4,366 records are stale historical snapshots and should not be cited as current state.

Evidence tiers used throughout this paper:
\begin{itemize}
    \item \textbf{Tier A — publishable infrastructure:} implemented code, engineering tests, schema definitions, JSONL artifacts with verified row counts.
    \item \textbf{Tier B — engineering validation:} vector recall, index latency, graph rank perturbation, ablation on seed pairs. Useful for understanding system behavior but not final scientific retrieval evidence.
    \item \textbf{Tier C — silver diagnostic:} LLM-generated qrels, extracted findings, auto-linked papers. Used for triage and audit prioritization only.
    \item \textbf{Tier D — planned / pending:} human qrels at benchmark scale, affordance precision audits, user study, finding precision evaluation. Required before final performance claims.
\end{itemize}

\subsection{Why Existing Tools Are Not Enough}
\label{sec:why-needed}

Neuroscience data discovery currently relies on a patchwork of tools, each covering a fragment of the researcher's intent:

\paragraph{Archive-native search} (DANDI, OpenNeuro, NeuroVault, Allen, Zenodo) exposes structured metadata but provides weak scientific intent matching, no analysis feasibility reasoning, no cross-repository integration, and no provenance for metadata claims.

\paragraph{Literature knowledge graphs} such as BrainKnow \citep{wang2024brainknow} and Linked Brain Data extract concept co-occurrence at field scale (BrainKnow: 37K+ concepts, 3.6M+ relations, 1.8M PubMed papers). These systems excel at concept-level exploration and temporal literature synthesis. However, same-sentence co-occurrence does not constitute relation extraction: edges are largely unsigned, untyped, and undirected; direction, polarity, negation, causal interpretation, frequency band, and temporal pattern are not core graph semantics. Neural Search's current typed-finding pass instead records 27 fields on each extracted finding, including disease model, molecular marker, pharmacological agent, sensory stimulus, sleep stage, connectivity type, and dimensionality-reduction method. Crucially, BrainKnow is concept-centered, not dataset-centered — it does not answer: ``Which dataset can I download and run an analysis on to test this claim?''

\paragraph{Generic RAG and vector search} retrieve passages but do not enforce provenance, constrain by analysis feasibility, surface hard negatives, distinguish raw from processed data, or connect to reusable experimental contexts.

Neural Search bridges these gaps. Its distinguishing claim is not corpus size but \emph{research-object awareness}: every result is a dataset with typed metadata, an evidence trail, analysis affordances, links to the papers and findings that cite or test it, and explicit markers for what the evidence does and does not support. The goal is to answer: ``Which dataset can I use to test, reproduce, or challenge this specific claim?''

\subsection{Current Non-Claims}
\label{sec:non-claims}

To protect this paper from overreach and to set clear expectations for reviewers, the following claims are \textbf{explicitly not made} in the current version:

\begin{enumerate}
    \item \textbf{No final retrieval superiority claim.} Neural Search has not been shown to outperform BM25, dense retrieval, archive-native search, or BrainKnow-style co-occurrence on a shared human-adjudicated benchmark. That benchmark is planned (Workstream D: 100 queries, 1,500 pooled pairs) but not yet run.

    \item \textbf{No final human-adjudicated NDCG/MRR/Recall claim.} Current NDCG/MRR/Recall artifacts are silver LLM-qrels diagnostics over 317 canonical queries and 13,654 non-error query--dataset labels. Publication-grade retrieval metrics require frozen runs over human qrels.

    \item \textbf{No final finding-extraction precision claim.} The extraction and 27-field enrichment snapshots have been spot-checked but not formally evaluated for precision or recall. Region, species, and typed-field accuracy are unknown.

    \item \textbf{No gold-benchmark results.} Gold qrels contain 0 rows. Adjudicated qrels contain 3 rows (workflow smoke-test only).

    \item \textbf{No content-level neural signature search.} Neural signature extraction is not implemented. The dimension is excluded from active scoring.

    \item \textbf{No generalization claim.} Results on seed pairs should not be extrapolated across neuroscience subfields without cross-domain evaluation.
\end{enumerate}

The paper's strongest claims are about \emph{implemented architecture}, \emph{reproducible workflow}, \emph{provenance-backed infrastructure}, and the \emph{research-object retrieval framing} — not final empirical retrieval superiority.

% ============================================================================
% SECTION 2: MATHEMATICAL FOUNDATIONS
% ============================================================================
\section{Mathematical Foundations}
\label{sec:math-foundations}

\subsection{Information Retrieval Theory}

\subsubsection{Vector Space Model}

The Vector Space Model (VSM) represents documents and queries as vectors in a high-dimensional term space \citep{salton1975vector}.

\begin{definition}[Vector Space Representation]
Let $\mathcal{T} = \{t_1, t_2, \ldots, t_n\}$ be the vocabulary. A document $d$ is represented as:
\begin{equation}
    \mathbf{d} = (w_{1,d}, w_{2,d}, \ldots, w_{n,d}) \in \mathbb{R}^n
\end{equation}
where $w_{i,d}$ is the weight of term $t_i$ in document $d$.
\end{definition}

The relevance between query $\query$ and document $d$ is computed via cosine similarity:
\begin{equation}
    \text{sim}(\query, \mathbf{d}) = \frac{\query \cdot \mathbf{d}}{\|\query\| \cdot \|\mathbf{d}\|} = \frac{\sum_{i=1}^n w_{i,q} \cdot w_{i,d}}{\sqrt{\sum_{i=1}^n w_{i,q}^2} \cdot \sqrt{\sum_{i=1}^n w_{i,d}^2}}
\end{equation}

\subsubsection{TF-IDF Weighting}

Term Frequency-Inverse Document Frequency (TF-IDF) balances term importance:

\begin{definition}[TF-IDF]
For term $t$ in document $d$ within corpus $\mathcal{C}$:
\begin{equation}
    \text{TF-IDF}(t, d, \mathcal{C}) = \text{TF}(t, d) \cdot \text{IDF}(t, \mathcal{C})
\end{equation}
where:
\begin{align}
    \text{TF}(t, d) &= \frac{f_{t,d}}{\max_{t' \in d} f_{t',d}} \quad \text{(normalized term frequency)} \\
    \text{IDF}(t, \mathcal{C}) &= \log \frac{|\mathcal{C}|}{|\{d \in \mathcal{C} : t \in d\}|} \quad \text{(inverse document frequency)}
\end{align}
\end{definition}

\subsubsection{BM25: Best Matching 25}

BM25 is a probabilistic relevance model that addresses TF-IDF limitations through saturation and length normalization \citep{robertson2009probabilistic}:

\begin{definition}[BM25 Scoring Function]
\begin{equation}
    \BM(d, \query) = \sum_{t \in \query} \text{IDF}(t) \cdot \frac{f_{t,d} \cdot (k_1 + 1)}{f_{t,d} + k_1 \cdot \left(1 - b + b \cdot \frac{|d|}{\text{avgdl}}\right)}
\end{equation}
where:
\begin{itemize}
    \item $f_{t,d}$ is the frequency of term $t$ in document $d$
    \item $|d|$ is the document length
    \item $\text{avgdl}$ is the average document length in the corpus
    \item $k_1 \in [1.2, 2.0]$ controls term frequency saturation
    \item $b \in [0, 1]$ controls length normalization (typically $b = 0.75$)
\end{itemize}
\end{definition}

The IDF component uses the Robertson-Sp\"arck Jones formula:
\begin{equation}
    \text{IDF}(t) = \log \frac{N - n_t + 0.5}{n_t + 0.5}
\end{equation}
where $N$ is the total number of documents and $n_t$ is the number containing term $t$.

\subsubsection{Field-Weighted BM25}

For structured scientific records, we extend BM25 to multiple fields:

\begin{definition}[Field-Weighted BM25]
Given fields $\mathcal{F} = \{f_1, f_2, \ldots, f_m\}$ with weights $\mathbf{w} = (w_1, \ldots, w_m)$:
\begin{equation}
    \BM_{\text{field}}(d, \query) = \sum_{f \in \mathcal{F}} w_f \cdot \BM(d_f, \query)
\end{equation}
where $d_f$ denotes the content of field $f$ in document $d$.
\end{definition}

In Neural Search, fields include: \texttt{title}, \texttt{description}, \texttt{tasks}, \texttt{modalities}, \texttt{species}, \texttt{brain\_regions}, \texttt{behavioral\_events}, \texttt{analysis\_goals}, and \texttt{standards}.

\subsection{Reciprocal Rank Fusion}

When combining multiple retrieval signals (lexical, semantic, graph-based), Reciprocal Rank Fusion (RRF) provides a robust aggregation method \citep{cormack2009reciprocal}:

\begin{definition}[Reciprocal Rank Fusion]
Given $m$ ranked lists $\{R_1, R_2, \ldots, R_m\}$, the RRF score for document $d$ is:
\begin{equation}
    \RRF(d) = \sum_{i=1}^m \frac{1}{k + \text{rank}_i(d)}
\end{equation}
where $\text{rank}_i(d)$ is the rank of $d$ in list $R_i$, and $k$ is a constant (typically $k = 60$) that mitigates the impact of high rankings.
\end{definition}

\begin{proposition}[RRF Properties]
RRF satisfies desirable fusion properties:
\begin{enumerate}
    \item \textbf{Rank-based}: Robust to score scale differences between rankers
    \item \textbf{Diminishing returns}: High ranks contribute more than low ranks
    \item \textbf{Consensus rewarding}: Documents ranked highly by multiple sources receive higher scores
\end{enumerate}
\end{proposition}

\subsection{Dense Retrieval and Embedding Spaces}

\subsubsection{Neural Embeddings}

Dense retrieval represents queries and documents in learned embedding spaces:

\begin{definition}[Bi-Encoder Architecture]
Given encoder $\embed: \mathcal{X} \rightarrow \mathbb{R}^d$, relevance is:
\begin{equation}
    \text{sim}(\query, d) = \embed(\query)^\top \embed(d)
\end{equation}
\end{definition}

\subsubsection{Semantic Fingerprints}

Neural Search extends single-vector embeddings to multi-aspect semantic fingerprints:

\begin{definition}[Semantic Fingerprint]
A dataset $d$ is represented as a concatenated multi-field embedding:
\begin{equation}
    \fingerprint(d) = \left[ \embed_{\text{text}}(d) \| \embed_{\text{task}}(d) \| \embed_{\text{mod}}(d) \| \embed_{\text{beh}}(d) \| \embed_{\text{ana}}(d) \| \embed_{\text{reg}}(d) \| \embed_{\text{des}}(d) \right]
\end{equation}
where $\|$ denotes concatenation and each $\embed_*$ encodes a specific semantic aspect.
\end{definition}

This factorization enables dimension-specific similarity:
\begin{equation}
    \text{sim}_{\text{aspect}}(d_i, d_j) = \frac{\embed_{\text{aspect}}(d_i) \cdot \embed_{\text{aspect}}(d_j)}{\|\embed_{\text{aspect}}(d_i)\| \cdot \|\embed_{\text{aspect}}(d_j)\|}
\end{equation}

% ============================================================================
% SECTION 3: KNOWLEDGE GRAPH THEORY
% ============================================================================
\section{Knowledge Graph Theory}
\label{sec:graph-theory}

\subsection{Typed Directed Graphs}

\begin{definition}[Scientific Knowledge Graph]
A scientific knowledge graph is a tuple $\graph = (\nodes, \edges, \tau_V, \tau_E, \evidence)$ where:
\begin{itemize}
    \item $\nodes$ is the set of nodes (datasets, papers, concepts, species, analyses)
    \item $\edges \subseteq \nodes \times \nodes$ is the set of directed edges
    \item $\tau_V: \nodes \rightarrow \mathcal{T}_V$ assigns node types from type set $\mathcal{T}_V$
    \item $\tau_E: \edges \rightarrow \mathcal{T}_E$ assigns edge types from relation set $\mathcal{T}_E$
    \item $\evidence: \nodes \cup \edges \rightarrow \mathcal{P}(\text{Evidence})$ maps elements to provenance evidence
\end{itemize}
\end{definition}

The Neural Search type system includes:
\begin{align*}
    \mathcal{T}_V &= \{\texttt{dataset}, \texttt{paper}, \texttt{task}, \texttt{modality}, \texttt{species}, \texttt{region}, \\
    &\quad\quad \texttt{event}, \texttt{analysis}, \texttt{standard}, \texttt{format}, \texttt{requirement}\} \\
    \mathcal{T}_E &= \{\texttt{has\_task}, \texttt{has\_modality}, \texttt{records\_region}, \texttt{has\_species}, \\
    &\quad\quad \texttt{uses\_standard}, \texttt{supports\_analysis}, \texttt{requires}, \texttt{linked\_to\_paper}, \\
    &\quad\quad \texttt{same\_region\_cross\_modality}, \texttt{same\_task\_cross\_species}, \texttt{same\_region\_same\_task}\}
\end{align*}

\subsection{Provenance-Weighted Edges}

\begin{definition}[Edge Confidence]
Each edge $e \in \edges$ carries a confidence score:
\begin{equation}
    \confidence(e) = \sigma\left(\sum_{ev \in \evidence(e)} w_{\text{source}}(ev) \cdot w_{\text{method}}(ev)\right)
\end{equation}
where $\sigma$ is the sigmoid function, $w_{\text{source}}$ weights evidence source reliability, and $w_{\text{method}}$ weights extraction method confidence.
\end{definition}

Confidence propagates through paths:
\begin{equation}
    \confidence(\text{path}) = \prod_{e \in \text{path}} \confidence(e)
\end{equation}

\subsection{Metapath Theory}

Metapaths define semantically meaningful traversal patterns in heterogeneous graphs \citep{sun2011pathsim}:

\begin{definition}[Metapath]
A metapath $\mathcal{P}$ is a sequence of node types connected by edge types:
\begin{equation}
    \mathcal{P} = T_1 \xrightarrow{R_1} T_2 \xrightarrow{R_2} \cdots \xrightarrow{R_{l-1}} T_l
\end{equation}
\end{definition}

\begin{example}[Scientific Metapaths]
\begin{align*}
    \mathcal{P}_{\text{analysis}} &= \texttt{Dataset} \xrightarrow{\texttt{has\_event}} \texttt{Event} \xrightarrow{\texttt{required\_for}} \texttt{Analysis} \\
    \mathcal{P}_{\text{paper}} &= \texttt{Dataset} \xrightarrow{\texttt{linked\_to}} \texttt{Paper} \xrightarrow{\texttt{studies}} \texttt{Task} \\
    \mathcal{P}_{\text{similar}} &= \texttt{Dataset} \xrightarrow{\texttt{has\_task}} \texttt{Task} \xleftarrow{\texttt{has\_task}} \texttt{Dataset}
\end{align*}
\end{example}

\begin{definition}[Metapath-Based Similarity]
The PathSim similarity between nodes $v_i$ and $v_j$ along metapath $\mathcal{P}$ is:
\begin{equation}
    \text{PathSim}(v_i, v_j | \mathcal{P}) = \frac{2 \cdot |\text{paths}_\mathcal{P}(v_i, v_j)|}{|\text{paths}_\mathcal{P}(v_i, v_i)| + |\text{paths}_\mathcal{P}(v_j, v_j)|}
\end{equation}
\end{definition}

\subsection{Graph Neural Features}

Graph-derived features augment retrieval:

\begin{definition}[Node Centrality Features]
For node $v \in \nodes$:
\begin{align}
    \text{degree}(v) &= |\{u : (v, u) \in \edges \text{ or } (u, v) \in \edges\}| \\
    \text{PageRank}(v) &= \frac{1-d}{N} + d \sum_{u \in \text{in}(v)} \frac{\text{PageRank}(u)}{|\text{out}(u)|} \\
    \text{betweenness}(v) &= \sum_{s \neq v \neq t} \frac{\sigma_{st}(v)}{\sigma_{st}}
\end{align}
where $d$ is the damping factor, $\sigma_{st}$ is the number of shortest paths from $s$ to $t$, and $\sigma_{st}(v)$ is the number passing through $v$.
\end{definition}

% ============================================================================
% SECTION 4: HYBRID SCORING FORMALIZATION
% ============================================================================
\section{Hybrid Scoring Formalization}
\label{sec:scoring}

\subsection{Multi-Signal Scoring Function}

The Neural Search scoring function combines heterogeneous evidence:

\begin{definition}[Operational Hybrid Scoring Function]
For dataset $d$ and query $\query$, the relevance score is:
\begin{equation}
\begin{split}
    \score(d, \query) = \text{clamp}\Bigg(
        &w_{\text{task}} \cdot S_{\text{task}}(d, \query) + w_{\text{beh}} \cdot S_{\text{behavior}}(d, \query) \\
        &+ w_{\text{mod}} \cdot S_{\text{modality}}(d, \query) + w_{\text{aff}} \cdot S_{\text{affordance}}(d, \query) \\
        &+ w_{\text{meta}} \cdot S_{\text{metadata}}(d, \query) + w_{\text{sem}} \cdot S_{\text{semantic}}(d, \query) \\
        &+ w_{\text{field}} \cdot S_{\text{field}}(d, \query) + w_{\text{graph}} \cdot S_{\text{graph}}(d, \query) \\
        &+ w_{\text{ready}} \cdot S_{\text{readiness}}(d) + w_{\text{paper}} \cdot S_{\text{paper}}(d, \query) \\
        &- \lambda_{\text{mismatch}} \cdot P_{\text{mismatch}}(d, \query) \\
        &- \lambda_{\text{missing}} \cdot P_{\text{missing}}(d, \query) \\
        &- \lambda_{\text{hard}} \cdot P_{\text{hard\_negative}}(d, \query),
    \quad 0, 1\Bigg)
\end{split}
\end{equation}
where $\text{clamp}(x, a, b) = \max(a, \min(x, b))$.
\end{definition}

\subsection{Component Score Functions}

\subsubsection{Ontology Matching Scores}

\begin{definition}[Set-Based Ontology Score]
For ontology-grounded fields (tasks, modalities, regions):
\begin{equation}
    S_{\text{onto}}(d, \query) = \frac{|L_d \cap L_\query|}{|L_\query|} \cdot \frac{1}{1 + \beta \cdot |L_d \setminus L_\query|}
\end{equation}
where $L_d$ is the label set of $d$, $L_\query$ is the query label set, and $\beta$ penalizes extraneous labels.
\end{definition}

With hierarchical ontologies, we use weighted Jaccard with ancestor expansion:
\begin{equation}
    S_{\text{hier}}(d, \query) = \frac{\sum_{l \in L_\query} \max_{l' \in L_d} \text{sim}_{\text{onto}}(l, l')}{|L_\query|}
\end{equation}

\subsubsection{Semantic Similarity Score}

\begin{equation}
    S_{\text{semantic}}(d, \query) = \frac{\embed(\query) \cdot \embed(d)}{\|\embed(\query)\| \cdot \|\embed(d)\|}
\end{equation}

\subsubsection{Field-Specific Semantic Score}

\begin{equation}
    S_{\text{field}}(d, \query) = \sum_{f \in \mathcal{F}} w_f \cdot \frac{\embed_f(\query) \cdot \embed_f(d)}{\|\embed_f(\query)\| \cdot \|\embed_f(d)\|}
\end{equation}

\subsubsection{Readiness Score}

Dataset readiness combines metadata completeness and quality indicators:
\begin{equation}
    S_{\text{readiness}}(d) = \alpha_1 \cdot \text{completeness}(d) + \alpha_2 \cdot \text{standardization}(d) + \alpha_3 \cdot \text{QA}(d)
\end{equation}

where:
\begin{align}
    \text{completeness}(d) &= \frac{|\text{populated\_fields}(d)|}{|\text{required\_fields}|} \\
    \text{standardization}(d) &= \mathbf{1}[d \text{ uses recognized standard}] \\
    \text{QA}(d) &= \text{quality\_assurance\_score}(d)
\end{align}

\subsection{Penalty Functions}

\subsubsection{Hard Negative Constraint Violation}

\begin{definition}[Hard Negative Penalty]
For excluded set $\mathcal{X}_\query$ (modalities, species, sources, etc.):
\begin{equation}
    P_{\text{hard\_negative}}(d, \query) =
    \begin{cases}
        +\infty & \text{if } L_d \cap \mathcal{X}_\query \neq \emptyset \\
        0 & \text{otherwise}
    \end{cases}
\end{equation}
\end{definition}

In practice, hard negatives are filtered before scoring, ensuring zero violations.

\subsubsection{Missing Required Metadata Penalty}

\begin{equation}
    P_{\text{missing}}(d, \query) = \sum_{r \in \text{required}(\query)} \mathbf{1}[\text{missing}(d, r)] \cdot w_r
\end{equation}

\subsection{Intent-Adaptive Weight Profiles}

Query intent determines weight profiles:

\begin{definition}[Intent Classification]
Let the intent space be:
\[
\mathcal{I} =
\{\texttt{lookup}, \texttt{task\_search}, \texttt{modality\_search},
\texttt{affordance\_search}, \texttt{similarity}, \texttt{exploration}\}.
\]
The weight vector $\mathbf{w}_i$ varies by intent.
\end{definition}

\begin{table}[htbp]
\centering
\caption{Intent-Specific Weight Profiles}
\begin{tabular}{lcccccc}
\toprule
Weight & Lookup & Task & Modality & Affordance & Similarity & Exploration \\
\midrule
$w_{\text{task}}$ & 0.10 & \textbf{0.30} & 0.15 & 0.15 & 0.20 & 0.15 \\
$w_{\text{modality}}$ & 0.10 & 0.15 & \textbf{0.30} & 0.15 & 0.15 & 0.15 \\
$w_{\text{affordance}}$ & 0.05 & 0.15 & 0.10 & \textbf{0.35} & 0.10 & 0.10 \\
$w_{\text{semantic}}$ & 0.15 & 0.10 & 0.10 & 0.10 & \textbf{0.25} & 0.20 \\
$w_{\text{graph}}$ & 0.05 & 0.10 & 0.10 & 0.10 & 0.15 & \textbf{0.20} \\
\bottomrule
\end{tabular}
\end{table}

% ============================================================================
% SECTION 4b: LATENT USEFULNESS SCORING (v0.8)
% ============================================================================
\section{Latent Usefulness Scoring}
\label{sec:latent-usefulness}

\subsection{Motivation}

Relevance in scientific dataset search is intent-dependent. A dataset that perfectly matches a \emph{replication} query may be poorly suited for \emph{meta-analysis} even if both queries share surface-level keywords. Latent usefulness scoring formalizes this by defining seven researcher intents, scoring candidates along ten orthogonal dimensions, and weighting those dimensions per intent.

\subsection{Researcher Intent Taxonomy}

\begin{definition}[UsefulnessIntent]
Let $\mathcal{I}_U = \{I_1, \ldots, I_7\}$ be the researcher intent space:
\begin{align*}
    I_1 &= \texttt{STRICT\_LOOKUP} & &\text{(find exact dataset by ID/name)} \\
    I_2 &= \texttt{REPLICATION}    & &\text{(reproduce a specific experiment)} \\
    I_3 &= \texttt{META\_ANALYSIS} & &\text{(aggregate across multiple studies)} \\
    I_4 &= \texttt{PIPELINE\_REUSE}& &\text{(reuse existing analysis pipeline)} \\
    I_5 &= \texttt{CROSS\_DATASET\_COMPARISON} & &\text{(compare modalities or species)} \\
    I_6 &= \texttt{EXPLORATION}    & &\text{(discover datasets by topic)} \\
    I_7 &= \texttt{METHOD\_TRANSFER} & &\text{(apply method from one domain to another)}
\end{align*}
\end{definition}

Intent is classified rule-based via regex patterns and keyword heuristics with a confidence score $\gamma \in [0, 1]$.

\subsection{Ten-Dimension Usefulness Score}

\begin{definition}[DatasetContext]
A dataset candidate is represented as a context vector:
\begin{equation}
    C = (\text{modalities}, \text{tasks}, \text{species}, \text{regions}, \text{affordances}, \text{standards},
         N_s, N_t, N_u, b_{\text{ts}}, q)
\end{equation}
where $N_s, N_t, N_u$ are session/trial/subject counts, $b_{\text{ts}}$ indicates timestamp availability, and $q$ is a quality score.
\end{definition}

\begin{definition}[Operational Usefulness Scoring Function]
For query context $C_q$, candidate $C_d$, and intent $i \in \mathcal{I}_U$:
\begin{equation}
    U(C_q, C_d, i) = \sum_{k \in A} \tilde{w}_k(i) \cdot s_k(C_q, C_d)
\end{equation}
where $A$ is the active feature set and inactive dimensions are excluded. Effective weights are renormalized:
\begin{equation}
    \tilde{w}_k(i) = \frac{w_k(i)}{\sum_{j \in A} w_j(i)}.
\end{equation}
The dimension scores are:
\begin{align}
    s_1 &= \text{Jaccard}(\text{modalities}_q, \text{modalities}_d) \\
    s_2 &= \text{Jaccard}(\text{tasks}_q, \text{tasks}_d) \\
    s_3 &= \text{Jaccard}(\text{species}_q, \text{species}_d) \\
    s_4 &= \text{Jaccard}(\text{regions}_q, \text{regions}_d) \\
    s_5 &= \text{Jaccard}(\text{affordances}_q, \text{affordances}_d) \\
    s_6 &= \text{Jaccard}(\text{standards}_q, \text{standards}_d) \\
    s_7 &= f_{\log}(N_s) \\
    s_8 &= f_{\log}(N_t) \\
    s_9 &= \text{PathSim}(\cdot \mid \mathcal{G}) \\
    s_{10} &= \text{inactive}.
\end{align}
These correspond to modality, task, species, region, affordance, pipeline compatibility, session scale, trial scale, graph proximity, and neural-signature similarity respectively. The graph term uses a neutral prior if no graph is available, and neural-signature similarity is inactive until content extraction exists.
\end{definition}

The neural signature dimension $s_{10}$ is excluded from $A$ until content-level neural signature extraction is implemented. This prevents unimplemented features from lowering scores through non-zero intent weights.

The log-power scale function handles count fields with wide dynamic range:
\begin{equation}
    f_{\log}(n) = \begin{cases} 0.3 & n = 0 \text{ or None} \\ \min\!\left(1,\, \frac{\ln(1+n)}{\ln(1+10000)}\right) & \text{otherwise} \end{cases}
\end{equation}

Jaccard similarity does not reward missing metadata as similarity:
\begin{equation}
    \text{Jaccard}(A, B) = \begin{cases} 0 & A = B = \emptyset \\ 0 & A = \emptyset \text{ XOR } B = \emptyset \\ \frac{|A \cap B|}{|A \cup B|} & \text{otherwise} \end{cases}
\end{equation}
Missing fields are tracked as uncertainty and should contribute to result-card warnings, missingness penalties, and lower evidence confidence rather than increasing similarity.

\subsection{Intent-Specific Weight Profiles}

\begin{table}[htbp]
\centering
\caption{Base intent weight profiles before inactive-dimension renormalization.}
\small
\begin{tabular}{lrrrrrrr}
\toprule
Dimension & Lookup & Replication & Meta & Pipeline & Cross & Explore & Transfer \\
\midrule
Modality & 0.05 & 0.15 & 0.10 & 0.15 & 0.20 & 0.10 & 0.15 \\
Task     & 0.05 & 0.20 & 0.15 & 0.10 & 0.15 & 0.20 & 0.15 \\
Species  & 0.05 & 0.15 & 0.10 & 0.05 & 0.15 & 0.10 & 0.15 \\
Region   & 0.05 & 0.10 & 0.10 & 0.05 & 0.10 & 0.10 & 0.10 \\
Affordance & 0.05 & 0.05 & 0.15 & 0.20 & 0.10 & 0.15 & 0.15 \\
Standards & 0.05 & 0.05 & 0.05 & 0.20 & 0.05 & 0.05 & 0.05 \\
Scale-S  & 0.05 & 0.05 & 0.15 & 0.05 & 0.05 & 0.10 & 0.05 \\
Scale-T  & 0.05 & 0.05 & 0.10 & 0.05 & 0.05 & 0.05 & 0.05 \\
Graph    & 0.30 & 0.10 & 0.05 & 0.10 & 0.10 & 0.10 & 0.10 \\
Signature& 0.25 & 0.10 & 0.05 & 0.05 & 0.05 & 0.05 & 0.05 \\
\midrule
Sum      & 1.00 & 1.00 & 1.00 & 1.00 & 1.00 & 1.00 & 1.00 \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Graph-Derived Usefulness Signals}

Four graph-derived signals augment the usefulness scorer:

\begin{definition}[Affordance Overlap]
\begin{equation}
    \text{aff\_overlap}(d_i, d_j) = \frac{|\text{affordances}(d_i) \cap \text{affordances}(d_j)|}{|\text{affordances}(d_i) \cup \text{affordances}(d_j)|}
\end{equation}
\end{definition}

\begin{definition}[Pipeline Overlap]
\begin{equation}
    \text{pipe\_overlap}(d_i, d_j) = \frac{|\text{standards}(d_i) \cap \text{standards}(d_j)|}{|\text{standards}(d_i) \cup \text{standards}(d_j)|}
\end{equation}
\end{definition}

\begin{definition}[Complementarity Score]
\begin{equation}
    \text{complement}(d_i, d_j) = 0.7 \cdot \frac{|A_i \triangle A_j|}{|A_i \cup A_j|} + 0.3 \cdot \frac{|A_i \cap A_j|}{|A_i \cup A_j|}
\end{equation}
where $A_k = \text{affordances}(d_k)$. High unique fractions reward complementary pairs; modest overlap rewards shared foundation.
\end{definition}

\begin{definition}[Hub-Normalized PathSim]
For metapath $\mathcal{P}$ and shared-neighbor sets $N_\mathcal{P}(\cdot)$ derived from the knowledge graph:
\begin{equation}
    \text{PathSim}_\mathcal{P}(d_i, d_j) = \frac{2 \cdot |N_\mathcal{P}(d_i) \cap N_\mathcal{P}(d_j)|}{|N_\mathcal{P}(d_i)| + |N_\mathcal{P}(d_j)|}
\end{equation}
The self-exclusion guard $\{v \neq \text{src}\}$ prevents hub nodes from inflating denominator terms.
\end{definition}

% ============================================================================
% SECTION 5: COMPATIBILITY TENSOR
% ============================================================================
\section{Dataset Compatibility Theory}
\label{sec:compatibility}

\subsection{Slot-Based Dataset Representation}

\begin{definition}[Experimental Slot Vector]
A dataset $d$ is represented as a slot vector:
\begin{equation}
    E(d) = \begin{pmatrix}
        \text{organism}(d) \\
        \text{taxon\_group}(d) \\
        \text{preparation}(d) \\
        \text{task\_family}(d) \\
        \text{task\_variant}(d) \\
        \text{event\_schema}(d) \\
        \text{modality\_family}(d) \\
        \text{device}(d) \\
        \text{signal\_type}(d) \\
        \text{brain\_region}(d) \\
        \text{region\_hierarchy}(d) \\
        \text{species\_homolog}(d) \\
        \text{data\_standard}(d) \\
        \text{file\_format}(d) \\
        \text{sampling\_resolution}(d) \\
        \text{analysis\_affordances}(d) \\
        \text{readiness}(d) \\
        \text{provenance\_confidence}(d)
    \end{pmatrix}
\end{equation}
\end{definition}

\subsection{Compatibility Function}

\begin{definition}[Compatibility Scoring Model]
The compatibility between datasets $d_i$ and $d_j$ given query intent is:
\begin{equation}
\begin{split}
    \compatibility(d_i, d_j | \query) = &\sum_{k} w_k(\query) \cdot \text{match}_k(E_k(d_i), E_k(d_j)) \\
    &+ \lambda_g \cdot S_{\text{graph\_path}}(d_i, d_j) \\
    &+ \lambda_a \cdot S_{\text{analysis\_complement}}(d_i, d_j) \\
    &+ \lambda_p \cdot S_{\text{provenance\_agreement}}(d_i, d_j) \\
    &- \mu_c \cdot P_{\text{contradiction}}(d_i, d_j) \\
    &- \mu_m \cdot P_{\text{missing\_slot}}(d_i, d_j)
\end{split}
\end{equation}
\end{definition}

\subsection{Compatibility Types}

\begin{definition}[Compatibility Categories]
We define five compatibility relations:

\begin{enumerate}
    \item \textbf{Equivalent context} ($\equiv$): Same or homologous task, modality, species, region, and event schema. Useful for replication and meta-analysis.
    \begin{equation}
        d_i \equiv d_j \iff \forall k \in \mathcal{K}_{\text{core}}: \text{match}_k(E_k(d_i), E_k(d_j)) > \theta_{\text{equiv}}
    \end{equation}

    \item \textbf{Complementary context} ($\oplus$): Same task but different modality, or same region but different species.
    \begin{equation}
        d_i \oplus d_j \iff \exists k_1, k_2: E_{k_1}(d_i) = E_{k_1}(d_j) \land E_{k_2}(d_i) \neq E_{k_2}(d_j)
    \end{equation}

    \item \textbf{Translational context} ($\rightarrow$): Cross-species mapping via homologous behaviors and regions.

    \item \textbf{Analysis-compatible} ($\sim_A$): Both satisfy requirements for analysis $A$.
    \begin{equation}
        d_i \sim_A d_j \iff \text{affordance}_A(d_i) = \text{affordance}_A(d_j) = \text{True}
    \end{equation}

    \item \textbf{Contrastive context} ($\bot$): Intentionally differ on one axis for controlled comparison.
\end{enumerate}
\end{definition}

% ============================================================================
% SECTION 6: ANALYSIS AFFORDANCES
% ============================================================================
\section{Analysis Affordance Theory}
\label{sec:affordances}

\subsection{Affordance as Constraint Satisfaction}

\begin{definition}[Analysis Affordance]
An analysis affordance $A$ is a predicate over dataset features:
\begin{equation}
    \text{affordance}_A(d) = \bigwedge_{r \in \text{requirements}(A)} \text{satisfies}(d, r)
\end{equation}
\end{definition}

\begin{example}[Choice Decoding Affordance]
\begin{equation}
    \text{affordance}_{\text{choice\_decoding}}(d) = \text{has\_neural}(d) \land \text{has\_choice\_labels}(d)
\end{equation}
\end{example}

\begin{example}[Q-Learning Modeling Affordance]
\begin{equation}
    \text{affordance}_{\text{Q-learning}}(d) = \text{has\_choice}(d) \land \text{has\_reward}(d) \land \text{has\_trial\_outcome}(d)
\end{equation}
\end{example}

\subsection{Hyperedge Requirements}

Some analyses require conjunctions that cannot be represented as simple edges:

\begin{definition}[Requirement Hyperedge]
A hyperedge $h = (A, \{r_1, \ldots, r_n\})$ represents that analysis $A$ requires \textit{all} of $r_1, \ldots, r_n$ simultaneously:
\begin{equation}
    \text{satisfies}(d, h) = \bigwedge_{i=1}^n \text{has\_feature}(d, r_i)
\end{equation}
\end{definition}

\subsection{Registered Affordance Types (v1.1)}
\label{sec:affordance-registry}

Table~\ref{tab:affordances} enumerates the 21 registered analysis affordance types as of June~2026. Types are organised into four categories: core experimental, population dynamics, clinical/BCI, and extended behavioral.

\begin{table}[h]
\small
\centering
\begin{tabular}{llp{5.5cm}}
\toprule
\textbf{ID} & \textbf{Category} & \textbf{Required signals} \\
\midrule
\texttt{event\_aligned\_psth}        & Core experimental & spike times, event timestamps \\
\texttt{trial\_aligned\_neural}       & Core experimental & neural data, trial structure, event timestamps \\
\texttt{choice\_decoding}            & Core experimental & neural data, choice labels, trial structure \\
\texttt{stimulus\_response\_modeling} & Core experimental & neural data, stimulus presentation, timing \\
\texttt{behavioral\_state\_decoding} & Core experimental & neural data, behavioral state labels \\
\texttt{q\_learning}                  & Core experimental & trial structure, choice sequence, reward signal \\
\texttt{motor\_decoding}             & Core experimental & neural data, motor action labels \\
\texttt{delay\_discounting\_modeling} & Core experimental & choice sequence, delay variable, reward magnitude \\
\texttt{cross\_session\_generalization} & Core experimental & neural data, multiple sessions \\
\midrule
\texttt{dimensionality\_reduction}   & Population dynamics & neural population data, multiple units/voxels \\
\texttt{functional\_connectivity}    & Population dynamics & multi-channel neural data \\
\texttt{fmri\_glm\_analysis}         & Population dynamics & fMRI BOLD data, BIDS events \\
\texttt{trial\_aligned\_calcium}     & Population dynamics & calcium imaging, ROI traces, trial structure \\
\texttt{latent\_dynamics\_modeling}  & Population dynamics & neural population data, time series \\
\texttt{representational\_similarity\_analysis} & Population dynamics & neural population, task conditions \\
\midrule
\texttt{speech\_decoding}            & Clinical / BCI & neural data, speech events, temporal alignment \\
\texttt{seizure\_detection}          & Clinical / BCI & neural data, seizure annotations \\
\texttt{sleep\_stage\_classification} & Clinical / BCI & neural data, sleep stage labels \\
\texttt{bci\_decoding}               & Clinical / BCI & neural data, BCI context \\
\midrule
\texttt{cross\_area\_interaction}    & Extended behavioral & neural data, multiple brain regions, simultaneous recording \\
\texttt{pose\_neural\_correlation}   & Extended behavioral & neural data, pose tracking data \\
\bottomrule
\end{tabular}
\caption{The 21 registered analysis affordance types in Neural Search v1.1. Each affordance specifies required features, optional features, negative conditions, and validation methods. Support levels (HIGH/MEDIUM/LOW/UNSUPPORTED) depend on whether features are verified from structured NWB/BIDS data or inferred from metadata.}
\label{tab:affordances}
\end{table}

\subsection{Affordance Confidence}

\begin{equation}
    \confidence_A(d) = \prod_{r \in \text{requirements}(A)} \confidence(\text{evidence}(d, r))
\end{equation}

% ============================================================================
% SECTION 7: PROVENANCE AND CONFIDENCE
% ============================================================================
\section{Provenance and Confidence Modeling}
\label{sec:provenance}

\subsection{Evidence Labels}

\begin{definition}[Evidence Label]
An evidence label is a tuple:
\begin{equation}
    \ell = (\text{id}, \text{label}, \text{type}, \confidence, \text{evidence\_text}, \text{source\_field}, \text{extractor}, \text{version})
\end{equation}
\end{definition}

\subsection{Confidence Calibration}

Confidence scores should be calibrated such that:
\begin{equation}
    \mathbb{P}[\text{label correct} | \confidence = c] \approx c
\end{equation}

We measure calibration via reliability diagrams and Expected Calibration Error (ECE):
\begin{equation}
    \text{ECE} = \sum_{b=1}^B \frac{|B_b|}{N} \left| \text{acc}(B_b) - \text{conf}(B_b) \right|
\end{equation}
where $B_b$ are confidence bins, $\text{acc}(B_b)$ is accuracy within bin $b$, and $\text{conf}(B_b)$ is mean confidence.

\subsection{Confidence Source Hierarchy}

\begin{equation}
    \confidence_{\text{base}}(\ell) =
    \begin{cases}
        0.95 & \text{structured metadata field} \\
        0.85 & \text{curated synonym match} \\
        0.75 & \text{strong title pattern} \\
        0.60 & \text{description inference} \\
        0.40 & \text{path or filename evidence}
    \end{cases}
\end{equation}

\section{Field-State Memory for Scientific Retrieval Validation}
\label{sec:field-state-memory}

Neural Search now includes a Field-State validation layer that records what the system claims, what evidence exists, and what benchmark gaps remain. This layer is intentionally lightweight: generated truth remains in JSONL artifacts under \texttt{artifacts/field\_state/} and Markdown reports under \texttt{reports/field\_state/}. The memory layer does not ingest papers, parse PDFs, call external APIs, or replace human evaluation.

\subsection{Claims, Gaps, and Opportunities}

Field-State records three first-class artifact types. A \texttt{FieldClaim} tracks a scientific or evaluation claim, including evidence level, confidence, related artifacts, missing tests, and status. A \texttt{BenchmarkGap} tracks validation work that is still required before a claim can be defended. A \texttt{FieldOpportunity} tracks high-value next work and is ranked by a transparent heuristic score:

\begin{equation}
\begin{split}
    \text{score} ={}& 0.20\,s_{\text{novelty}} + 0.25\,s_{\text{feasibility}}
    + 0.20\,s_{\text{impact}} \\
    &+ 0.15\,s_{\text{uncertainty}} + 0.15\,s_{\text{fit}}
    - 0.10\,s_{\text{risk}} .
\end{split}
\end{equation}

The seeded Field-State ledger currently tracks claims about dense retrieval, hard negatives, human qrels, metadata richness, analysis affordance extraction, and graph proximity. These claims are not automatically upgraded by the presence of infrastructure. Evidence suggestions are written separately so that human review can decide whether a claim should change.

\subsection{Obsidian Review Memory}

The Obsidian memory base is a local Markdown review spine. Every generated note has stable \texttt{field\_state\_id} frontmatter and two explicit blocks:

\begin{lstlisting}
<!-- FIELDSTATE:BEGIN generated -->
Machine-generated content.
<!-- FIELDSTATE:END generated -->

<!-- FIELDSTATE:BEGIN human -->
Human review notes.
<!-- FIELDSTATE:END human -->
\end{lstlisting}

On re-export, Neural Search replaces only the generated block and preserves the human block and safe human-edited frontmatter fields such as \texttt{review\_status}, \texttt{human\_reviewer}, \texttt{reviewed\_at}, \texttt{human\_priority}, and \texttt{human\_tags}. Imported review overlays are written to separate JSONL files; they annotate reports and snapshots but do not mutate the generated claim, gap, opportunity, or qrels artifacts.

\subsection{Evaluation Memory and Qrels Review}

The evaluation-memory layer connects retrieval candidates to human qrels review without treating generated labels as gold. Candidate query--dataset pairs are exported as Obsidian qrels notes. Human reviewers can enter 0--3 graded relevance and usefulness scores, hard-negative violation flags, label confidence, annotator identifiers, rationales, and reviewer notes. Imports write \texttt{artifacts/field\_state/qrels\_reviews.jsonl}; adjudication writes \texttt{artifacts/field\_state/adjudicated\_qrels.jsonl}; agreement summaries and evaluation snapshots are written to \texttt{reports/field\_state/}.

The intended validation loop is:

\begin{center}
\small
\texttt{Retrieval candidates}
$\rightarrow$ \texttt{Obsidian qrels review}
$\rightarrow$ \texttt{human graded labels}
$\rightarrow$ \texttt{adjudicated qrels}
$\rightarrow$ \texttt{metrics and agreement}
$\rightarrow$ \texttt{eval snapshot}
$\rightarrow$ \texttt{claim evidence update}
$\rightarrow$ \texttt{whitepaper validation report}.
\end{center}

As of the current artifact snapshot, this human-review loop has been smoke-tested and remains separate from the larger canonical LLM-judged qrels snapshot. The smoke test demonstrates artifact integrity and review-loop feasibility, while the 317-query LLM qrels snapshot exercises retrieval metrics and the regression gate. Neither constitutes final publication-grade human relevance validation until duplicate review and adjudication are complete.

\subsection{Field-State Memory Graph (v0.9)}
\label{sec:field-state-memory-graph}

The v0.9 sprint replaces the flat JSONL field-state layer with a \textbf{Field-State Memory Graph}: a typed, directed, incrementally-updatable relational graph that stores every first-class artifact as a node and every relationship as a named, provenance-carrying edge.

\subsubsection{Graph Schema}

The graph is defined in \texttt{neural\_search/graph/schema.py} and implemented in \texttt{neural\_search/field\_state/}. Nodes are instances of \texttt{KnowledgeGraphNode}; edges are instances of \texttt{KnowledgeGraphEdge}. Every node carries a stable deterministic ID (\texttt{node:<type>:<slug>}), label, aliases, source IDs, properties, evidence list, confidence float, and created/updated timestamps. Every edge carries a stable ID (\texttt{edge:<type>:<source>:<target>}), evidence, confidence, and properties.

Node types include: \texttt{dataset}, \texttt{source\_archive}, \texttt{paper}, \texttt{modality}, \texttt{species}, \texttt{brain\_region}, \texttt{task}, \texttt{behavioral\_event}, \texttt{data\_standard}, \texttt{file\_format}, \texttt{analysis\_affordance}, \texttt{raw\_data\_signal}, \texttt{processed\_data\_signal}, \texttt{concept}, \texttt{pipeline}, \texttt{query}, \texttt{retrieval\_run}, \texttt{neuro\_judge\_evidence\_packet}, \texttt{neuro\_judge\_judgment}, \texttt{feedback\_signal}, \texttt{curation\_issue}, and \texttt{snapshot\_manifest}.

\subsubsection{Provenance Invariants}

The graph enforces strict provenance separation at the field level. Neuro-judge outputs are stored with \texttt{provenance: "neuro\_judge\_silver\_not\_human\_gold"} and are rejected at ingestion time if submitted with \texttt{provenance: "human\_gold"}. User feedback events are stored with \texttt{provenance: "user\_feedback\_downstream\_signal"}. No LLM label or feedback event can be promoted to gold label status through any pipeline path.

\subsubsection{Conservative Normalizers}

The \texttt{neural\_search/field\_state/normalizers.py} module extracts species, modality, brain region, and raw/processed data evidence with explicit disambiguation guardrails:

\begin{itemize}
    \item Calcium imaging does \emph{not} satisfy extracellular spike sorting affordances.
    \item fMRI does \emph{not} satisfy LFP or spike affordances.
    \item MT/V5 is \emph{not} interchangeable with lateral intraparietal cortex.
    \item Entorhinal cortex is \emph{not} equivalent to hippocampus unless marked as related (not equivalent).
    \item Human imaging is \emph{not} interchangeable with rodent electrophysiology.
    \item Raw data availability is \emph{not} inferred from archive membership or standard alone; it requires positive keyword evidence (e.g., raw AP bin files, continuous voltage traces).
\end{itemize}

Text-inferred labels receive confidence $\leq 0.6$; structured metadata labels receive confidence $\geq 0.7$.

\subsubsection{Corpus Enrichment and Regional Coverage Management}

Corpus growth is handled as a controlled evidence-management process rather than as indiscriminate scraping. The expansion layer admits high-volume sources only when records satisfy four inclusion gates: evidence of raw or processed data, a neuroscience signal, an open or usable license, and a persistent identifier or stable source identifier. Source-specific adapters normalize records from DANDI, OpenNeuro, NeuroVault, Zenodo, Figshare, NeuroMorpho, Allen, GIN, Brain Image Library, BlueBrain, IBL, CRCNS, OSF, and smaller sources into the same schema. Deduplication is performed by stable \texttt{dataset\_id}; rejected Tier-2 candidates are logged rather than silently discarded.

The current expansion also adds targeted enrichment passes for metadata dimensions that are weak in repository-level records. DANDI records can be refreshed from rich metadata fields such as \texttt{assetsSummary.variableMeasured}, \texttt{assetsSummary.approach}, \texttt{assetsSummary.species}, and \texttt{studyTarget}. NWB-specific extractors can stream representative assets to read electrode-table \texttt{location} fields and surgery or experiment-description text. Paper-linked records can be enriched through DOI extraction, CrossRef abstract retrieval, and rule-based region extraction. These enrichments are marked as silver provenance, for example \texttt{nwb\_surgery\_text\_silver\_not\_human\_gold} or \texttt{crossref\_abstract\_rule\_based\_silver\_not\_human\_gold}, and therefore cannot be promoted to gold labels without human review.

Regional memory is managed through a separate coverage-map artifact. The regional map distinguishes verified region labels from candidate-only textual mentions, summarizes coverage by source, species, modality, and anatomical system, and emits a regionless review queue for acquisition and curation. The current regional map covers 625 curated/depth records, with 223 records containing verified regions (35.7\%), 94 atlas target regions, 43 verified targets, 16 candidate-only targets, and 35 uncovered targets. This is used as a curation steering signal, not as a final anatomical benchmark.

\subsubsection{Incremental Update Pipeline}

The \texttt{update\_field\_state.py} script detects corpus record changes by SHA-256 content hash over $\{\text{title}, \text{description}, \text{source\_id}, \text{source}\}$. Each run writes a versioned snapshot directory under \texttt{artifacts/field\_state/snapshots/<timestamp>/} containing:
\begin{itemize}
    \item \texttt{corpus\_manifest.json} --- record hashes and change counts
    \item \texttt{memory\_graph\_manifest.json} --- node/edge counts by type
    \item \texttt{index\_manifest.json} --- which stages were skipped
    \item \texttt{update\_report.md} --- human-readable summary
\end{itemize}

The \texttt{compare\_snapshots.py} script computes node/edge count deltas by type between any two snapshots and warns on $>10\%$ unexpected removals.

\subsubsection{Review Overlays, Memory Indexing, and Diff Reports}

Human review is stored as an overlay on top of generated artifacts. The Obsidian mirror preserves machine-generated and human-authored blocks separately, imports reviewer status, priority, tags, and notes into overlay JSONL files, and then annotates generated records at report time without mutating the canonical claim, gap, opportunity, or qrels artifacts. A memory index records a SHA-256 hash of each full Markdown note, including human content, so the system can detect human edits, schema mismatches, missing markers, duplicate identifiers, notes missing from the index, and indexed notes missing from the vault. The resulting memory diff report makes review drift explicit and audit-ready.

\subsubsection{Learning Signals}

The \texttt{LearningSignalGenerator} in \texttt{neural\_search/field\_state/learning\_signals.py} derives provisional signals from feedback aggregation and neuro-judge disagreements. Signal types include: \texttt{false\_high\_candidate}, \texttt{false\_low\_candidate}, \texttt{missing\_metadata\_candidate}, \texttt{hard\_negative\_candidate}, \texttt{concept\_link\_candidate}, \texttt{affordance\_link\_candidate}, \texttt{source\_quality\_warning}, \texttt{query\_failure\_mode}, and \texttt{dataset\_reuse\_success\_signal}. All signals are provisional---they update priority queues and priors for review, but do not overwrite canonical labels or generate new qrels entries.

\subsubsection{Current Artifact State}

\paragraph{Full-corpus knowledge graph (rebuilt 30~June~2026).}
\texttt{data/graph/neural\_search\_graph.real\_corpus.json} was rebuilt from all 7,171 flat corpus records using \texttt{scripts/build\_real\_corpus\_graph.py}. The graph contains \textbf{7,438 nodes} and \textbf{81,518 edges}. Node type breakdown: 7,171 dataset, 118 brain\_region, 85 task, 46 modality, and 18 species. Edge type breakdown: \texttt{dataset\_has\_modality} (8,083), \texttt{dataset\_records\_region} (7,098), \texttt{dataset\_has\_species} (5,393), \texttt{dataset\_has\_task} (2,423), \texttt{same\_region\_cross\_modality} (48,212), and \texttt{same\_task\_cross\_species} (10,309). Each dataset node now uses the real corpus ID pattern \texttt{node:dataset:\{source\}:\{source\_id\}} and includes aliases such as \texttt{dandi:000785}, \texttt{000785}, and \texttt{dataset:dandi:000785}. This aligns graph node identity with retrieval run IDs and restores a nonzero graph-context contribution for real corpus records.

\paragraph{Field-state memory graph (evidence and curation layer).}
The \texttt{update\_20260613T053153Z} build covers the 629-record reviewed slice: 2,200 nodes and 3,788 typed edges, including 675 neuro-judge evidence-packet nodes, 618 raw-data-signal nodes, 22 paper nodes, and 17 feedback-signal nodes. This graph tracks curation state, evidence packets, and feedback; it is complementary to the full-corpus knowledge graph above.

\paragraph{Ontology coverage.}
The brain region ontology now covers 106 canonical region IDs with UBERON, Allen CCF v3 (mouse/rat), and Allen Human Atlas cross-references. The species taxonomy covers 10 entries with NCBITaxon identifiers. The task atlas covers 87 tasks mapped to Cognitive Atlas identifiers.

% ============================================================================
% SECTION 7b: COVERAGE LEDGER AND RETRIEVAL POST-PROCESSING
% ============================================================================
\section{Coverage Ledger, Gap Boosting, and Source Diversity}
\label{sec:coverage-ledger}

\subsection{DuckDB Coverage Ledger}

Dense retrieval alone does not distinguish between a brain region that is well-represented in the corpus and one that has only a single dataset. Without explicit coverage tracking, underrepresented anatomical or experimental contexts are systematically deprioritized by frequency-weighted scoring. The Neural Search coverage ledger addresses this with a persistent analytical database.

\begin{definition}[Coverage Ledger]
Let $\mathcal{C}$ be the corpus of $N$ datasets. The coverage ledger $\mathcal{L}$ is a 6-tuple:
\begin{equation}
    \mathcal{L} = (\mathcal{C},\; D_{\text{brain\_region}},\; D_{\text{modality}},\; D_{\text{recording\_scale}},\; D_{\text{species}},\; D_{\text{task}},\; D_{\text{behavioral\_event}})
\end{equation}
where each $D_{\text{dim}}$ is a mapping from label value to dataset identifier. The current ledger (15~June~2026) contains 7,176 datasets and 29,561 entries across six dimensions: brain region coverage 48.1\% (3,453/7,176 datasets), modality coverage 81.9\%, recording scale 80.6\%, species 72.8\%, and task 23.4\%. The task dimension is the primary coverage gap: many recording datasets carry no explicit task structure, and the 23.4\% rate counts passive recordings classified as \texttt{visual\_stimulation}, \texttt{resting\_state}, and \texttt{spontaneous\_activity} alongside active behavioral tasks.
\end{definition}

The ledger supports three analytical queries:
\begin{itemize}
    \item \textbf{Coverage rates by source}: fraction of records per archive with at least one label per dimension.
    \item \textbf{Dark pairs}: dimension value pairs $(v_1, v_2)$ with no co-occurring dataset (e.g., retina $\times$ fMRI = 0 datasets).
    \item \textbf{Gap matrix}: per-dimension coverage sorted ascending to surface rarest labels for acquisition targeting.
\end{itemize}

\subsection{Coverage Gap Boost}

\begin{definition}[Rarity Boost]
For a dataset $d$ with label set $L_d$ across dimension $k$, the rarity score is:
\begin{equation}
    r_k(d) = \frac{1}{\log_2(1 + \text{count}_k(L_d))}
\end{equation}
where $\text{count}_k(\ell)$ is the number of corpus datasets sharing label $\ell$ in dimension $k$. The additive gap boost is:
\begin{equation}
    B_{\text{gap}}(d) = \min\!\left(\text{GAP\_BOOST\_MAX},\; \frac{1}{|L_d|} \sum_{\ell \in L_d} r_k(\ell)\right)
\end{equation}
with $\text{GAP\_BOOST\_MAX} = 0.10$.
\end{definition}

The boost is additive and appears as a named component in the score breakdown, making it auditable per result. It is computed from the DuckDB ledger at query time with a per-dimension LRU cache.

\subsection{Source Diversity Post-Processor}

High-volume repositories (e.g., NeuroVault with 2,000 records) can dominate top-$k$ results for broad queries. A source diversity post-processor addresses this with a greedy windowed algorithm:

\begin{definition}[Source Diversity Reranking]
Given a ranked list $R = [d_1, d_2, \ldots, d_n]$ and source function $\sigma(d)$, the diversity-reranked list $R^*$ is built greedily:
\begin{equation}
    R^* = \text{greedyDiversify}(R,\; \text{max\_per\_source} = 3)
\end{equation}
where each result is placed in the accept queue if $|\{d' \in R^*_{\text{accepted}} : \sigma(d') = \sigma(d)\}| < \text{max\_per\_source}$, and deferred otherwise. Deferred results are appended after all accepted results.
\end{definition}

This is applied after SPECTER2 fusion and before the final $k$-truncation, ensuring that score-ordered diversity is preserved.

\subsection{SPECTER2 Hybrid Fusion (Config-Gated)}

Scientific literature embeddings from SPECTER2~\citep{specter2_2023} encode the conceptual space of neuroscience papers. A config-gated module augments retrieval scores with paper embedding similarity:

\begin{equation}
    S_{\text{specter2}}(d, q) = \cos\!\left(\mathbf{e}_{\text{specter2}}(d),\; \mathbf{e}_{\text{specter2}}(q)\right) \cdot w_{\text{specter2}}
\end{equation}

where $w_{\text{specter2}} = 0.20$ by default. SPECTER2 corpus embeddings are pre-computed and LRU-cached; query embeddings are computed on demand. The module is disabled by default and must be explicitly enabled in \texttt{data/config/retrieval.yaml}.

\subsection{LLM Query Expansion (Fallback)}

When rule-based query parsing extracts zero brain regions and zero tasks from a query string, a fallback LLM expansion fires to recover structured terms. The expansion:
\begin{enumerate}
    \item Calls Claude Haiku with a structured JSON-output prompt specifying $\{$\texttt{brain\_regions}, \texttt{tasks}, \texttt{modalities}, \texttt{species}$\}$.
    \item Merges returned terms with any existing parsed terms (deduplication-preserving order).
    \item Fails safe: any exception returns the original parsed query unchanged.
    \item Is LRU-cached (512 entries) so repeated identical queries are not billed.
\end{enumerate}

The expansion sets \texttt{llm\_expansion\_applied: True} in the parsed query for downstream auditability. It is disabled by default and enabled via \texttt{llm\_expansion.enabled: true} in the retrieval config.

% ============================================================================
% SECTION 8: EVALUATION METHODOLOGY
% ============================================================================
\section{Evaluation Methodology}
\label{sec:evaluation}

\subsection{Ranking Metrics}

\subsubsection{Precision at K}

\begin{definition}[Precision@K]
\begin{equation}
    P@K = \frac{|\text{relevant} \cap \text{top-}K|}{K}
\end{equation}
\end{definition}

\subsubsection{Recall at K}

\begin{definition}[Recall@K]
\begin{equation}
    R@K = \frac{|\text{relevant} \cap \text{top-}K|}{|\text{relevant}|}
\end{equation}
\end{definition}

\subsubsection{Mean Reciprocal Rank}

\begin{definition}[MRR]
For queries $Q$:
\begin{equation}
    \MRR = \frac{1}{|Q|} \sum_{q \in Q} \frac{1}{\text{rank}_q(\text{first relevant})}
\end{equation}
\end{definition}

\subsubsection{Normalized Discounted Cumulative Gain}

\begin{definition}[NDCG@K]
\begin{equation}
    \text{DCG}@K = \sum_{i=1}^K \frac{2^{\rel_i} - 1}{\log_2(i + 1)}
\end{equation}
\begin{equation}
    \NDCG@K = \frac{\text{DCG}@K}{\text{IDCG}@K}
\end{equation}
where IDCG is the ideal (maximum possible) DCG.
\end{definition}

\subsubsection{Graded Relevance for Usefulness Evaluation (v0.8)}

Standard binary relevance conflates degrees of usefulness. We define a 4-level graded gain function:

\begin{definition}[Graded Usefulness Gain]
\begin{equation}
    \text{GAIN}(\ell) = \begin{cases}
        0 & \ell = \texttt{not\_useful} \\
        1 & \ell = \texttt{weakly\_useful} \\
        2 & \ell = \texttt{useful} \\
        3 & \ell = \texttt{highly\_useful}
    \end{cases}
\end{equation}
\end{definition}

This enables NDCG to distinguish between a result that is merely relevant and one that is highly reusable.

\subsubsection{Hard-Negative Violation Rate}

\begin{definition}[HN-VIOL]
Let $\mathcal{H}$ be the set of hard-negative candidates and $r_1$ be the rank of the first relevant result:
\begin{equation}
    \text{HN-VIOL} = \frac{|\{h \in \mathcal{H} : \text{rank}(h) < r_1\}|}{|\mathcal{H}|}
\end{equation}
A system with $\text{HN-VIOL} = 0$ never surfaces a hard-negative above the first relevant result.
\end{definition}

\subsubsection{Mean Reciprocal Rank (Relevance-Thresholded)}

For usefulness evaluation, MRR counts only labels $\ell \in \{\texttt{useful}, \texttt{highly\_useful}\}$ as relevant, preventing \texttt{weakly\_useful} labels from inflating the score.

\subsection{Current Corpus and Validation Results}

This section reports only artifacts aligned with the current repository state. Retrieval metrics from smaller historical corpora have been removed from the main paper; they should remain in archived reports until a new benchmark is run on a frozen expanded-corpus snapshot.

\noindent\textbf{Artifact reconciliation note.} Some generated reports still describe an earlier 10,404-record target snapshot. The live corpus artifact inspected for this paper update is \texttt{data/corpus/normalized/combined\_corpus.jsonl/full\_corpus\_v09.jsonl}, which contains 7,171 normalized rows and 7,121 unique record identifiers. The current dense field embedding cache has shape $7{,}171 \times 1024$ in \texttt{real\_all.dense.field\_embeddings.agg.mat.npy}; the older on-disk turbovec vector file still has shape $625 \times 1024$ and should be rebuilt before approximate-index latency or recall is cited as current full-corpus evidence.

\noindent\textit{The generated table bundle \texttt{generated/current\_results\_tables.tex} currently contains stale 10,404-record numbers and should be regenerated before use. The compact tables below preserve the current reported artifact state.}

    \begin{table}[htbp]
    \centering
    \caption{Current corpus and index artifact summary (as of 30~June~2026).}
    \small
    \begin{tabular}{lc}
    \toprule
    Metric & Value \\
    \midrule
    Unique normalized dataset rows & 7,171 \\
    Unique normalized dataset ids & 7,121 \\
    Dense field embedding cache & 7,171 records \\
    Embedding model & BGE-large-en-v1.5 \\
    Embedding dimension & 1,024 \\
    Knowledge graph nodes & 7,438 \\
    Knowledge graph edges & 81,518 (incl.\ 58,521 cross-dataset) \\
    Coverage ledger datasets & 7,176 \\
    Coverage ledger entries & 29,561 \\
    Indexed ids (turbovec, requires rebuild) & 625 (stale — rebuild pending) \\
    Vector index recall@50 (prior snapshot) & 1.0 \\
    Tier-2 rejection summary & 24,160 (CSV report) \\
    \midrule
    \multicolumn{2}{l}{\textit{Literature layer (added 17~June~2026)}} \\
    OpenAlex papers ingested (tier~1, $\geq$100 citations) & 255,940 \\
    Paper-dataset linker input records & 7,171 \\
    Successful paper-dataset links & 393 (168 DOI exact; 225 title fuzzy local) \\
    Unlinked dataset records & 6,778 \\
    Finding extraction model & qwen2.5:7b-instruct-q4\_K\_M (Ollama) \\
    Finding extraction yield & 74.6\% (16,150 papers processed; 12,050 findings) \\
    Typed-finding enrichment & 27 fields over 122K findings; highest coverage among new fields: molecular marker 4.1\%, metabolic context 3.3\%, pharmacological agent 3.0\% \\
    Expected tier~1 findings if maintained & $\sim$190K \\
    Tier~2 papers staged (open-access, with abstract) & $\sim$1.39M \\
    \bottomrule
    \end{tabular}
    \end{table}

    \begin{table}[htbp]
    \centering
    \caption{Current validation artifacts.}
    \small
    \begin{tabular}{lc}
    \toprule
    Artifact & Value \\
    \midrule
    Usefulness labeled pairs & 270 \\
    Usefulness Spearman r & 0.3999 \\
    Mean useful / irrelevant score & 0.2244 / 0.1678 \\
    Graph pairs changed & 39.0\% \\
    Expanded-corpus retrieval benchmark & Pending benchmark artifact \\
    \bottomrule
    \end{tabular}
    \end{table}

\noindent\textit{The vector-index recall result validates the approximate vector index against exact nearest-neighbor retrieval under the current test configuration. It does not establish scientific relevance or usefulness. The next validation requirement is retrieval benchmarking on this exact snapshot, with exact-lookup checks, source-skew analysis, and independent relevance labels.}

\begin{table}[htbp]
\centering
\caption{Per-source corpus quality artifacts required before publication claims.}
\small
\begin{tabular}{p{0.37\linewidth}p{0.55\linewidth}}
\toprule
Artifact & Current status \\
\midrule
Per-source field completeness & Generated in \texttt{reports/eval/field\_completeness.csv} \\
Per-source DOI/PID coverage & Generated in \texttt{reports/eval/pid\_license\_coverage.csv} \\
Per-source license coverage & Generated in \texttt{reports/eval/pid\_license\_coverage.csv} \\
Per-source modality extraction precision & Pending manual audit artifact \\
Per-source species extraction precision & Pending manual audit artifact \\
Per-source task extraction precision & Pending manual audit artifact \\
Top Tier-2 rejection reasons & Generated in \texttt{reports/eval/rejection\_summary.csv} \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Current Usefulness Score Calibration}

Running 30 real-corpus benchmark queries through the usefulness scorer and measuring Spearman rank correlation between \texttt{usefulness\_score.total\_score} and benchmark relevance labels (270 pairs):

\begin{itemize}
    \item Spearman $r$ = 0.3999 ($p < 0.001$, $n = 270$ pairs)
    \item Mean usefulness score for relevant results: 0.2244
    \item Mean usefulness score for irrelevant results: 0.1678
\end{itemize}

The positive correlation indicates that the 10-dimension usefulness scorer assigns higher scores to relevant results on average. This is current preliminary evidence of discriminative signal, not final validation of downstream scientific utility. Independent labels, larger query coverage, and content-derived affordance labels are required before making strong publication claims.

\subsection{Ablation Analysis}

\begin{definition}[Ablation Study]
For signal set $\mathcal{S}$, the ablation of signal $s$ measures:
\begin{equation}
    \Delta_s = \text{metric}(\mathcal{S}) - \text{metric}(\mathcal{S} \setminus \{s\})
\end{equation}
\end{definition}

The current ablation artifacts include both graph-rank perturbation diagnostics and a silver LLM-qrels retrieval preview. A publishable ablation should be rerun after a human-adjudicated benchmark snapshot is frozen and labeled.

\subsection{Graph Proximity Ablation}
\label{sec:s9ablation}

The current graph proximity ablation confirms that the $s_9$ signal changes rank orderings, and the calibrated graph rung shows a small silver LLM-qrels-backed metric gain. This remains preliminary until human-adjudicated qrels are available:

\begin{itemize}
    \item Ranking pairs changed when real graph enabled vs.\ ablated: \textbf{39\%}
    \item This confirms the graph proximity signal is non-trivial and reshapes result ordering for a substantial fraction of queries.
    \item The June 30 graph rebuild addressed a namespace mismatch where retrieval runs used IDs such as \texttt{dandi:000785} while the old graph used demo-era dataset nodes. Real corpus graph nodes now use \texttt{node:dataset:\{source\}:\{source\_id\}} with alias fallback.
    \item The ablation should be rerun against a frozen expanded-corpus snapshot with human-adjudicated labels.
\end{itemize}

\subsection{Current Expanded-Corpus Evaluation Status}
\label{sec:v2eval}

The generated artifact tables summarize the current repository state after corpus expansion and the upgrade to BGE-large-en-v1.5 field embeddings. They also expose the current manifest reconciliation requirement: the expanded corpus, embedding cache, index metadata, and recall report must be regenerated as a single frozen bundle before any end-to-end retrieval claim is made.

\noindent\textbf{Note on usefulness calibration:} The BGE-large scorer shows a relevant-vs-irrelevant score gap (0.2244 vs.\ 0.1678). This supports the usefulness scorer as a current signal, while still requiring expanded-snapshot retrieval evaluation and independent labels.

\subsection{Engineering and Artifact Validation}

The Field-State, Obsidian memory, and evaluation-memory infrastructure is currently covered by targeted engineering validation. The repository passes focused tests for schema validation, JSONL read/write, report generation, Obsidian export/import roundtrips, human-block preservation, memory validation, memory diffs, qrels candidate export, qrels review import, adjudication logic, agreement summaries, evaluation snapshots, claim-evidence suggestions, and whitepaper validation reports. These tests validate artifact integrity and reproducibility of the review workflow. They do not validate ranking quality or scientific usefulness.

\subsection{Human Relevance Labeling Protocol}

Human relevance validation uses 0--3 graded labels for each query--dataset pair. The evaluation-memory layer exports qrels candidates to review notes and imports human labels into \texttt{artifacts/field\_state/qrels\_reviews.jsonl}. Adjudicated labels are written separately to \texttt{artifacts/field\_state/adjudicated\_qrels.jsonl}. The current canonical LLM-judged qrels snapshot is useful for ablation preview and regression testing, but human relevance validation still requires duplicate review, disagreement adjudication, and agreement reporting.

\noindent\textbf{Neuro-judge labels are preliminary silver labels.} The repository includes a RAG-grounded neuro-judge workflow for prioritizing query--dataset pairs for review and for diagnosing evidence gaps. These labels are not human gold. The mock neuro-judge backend is a deterministic plumbing and regression-test instrument, not a scientific relevance labeler. Judge-derived metrics must remain diagnostic until collapse diagnostics, stratified validation sampling, expert audit, and calibration against adjudicated human labels are complete.

\noindent\textbf{Neuro-judge v2 prompt.} The judge prompt was upgraded to PROMPT\_VERSION v2. Changes include: (1) description evidence extended from 600 to 1,200 characters, giving the judge access to more experimental detail; (2) a structured \emph{reasoning trace} field requiring the judge to work through five explicit steps (hard-negative check, dimension-by-dimension check, raw-data availability check, evidence completeness estimate, final score) before committing to a label---empirically improving accuracy on borderline cases; (3) a \emph{metadata signal quality} section distinguishing explicitly stated metadata fields from signals inferred by keyword heuristics, allowing the judge to calibrate confidence accordingly; and (4) tightened abstain calibration---a label of 3 with no \texttt{required\_dimensions\_present} now also triggers an abstain recommendation, not only the previous \texttt{label $\geq$ 2 + missing dimensions} condition.

\noindent\textbf{Frontend feedback loop and downstream retrieval success reporting.} The existing React/Vite frontend has been extended with: evidence inspection panels on every result card (showing evidence for/against, matched dimensions, affordance matches, concept-memory signals, failure modes, and raw evidence JSON); neuro-judge badges with provenance watermarks distinguishing silver from expert-audited labels; sorting and filtering by judge label, confidence, evidence completeness, abstain status, modality, species, source, and feedback score; and inline feedback controls capturing usefulness, would-use-for-analysis, reason tags, and free-text notes. Feedback is written to \texttt{artifacts/frontend/retrieval\_feedback.jsonl} with provenance \texttt{user\_feedback\_downstream\_signal}. A downstream retrieval success report summarising usefulness by rank bucket, method, judge label, and evidence completeness is regenerated with \texttt{scripts/eval/compute\_downstream\_retrieval\_success.py}. A feedback audit queue (\texttt{build\_feedback\_audit\_queue.py}) and transparent reranking prior layer (\texttt{build\_feedback\_rerank\_priors.py}) close the human-in-the-loop triage cycle. Dataset organisation views grouping the judged corpus by relevance tier, modality, and metadata quality are generated by \texttt{build\_dataset\_organization\_views.py}. These are downstream usage signals; they supplement but do not replace expert qrels annotation.

\subsection{Hard-Negative Evaluation}

The benchmark design treats hard negatives as scientific failures rather than ordinary low-relevance results. The current infrastructure records hard-negative rationales in qrels notes and supports future hard-negative review notes under \texttt{Field-State/60\_Evaluation/hard\_negatives/}. However, a qrels-backed hard-negative violation rate has not yet been computed for the expanded corpus. Any claim that hard-negative handling improves retrieval quality should remain caveated until reviewed examples and violation-rate reports exist.

\subsection{Calibration and Source-Skew Analysis}

Calibration and source skew remain planned scientific retrieval validations. Expected Calibration Error (ECE) is the target metric for whether predicted usefulness scores reflect empirical relevance probability. Source-skew reports are required because high-volume sources such as Zenodo and Figshare can dominate candidate pools. The current reports include source distribution and corpus manifests, but a qrels-backed calibration report and source-skew metric report are still pending.

\subsection{Claim-Evidence Tracking}

Field-State generates claim evidence suggestions rather than automatically modifying claims. Current suggestions keep dense retrieval, hard-negative, affordance extraction, metadata richness, and graph proximity claims caveated until qrels-backed metrics, hard-negative reviews, content-validated affordance labels, metadata-quality analyses, or graph ablations are available. The system is designed so every whitepaper claim can be traced to an artifact, metric, human label, or explicit future-validation item.

\begin{table}[htbp]
\centering
\caption{Claim ledger. Infrastructure existence is separated from scientific validation and publication readiness.}
\label{tab:claim-ledger}
\scriptsize
\begin{tabular}{p{0.22\linewidth}p{0.15\linewidth}p{0.22\linewidth}p{0.27\linewidth}p{0.08\linewidth}}
\toprule
Claim & Status & Evidence artifact & Missing validation & Publishable now? \\
\midrule
Corpus contains 7,171 normalized dataset rows & Implemented & \texttt{full\_corpus\_v09.jsonl} & Freeze manifest with commit/hash metadata & Yes \\
Dense vector index supports approximate nearest-neighbor lookup & Partial/stale & Turbovec index metadata and prior recall report & Rebuild embeddings, index, and recall report as one frozen bundle & Cautious \\
Usefulness scorer separates useful from irrelevant examples & Preliminary & 270 labeled pairs; Spearman $r=0.3999$ & Independent labels, larger query coverage, calibration & Cautious \\
Graph proximity improves retrieval quality & Unproven & 39\% rank perturbation artifact & Qrels-backed graph ablation on frozen snapshot & No \\
Paper-dataset linker covered the corpus & Implemented run & \texttt{paper\_dataset\_links.jsonl} & Precision audit; distinguish 393 successful links from 6,778 not-found records & Cautious \\
Finding extraction yields reusable structured findings & Operational estimate & 16,150 papers processed; 12,050 findings; 74.6\% yield; 27-field enrichment coverage over 122K findings & Manual finding audit against abstracts; ontology-field and typed-field precision/recall & No \\
Analysis affordance labels support recommendations & Implemented representation & 21-type registry and frontend panel & Content-level affordance precision audit & No \\
LLM qrels support final retrieval metrics & Silver only & LLM qrels and neuro-judge evidence packets & Human-adjudicated qrels for NDCG/MRR/Recall & No \\
Human qrels workflow is reproducible & Implemented workflow & qrels export/import/adjudication artifacts; three adjudicated labels & Benchmark-scale labels, agreement stats, metric report & Cautious \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Engineering Versus Scientific Retrieval Validation}
\label{sec:eng-vs-science}

Neural Search has passed engineering validation: ingestion, deduplication, embeddings, vector indexing, graph construction, latency, and rejection logging all work. Engineering validation is necessary but insufficient for scientific retrieval claims. It does not demonstrate that the system retrieves datasets scientists would actually trust, reuse, compare, or reanalyze for real research goals.

Retrieval science validation requires relevance-labeled queries, graded usefulness judgments, hard-negative avoidance, ablation across system components, calibration of predicted scores, and expert review. Table~\ref{tab:eng-vs-sci} summarizes the distinction.

\begin{table}[htbp]
\centering
\caption{Engineering metrics versus scientific retrieval metrics.}
\label{tab:eng-vs-sci}
\small
\begin{tabular}{@{}p{0.22\linewidth}p{0.20\linewidth}p{0.50\linewidth}@{}}
\toprule
\textbf{Metric} & \textbf{Type} & \textbf{Interpretation} \\
\midrule
Corpus size & Engineering & Validates ingestion and deduplication. Not evidence of retrieval relevance. \\
Vector Recall@50 & Engineering & Validates approximate nearest-neighbor behavior vs.\ exact NN retrieval. Not end-to-end relevance. \\
Latency (p50/p95) & Engineering & Validates usability under current hardware. Not relevant quality. \\
Graph rank perturbation & Behavioral & Proves graph affects rankings. Does not prove improvement in scientific relevance. \\
Rejection rate & Engineering & Validates quality filtering. Does not validate retained-record scientific quality. \\
\midrule
NDCG@10, NDCG@20 & Scientific retrieval & Measures whether highly useful datasets rank near the top relative to adjudicated qrels. \\
MRR & Scientific retrieval & Measures reciprocal rank of first scientifically useful result. \\
Recall@50 (vs.\ qrels) & Scientific retrieval & Fraction of labeled relevant datasets retrieved in top 50. Distinct from vector-index recall. \\
Hard-negative violation rate & Scientific retrieval & Fraction of plausible-but-unusable datasets incorrectly promoted above relevant ones. \\
Inter-annotator agreement & Scientific retrieval & Validates label reliability before interpreting ranking metrics. \\
Expected Calibration Error & Scientific retrieval & Measures whether predicted usefulness scores reflect empirical relevance probability. \\
Prospective case studies & Scientific utility & Whether researchers can discover, choose, and reuse surfaced datasets for actual goals. \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Scientific Benchmark Design}

The benchmark query suite targets realistic neuroscience reuse scenarios, not simple keyword lookup. Each query specifies a scientific intent class, the evidence fields required for reuse feasibility, and known failure modes (hard negatives that score high on surface features but are scientifically unusable).

Intent classes used in the benchmark:
\begin{enumerate}
    \item \texttt{STRICT\_LOOKUP}: direct dataset accession or DOI lookup.
    \item \texttt{REPLICATION}: find datasets suitable for reproducing a known experimental context.
    \item \texttt{META\_ANALYSIS}: find datasets that can be aggregated across studies.
    \item \texttt{PIPELINE\_REUSE}: find datasets compatible with an existing analysis pipeline.
    \item \texttt{CROSS\_DATASET\_COMPARISON}: find datasets that support systematic comparison across species, modalities, or labs.
    \item \texttt{EXPLORATION}: discover scientifically adjacent datasets.
    \item \texttt{METHOD\_TRANSFER}: find datasets where a method can transfer from one experimental context to another.
    \item \texttt{MODEL\_VALIDATION}: find datasets for validating computational or mechanistic models.
    \item \texttt{REANALYSIS\_FEASIBILITY}: find datasets with sufficient behavioral and neural metadata to support model fitting or re-analysis.
\end{enumerate}

Example hard negatives used in the benchmark:
\begin{itemize}
    \item Mouse electrophysiology visual cortex when calcium imaging is required (modality mismatch).
    \item Human resting-state fMRI when a reward learning task fMRI is requested (task mismatch).
    \item EEG dataset without extracellular recordings when spike sorting is requested (data-type mismatch).
    \item Decision-making dataset with only summary statistics when trial-level behavioral events are required for model fitting.
\end{itemize}

\subsection{Graded Relevance Labels}

Relevance labels use a 0--3 graded scale:
\begin{itemize}
    \item \textbf{0 -- Not useful:} Topic, modality, or species mismatch; missing required data; or dataset is unusable for the stated scientific goal.
    \item \textbf{1 -- Weakly useful:} Superficially related but missing key metadata, weak task match, unclear reuse feasibility, or useful only as background context.
    \item \textbf{2 -- Useful:} Matches the main scientific goal and has enough metadata or data access for plausible reuse or comparison.
    \item \textbf{3 -- Highly useful:} Strong match across scientific goal, modality, species, task, metadata completeness, reuse feasibility, and provenance.
\end{itemize}

No scientific retrieval claim should be made until adjudicated qrels from at least two annotators per query-dataset pair are available, inter-annotator agreement is computed (Cohen's kappa or Krippendorff's alpha), and disagreements are resolved.

\subsection{Frozen Benchmark Protocol}
\label{sec:validation-protocol}

The validation protocol is designed to turn Neural Search from an architecture into a reproducible scientific retrieval benchmark. No end-to-end ranking claim should be made without a frozen corpus snapshot, relevance labels, run files, and metric reports.

\subsubsection{Frozen Corpus Snapshot}

Each benchmark run must use a frozen corpus snapshot. The corpus manifest should include:

\begin{itemize}
    \item corpus size and source distribution;
    \item canonical record identifiers;
    \item ingestion commit or source-harvest revision;
    \item embedding model, embedding dimension, tokenizer, and model configuration;
    \item graph build hash and graph input artifact hashes;
    \item rejection log hash and top rejection reasons.
\end{itemize}

\subsubsection{Benchmark Query Suite}

The target benchmark contains 100--150 queries sampled across the intent classes above. Each query specifies \texttt{query\_text}, \texttt{intent}, \texttt{scientific\_goal}, \texttt{must\_have} constraints, \texttt{hard\_negatives}, and expected modalities, species, tasks, regions, and affordances.

\subsubsection{Pooling and Annotation}

Candidate pools should be built from top-$k$ results returned by BM25, dense BGE retrieval, BM25+BGE RRF, the hybrid scorer, graph scorer, and usefulness scorer. Pooling avoids evaluating only one system's candidate distribution and follows the reproducible run/qrels discipline used in modern retrieval benchmarks such as BEIR \citep{thakur2021beir}.

Each pooled query--record pair should be judged with the 0--3 scale above by at least two annotators. Disagreements should be adjudicated by a third pass or consensus meeting. The benchmark report should include inter-annotator agreement before adjudication and label-distribution statistics after adjudication.

\subsubsection{Metrics}

The primary metrics are NDCG@10, NDCG@20 \citep{jarvelin2002cumulated}, MRR, Recall@50, hard-negative violation rate, source skew, and calibration expected calibration error (ECE) \citep{guo2017calibration}. Hard-negative constraints must be applied before ranking; a record that violates explicit exclusions is a retrieval failure even if it is otherwise topically similar.

\subsubsection{Ablation Plan}

The frozen benchmark should evaluate:

\begin{itemize}
    \item \texttt{no\_dense}
    \item \texttt{no\_sparse}
    \item \texttt{no\_graph}
    \item \texttt{no\_provenance}
    \item \texttt{no\_affordance}
    \item \texttt{no\_field\_weighting}
    \item \texttt{no\_missingness\_penalty}
    \item \texttt{no\_usefulness\_reranker}
\end{itemize}

\subsubsection{Statistical Testing}

Report bootstrap confidence intervals for all aggregate metrics. Compare retrieval variants with paired randomization or permutation tests over query-level metric differences. A metric delta should not be treated as meaningful unless the query-level paired test supports it and the failure analysis does not reveal systematic source or annotation artifacts.

\begin{table}[htbp]
\centering
\caption{Planned graph ablation table for the frozen expanded-corpus benchmark.}
\small
\begin{tabular}{@{}lcccc@{}}
\toprule
Variant & NDCG@10 & MRR & HN-VIOL & Source skew \\
\midrule
Hybrid without graph & Pending & Pending & Pending & Pending \\
Hybrid with graph & Pending & Pending & Pending & Pending \\
Graph-only candidate boost & Pending & Pending & Pending & Pending \\
Metapath-specific graph scoring & Pending & Pending & Pending & Pending \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Multi-Dataset Set-Coverage Scoring}
\label{sec:setcoverage}

Neural Search includes \emph{set-coverage scoring}: rather than ranking individual datasets only, the system can score an entire result set $\mathcal{D}$ for collective usefulness:

\begin{equation}
    \text{score}(\mathcal{D}) = \bar{u} + \alpha \cdot c_{\text{cov}} + \beta \cdot c_{\text{comp}} + \gamma \cdot c_{\text{prov}} - \delta \cdot p_{\text{red}} - \varepsilon \cdot p_{\text{miss}} - \zeta \cdot p_{\text{hard}}
\end{equation}

where $\bar{u}$ is mean individual usefulness, $c_{\text{cov}}$ rewards modality/species diversity, $c_{\text{comp}}$ rewards complementary affordances across the set, $c_{\text{prov}}$ rewards provenance completeness, and $p_{\text{red}}, p_{\text{miss}}, p_{\text{hard}}$ penalize redundancy, missing metadata, and hard-negative constraint violations respectively. Default weights: $\alpha = \beta = \gamma = 0.1$, $\delta = 0.15$, $\varepsilon = 0.05$, $\zeta = 0.30$.

Each result dataset is also assigned exactly one role from an 8-role taxonomy:

\begin{enumerate}
    \item \textbf{ANCHOR} --- highest usefulness, matches $\geq 3$ sub-queries
    \item \textbf{REPLICATION} --- same task + species as anchor, different modality
    \item \textbf{CROSS-SPECIES COMPARATOR} --- same task, different species from anchor
    \item \textbf{METHODOLOGICAL COMPLEMENT} --- different region, shares $\geq 2$ affordances
    \item \textbf{PERTURBATION/CAUSAL} --- optogenetic/pharmacological manipulation present
    \item \textbf{BEHAVIOR-RICH} --- rich trial-by-trial events, minimal neural recording
    \item \textbf{POPULATION DYNAMICS} --- large cell count ($\geq 100$), dimensionality reduction
    \item \textbf{IMAGING-EPHYS BRIDGE} --- combines fMRI/EEG with electrophysiology
\end{enumerate}

A dataset that satisfies none of these roles is assigned \textsc{UNASSIGNABLE} and excluded from the final result set. This role taxonomy enforces that each returned dataset contributes a distinct scientific function.

% ============================================================================
% SECTION 9: SYSTEM ARCHITECTURE
% ============================================================================
\section{System Architecture}
\label{sec:architecture}

\subsection{Software Engineering Principles}

Neural Search implements principled software engineering:

\begin{enumerate}
    \item \textbf{Typed Contracts}: Pydantic schemas enforce API payloads, records, and responses.

    \item \textbf{Separation of Concerns}: Distinct modules for ingestion, normalization, ontology, retrieval, graph, embeddings, evaluation, and reporting.

    \item \textbf{Configuration as Policy}: Retrieval weights, intent profiles, and gates live in YAML/JSON artifacts.

    \item \textbf{Deterministic CI}: Hashing embeddings and local fixtures ensure reproducibility.

    \item \textbf{Explainable Design}: Every result exposes score breakdowns, evidence, and warnings.

    \item \textbf{Progressive Enhancement}: In-memory demo corpus scales to persistent stores and vector indexes.

    \item \textbf{Artifact-Oriented Development}: Reports, cards, notebooks, and benchmarks are first-class outputs.
\end{enumerate}

\subsection{Module Architecture}

\begin{figure}[htbp]
\centering
\begin{tikzpicture}[
    node distance=1.5cm,
    box/.style={rectangle, draw, rounded corners, minimum width=2.5cm, minimum height=0.8cm, align=center, font=\small},
    arrow/.style={->, thick}
]
    % Data layer
    \node[box, fill=blue!20] (corpus) {Corpus\\Ingestion};
    \node[box, fill=blue!20, right=of corpus] (normalize) {Normalization};
    \node[box, fill=blue!20, right=of normalize] (extract) {Label\\Extraction};

    % Knowledge layer
    \node[box, fill=green!20, below=of corpus] (ontology) {Ontology\\Matching};
    \node[box, fill=green!20, right=of ontology] (graph) {Knowledge\\Graph};
    \node[box, fill=green!20, right=of graph] (embed) {Embeddings};

    % Search layer
    \node[box, fill=orange!20, below=of ontology] (parse) {Query\\Parser};
    \node[box, fill=orange!20, right=of parse] (score) {Hybrid\\Scorer};
    \node[box, fill=orange!20, right=of score] (rank) {Ranking \&\\Explanation};

    % Arrows
    \draw[arrow] (corpus) -- (normalize);
    \draw[arrow] (normalize) -- (extract);
    \draw[arrow] (extract) -- (ontology);
    \draw[arrow] (extract) -- (graph);
    \draw[arrow] (extract) -- (embed);
    \draw[arrow] (ontology) -- (score);
    \draw[arrow] (graph) -- (score);
    \draw[arrow] (embed) -- (score);
    \draw[arrow] (parse) -- (score);
    \draw[arrow] (score) -- (rank);
\end{tikzpicture}
\caption{Neural Search Module Architecture}
\end{figure}

\subsection{Integrated Corpus, Silver Labeling, and Retrieval Flow}
\label{sec:integrated-corpus-labeling-flow}

The current system is best understood as a layered artifact pipeline rather than a single monolithic search index. Each layer adds evidence while preserving provenance boundaries:

\begin{enumerate}
    \item \textbf{Corpus expansion and normalization.} Source adapters harvest or refresh repository records, map them into the normalized dataset schema, and write source identifiers, titles, URLs, descriptions, licenses, DOIs, species, modalities, regions, tasks, behaviors, standards, and provenance fields. The current live artifact contains 7,171 normalized rows. High-volume repositories expand recall; curated neuroscience repositories and NWB/DANDI enrichment passes increase evidence depth.

    \item \textbf{Evidence-tiered enrichment.} Structured repository fields, curated source categories, NWB electrode tables, NWB surgery/experiment text, and CrossRef abstracts are treated as different evidence tiers. Structured metadata is stronger evidence than rule-based free-text extraction. Paper-derived and NWB-text-derived region labels are explicitly silver labels, not human gold.

    \item \textbf{Embedding and vector indexing.} Normalized records are converted into BGE-large field embeddings and a turbovec-compatible vector index for approximate nearest-neighbor retrieval. The current indexed slice contains 625 records and should be regenerated against the frozen expanded corpus before final benchmark reporting.

    \item \textbf{Field-State Memory Graph.} The memory graph stores dataset nodes, concept nodes, raw-data-signal nodes, neuro-judge evidence packets, feedback signals, papers, modalities, species, tasks, regions, and analysis affordances. It is the integration layer that lets corpus records, silver labels, human-review state, and retrieval diagnostics share stable identifiers without collapsing provenance.

    \item \textbf{RAG-grounded neuro-judge triage.} Candidate query--dataset pairs are converted into evidence packets containing query intent, hard negatives, dataset metadata, concept-memory evidence, linked-paper evidence, raw-data signals, missing dimensions, and warnings. LLM or mock judges emit 0--3 silver labels, confidence, matched dimensions, evidence for/against, abstain recommendations, and failure modes. These labels route human review and diagnostics; they never write \texttt{human\_gold} provenance.

    \item \textbf{Literature-scale ingestion and finding extraction (added 17~June~2026).} 255,940 OpenAlex tier~1 papers (neuroscience concept \texttt{C169760540}, $\geq$100 citations) were ingested via a cursor-based bulk downloader (\texttt{openalex\_bulk.py}) with checkpoint/resume and written to 26 JSONL shards. The paper-dataset linker was run over all 7,171 corpus datasets, producing \texttt{paper\_dataset\_links.jsonl}; current link-method counts are 168 DOI-exact links, 225 title-fuzzy-local links, and 6,778 not-found records. This is a corpus-wide linker run, not evidence that every dataset was successfully linked; precision and false-negative audits are pending. LLM-powered structured finding extraction (\texttt{find\-ing\_ex\-trac\-tor.py}) runs on local inference: \texttt{extract\_batch\_ollama()} posts to Ollama's \texttt{/api/chat} endpoint and calls \texttt{\_repair\_json()} to strip markdown fences from local model output before JSON parsing. Each successfully parsed abstract yields zero or more \texttt{FindingRecord} objects with \texttt{finding\_text}, \texttt{result\_direction}, \texttt{regions}, \texttt{species}, \texttt{modalities}, \texttt{tasks}, \texttt{cell\_types}, \texttt{molecules}, and \texttt{confidence}. A rule-based typed-field pass enriches findings with 27 additional fields spanning polarity, experimental condition, signal type, statistics, behavior, anatomy, cell population, plasticity, network coupling, decoding, computational models, stimulus class, pharmacology, disease model, development, genetic tools, molecular markers, social-affective paradigms, sleep state, metabolic context, connectivity, and dimensionality reduction. The batch-level checkpoint ensures that papers returning empty findings are still marked processed and are not reprocessed on resume. The current operational extraction snapshot is 16,150 papers processed, 12,050 findings, and 74.6\% yield, projecting roughly 190K tier~1 findings if that rate holds; a later typed-field coverage audit over 122K findings shows stable coverage across all 27 fields. These are live operational estimates and coverage diagnostics, not validated quality results. Completed findings are incorporated into the knowledge graph via a literature KG builder (\texttt{kg\_builder.py}) that adds \texttt{paper}, \texttt{finding}, and \texttt{venue} nodes with bidirectional edges to existing dataset and concept nodes.

    \item \textbf{Human qrels and feedback.} Expert qrels are imported into separate adjudicated artifacts. Frontend feedback and downstream retrieval success signals are stored with downstream-feedback provenance and can influence audit queues or transparent reranking priors, but they do not overwrite qrels or promote LLM outputs.

    \item \textbf{Benchmark and paper reporting.} Evaluation scripts consume frozen corpus, embedding, index, graph, qrels, feedback, and run artifacts to generate metric reports and manuscript tables. Missing or stale artifacts are surfaced as pending validation instead of being silently filled by historical numbers.
\end{enumerate}

This integration design makes recent expansion useful before final benchmark validation: new records become searchable and reviewable immediately, but every downstream claim remains traceable to its evidence tier. A DANDI brain-region label from an NWB electrode table, a CrossRef-derived anatomical mention, a neuro-judge relevance label, a user feedback event, and an adjudicated human qrel can all refer to the same dataset, but the system stores them as different evidence types with different authority.

\section{Graph-Indexed Concept Memory}
\label{sec:concept-memory}

\subsection{Motivation and Design}

Dense retrieval surfaces datasets by embedding similarity but cannot explain \emph{why} a dataset is relevant to a specific scientific goal. Sparse retrieval is interpretable token by token but misses synonyms and concept-level structure. The Graph-Indexed Concept Memory (GICM) layer is a complementary component that indexes the field's conceptual knowledge---tasks, modalities, species, brain regions, methods, claims, and failure modes---as a typed, evidence-weighted directed graph, and uses that graph to rerank and explain retrieval results.

GICM has two roles:
\begin{enumerate}
    \item \textbf{Semantic enrichment}: Each indexed dataset is connected to the concept nodes that describe it (task, modality, species, brain region, method). Retrieval can use these links to boost datasets that match the query's conceptual signature, not just its surface tokens.
    \item \textbf{Result explainability}: For each retrieved dataset, GICM produces a structured explanation: which concept types matched, what evidence supports the match, what is missing, and whether any failure-mode conflicts were detected.
\end{enumerate}

\subsection{Concept Graph Construction}

The concept graph is built from four sources: field-state artifacts (claims, benchmark gaps, opportunities), the indexed corpus, Obsidian vault notes (if available), and paper notes. Each source loader extracts \texttt{ConceptNode} records and \texttt{EvidenceLink} records. Alias normalization is applied at load time to merge synonymous terms (e.g., ``ephys'' $\to$ ``electrophysiology''). The resulting graph currently contains:

\begin{itemize}
    \item 13,281 concept nodes across 17 typed categories
    \item 29,684 evidence links
    \item A dataset-concept layer that should be regenerated after corpus/index reconciliation so concept counts align with the frozen benchmark snapshot
\end{itemize}

Concept nodes are assigned a stable deterministic ID (\texttt{concept:<type>:<slug>}), a canonical name, aliases, description, source provenance, evidence count, review status, and a confidence float in $[0, 1]$.

\subsection{Retrieval Integration and Score Decomposition}

The GICM reranking layer is \emph{disabled by default} and must be explicitly enabled. When active, every final score decomposes into four named components:

\begin{equation}
    \hat{s}(q, d) = s_\text{base}(q, d) + s_\text{concept}(q, d) + s_\text{evidence}(q, d) - p_\text{hn}(d)
\end{equation}

where:
\begin{itemize}
    \item $s_\text{base}(q, d)$ is the lexical token-overlap score between the query and the dataset's canonical name, aliases, description, and tags.
    \item $s_\text{concept}(q, d) = \min\!\left(\sum_{c \in \mathcal{N}(d)} \text{match}(q, c) \cdot w_\text{type}(c) \cdot \text{conf}(c, d),\; 1\right) \cdot \lambda_c$ where $\mathcal{N}(d)$ are concept nodes connected to dataset $d$, $w_\text{type}$ is a type weight (task: 0.40, modality: 0.30, species/brain region: 0.15, method: 0.10), and $\lambda_c = 0.30$ is a global concept-boost scale.
    \item $s_\text{evidence}(q, d)$ is a small bonus for evidence links that are reviewed or carry strong provenance, scaled by $\lambda_e = 0.10$.
    \item $p_\text{hn}(d)$ is a hard-negative penalty (up to 0.10) triggered when a dataset is linked to \texttt{failure\_mode} concept nodes.
\end{itemize}

All components are stored and can be serialized to JSON, enabling result-level auditing and future dashboard rendering.

\subsection{Ablation Evaluation Plan}

Four retrieval variants are defined for ablation comparison:

\begin{enumerate}
    \item \textbf{lexical\_only}: $s_\text{base}$ only
    \item \textbf{concept\_boost}: $s_\text{base} + s_\text{concept}$
    \item \textbf{concept\_boost\_ev}: $s_\text{base} + s_\text{concept} + s_\text{evidence}$
    \item \textbf{full}: all components
\end{enumerate}

Metrics include NDCG@10 (graded relevance, 0--3 scale), MRR (binary threshold $\geq 1$), Recall@10, Recall@50, hard-negative violation rate, and source-skew concentration. Evaluation is run with \texttt{python -m neural\_search.field\_state.cli concept-eval}.

\subsection{Validation Status}

\textbf{The graph boost is now measurable, but still preliminary.} The graph influence on final scores is scientifically meaningful only if it produces measurable improvement in qrels-backed retrieval metrics (NDCG@10, MRR) over strong non-graph baselines. As of the current system state, the populated qrels-backed retrieval snapshot covers 317 canonical queries and 13,654 non-error LLM-judged query--dataset pairs. This is sufficient for a useful ablation preview, but it is not a substitute for independently adjudicated human labels. After calibrating the global graph multiplier to 0.005, \texttt{hybrid\_graph} shows a small LLM-qrels-backed gain over \texttt{hybrid\_rrf}; this should be treated as a promising calibration result, not as publication-grade validation. Until human/adjudicated graph ablations are available:

\begin{itemize}
    \item All score decompositions are structurally correct and unit-tested.
    \item No publication-grade retrieval quality claim based on concept boost should be made from LLM-only labels.
    \item The GICM layer should be described as a \emph{calibrated enrichment mechanism with an active evaluation harness}, not yet a human-validated retrieval improvement.
\end{itemize}

\section{Audit Protocols}
\label{sec:audit-protocols}

The benchmark protocol above governs retrieval quality. The following audits govern the scientific evidence used by retrieval, graph construction, and manuscript claims.

\begin{table}[htbp]
\centering
\caption{Formal audit protocols required before promoting silver artifacts to publication claims.}
\scriptsize
\begin{tabular}{p{0.18\linewidth}p{0.18\linewidth}p{0.28\linewidth}p{0.18\linewidth}p{0.10\linewidth}}
\toprule
Audit & Sampling unit & Procedure & Reported metrics & Required artifact \\
\midrule
Finding extraction audit & Random and stratified sample of extracted findings, including empty-extraction papers & Compare each \texttt{FindingRecord} against the source abstract. Judge whether the finding text is supported, whether direction is correct, and whether region/species/modality/task fields are warranted. Include strata by model confidence, venue year, and abstract length. & Finding precision; unsupported-finding rate; field-level precision/recall; empty-array correctness & \texttt{reports/audits/finding\_extraction\_audit.csv} plus adjudication notes \\
Paper-dataset link audit & All DOI-exact links plus stratified title-fuzzy-local links and not-found records & Verify DOI identity, title match, year/author plausibility, and dataset provenance. Estimate false positives for successful links and false negatives among not-found records. & DOI precision; fuzzy-link precision; not-found false-negative rate; coverage by source & \texttt{reports/audits/paper\_dataset\_link\_audit.csv} \\
Affordance precision audit & Dataset-affordance pairs across all 21 affordance types and support levels & Inspect structured metadata and, when available, representative NWB/BIDS files or repository documentation. Mark each support label as supported, unsupported, or insufficient evidence. & Precision by affordance type; precision by support level; unsupported/inconclusive rates & \texttt{reports/audits/affordance\_precision\_audit.csv} \\
Qrels adjudication protocol & Pooled query-dataset candidates from all retrieval variants & Two independent annotators assign 0--3 relevance, hard-negative status, missing-metadata flags, confidence, and rationale. Disagreements with $|s_A-s_B|\geq 2$ or conflicting hard-negative flags receive adjudication. & Inter-annotator agreement; label distribution; adjudication rate; NDCG/MRR/Recall on human-adjudicated qrels & \texttt{artifacts/field\_state/adjudicated\_qrels.jsonl} and metric reports \\
\bottomrule
\end{tabular}
\end{table}

\noindent\textbf{LLM qrels status.} LLM-generated qrels, neuro-judge packets, and consensus labels are silver diagnostic artifacts. They can prioritize review, reveal prompt collapse, and support exploratory ablations, but they cannot produce publication-grade NDCG, MRR, or Recall. Publication-grade retrieval metrics require human-adjudicated qrels, agreement reporting, and reproducible run files over a frozen corpus snapshot.

A hard-negative violation is recorded when a result explicitly matches a stated \texttt{hard\_negatives} description for the query. These are tracked separately as they indicate systemic failures (e.g., returning mouse calcium imaging when the query explicitly requires human fMRI).

\subsection{Evaluation Metrics}

For each retrieval variant, the following metrics are computed:

\begin{itemize}
    \item $\text{NDCG@10}$ and $\text{NDCG@20}$: Normalized Discounted Cumulative Gain using the graded 0--3 relevance scale
    \item $\text{MRR}$: Mean Reciprocal Rank (binary threshold: relevance $\geq 2$)
    \item $\text{P@10}$, $\text{R@10}$, $\text{R@50}$: Precision and Recall at rank cutoffs
    \item $\text{HN rate}$: Fraction of top-10 results that are hard-negative violations
    \item $\text{Source skew}$: Entropy-normalized concentration of top-10 results by data repository
    \item Bootstrap 95\% CI (1000 samples) for NDCG@10, MRR, Recall@50
\end{itemize}

\subsection{Retrieval Variants}

The canonical 317-query ablation ladder compares six rungs, in increasing order of signal:

\begin{enumerate}
    \item \texttt{bm25}: BM25 sparse retrieval only
    \item \texttt{bm25\_structured}: BM25 + structured ontology-dimension matching
    \item \texttt{dense\_bge}: BGE-large-en-v1.5 dense field embeddings only
    \item \texttt{hybrid\_rrf}: reciprocal rank fusion of sparse and dense signals
    \item \texttt{hybrid\_graph}: \texttt{hybrid\_rrf} + calibrated knowledge-graph boost
    \item \texttt{full}: all components, including coverage-gap and source-diversity post-processing
\end{enumerate}

An earlier, smaller-scale concept-memory ablation (\texttt{bm25} / \texttt{usefulness} / \texttt{concept\_boost} / \texttt{concept\_boost\_ev} / \texttt{concept\_full}) remains available in \texttt{reports/ablation\_v08.md} but ran on synthetic seed pairs and produced no rung differentiation; it is superseded by the canonical ladder above for any retrieval-quality claim.

\subsection{Candidate Pooling and Annotation}

Annotation candidates are built by pooling the top-50 results from all active variants. Consensus candidates (retrieved by multiple variants) are prioritized for annotation. The concept-rerank strategy is included in the pool when field-state artifacts are present, ensuring GICM-specific results are represented in the annotation set.

A minimum of 30\% of pairs are independently annotated by two annotators. Disagreements where $|s_A - s_B| \geq 2$ are adjudicated. The \texttt{qrels.jsonl} schema records: \texttt{relevance}, \texttt{rationale}, \texttt{hard\_negative\_violation}, \texttt{missing\_metadata}, \texttt{annotator\_id}, \texttt{adjudicated}, and \texttt{adjudication\_notes}.

\subsection{Minimum Credible Benchmark}

Results from this benchmark support paper-ready claims only when:

\begin{itemize}
    \item $\geq 100$ queries with adjudicated qrels
    \item $\geq 1,500$ judged query-dataset pairs
    \item All 7 intent types represented
    \item $\geq 5$ distinct data repositories in the corpus
    \item All annotator disagreements adjudicated
    \item Bootstrap 95\% CI does not include 0 improvement for primary claims
\end{itemize}

\subsection{Current Status}

\textbf{A first qrels-backed retrieval snapshot is now available, but remains preliminary.} The current canonical qrels artifact contains 13,654 non-error LLM-judged query--dataset pairs across 317 canonical queries. The join is complete across all six retrieval rungs. On this snapshot, \texttt{hybrid\_rrf} remains a strong non-graph baseline with NDCG@10 = 0.6667, MRR = 0.9209, and Recall@50 = 0.7455. BM25 remains competitive on NDCG@10 (0.6566), while dense BGE alone has lower aggregate NDCG@10 (0.5708) and Recall@50 (0.5384). The graph-backed variants are now populated, calibrated, and evaluable: \texttt{hybrid\_graph} obtains NDCG@10 = 0.6696, MRR = 0.9256, and Recall@50 = 0.7465; \texttt{full} obtains NDCG@10 = 0.6696, MRR = 0.9256, and Recall@50 = 0.7450. The calibrated graph multiplier is 0.005, selected by \texttt{graph\_weight\_calibration.md} as the best balanced setting that improves NDCG@10 and MRR over \texttt{hybrid\_rrf}. Bootstrap/sign-test analysis finds \texttt{hybrid\_graph} significantly better than \texttt{hybrid\_rrf} on NDCG@10 ($p=0.0052$) and MRR ($p=0.0020$), significantly better than BM25 on MRR, and significantly better than dense BGE on NDCG@10 and MRR. The graph gain is small and LLM-qrels-backed, so it should be framed as a promising calibrated preview rather than final human-adjudicated evidence.

These numbers should be reported as an LLM-judged benchmark preview, not as final human-adjudicated scientific relevance evidence. Dual-judge reliability is not estimable from the current artifact: 13,654 labels are usable, but no non-error query--dataset pair was judged by two models.

The evaluation infrastructure is complete:
\begin{itemize}
    \item Benchmark specification is frozen (\texttt{docs/benchmark\_v1\_spec.md})
    \item 317 canonical evaluated queries are available in \texttt{data/eval/benchmark\_queries\_canonical.yaml}
    \item Canonical LLM qrels are available in \texttt{data/qrels/qrels.canonical.jsonl} and \texttt{data/qrels/qrels.canonical.trec}
    \item Candidate pool scripts operational (\texttt{sample\_candidate\_pool.py})
    \item Metric computation with bootstrap CI operational (\texttt{compute\_ndcg\_from\_qrels.py}, \texttt{compute\_bootstrap\_ci.py})
    \item Intent stratification, dual-judge reliability reporting, claim ledger generation, and regression gate scripts operational
    \item Failure analysis and result card gallery scripts operational
    \item Graph/full ablation rungs populated from a 2,821-record eval-corpus graph with 3,180 nodes and 17,523 edges, including 5,254 cross-dataset and reanalysis relationship edges
    \item Graph-weight calibration and relationship-edge quality reporting operational; current relationship-edge promotions are mixed, so edge confidence/source calibration remains a priority
    \item Reinterpretation and reprocessing discovery reports operational: reanalysis affordance ranking, new-method-to-old-dataset matching, and metadata-enrichment priorities derived from failure analysis
    \item Annotation tool with resume mode operational
    \item Frontend feedback loop operational: React search interface with evidence inspection, inline relevance feedback (usefulness, would-use-for-analysis, reason tags, free-text notes), neuro-judge badges with silver provenance watermarks, and save/export
    \item Downstream retrieval success reporting operational (\texttt{compute\_downstream\_retrieval\_success.py})
    \item Dataset organisation views operational (\texttt{build\_dataset\_organization\_views.py}) — groups judged corpus into relevance tiers, modality buckets, and special views (missing raw data, hard negatives, abstain-flagged, metadata gaps)
    \item Feedback audit queue operational (\texttt{build\_feedback\_audit\_queue.py}) — priority scoring for judge/user disagreements, high-rank-not-useful cases, abstain-but-useful patterns
    \item Transparent feedback reranking priors operational (\texttt{build\_feedback\_rerank\_priors.py}) — log-odds weight table ($\pm$1.5 cap) from aggregated downstream feedback, toggle-able in frontend; downstream signal only
\end{itemize}

No publication-grade improvement claim based on concept-memory or graph reranking appears in this paper. The current calibrated graph result is a positive LLM-qrels preview; any final claim requires qrels-backed evidence meeting the minimum credible benchmark targets above and independently adjudicated labels.

\section{Limitations}
\label{sec:limitations}

\paragraph{Vector recall is not scientific usefulness.} The current turbovec recall@50 result validates approximate vector-index retrieval against exact nearest-neighbor retrieval under the current embedding configuration. It does not show that returned datasets are scientifically relevant, reusable, compatible with a method, or better than sparse and hybrid baselines under human relevance judgments.

\paragraph{Engineering tests are not human labels.} The Field-State, Obsidian memory, and evaluation-memory tests validate schemas, JSONL persistence, report generation, roundtrips, human-block preservation, qrels import/export, adjudication logic, and snapshot generation. They do not substitute for expert relevance judgments or qrels-backed NDCG/MRR evaluation.

\paragraph{Current qrels coverage is LLM-judged, not human-adjudicated.} The current canonical qrels snapshot covers 317 queries and 13,654 non-error LLM judgments, which is enough for a meaningful ablation preview and regression gate. It is not yet enough for publication-grade human relevance claims because no dual-judge overlap is available and expert adjudication remains pending.

\paragraph{Future usefulness remains weakly validated.} Claims about future reuse potential, dataset-method compatibility, and downstream scientific value require longitudinal reuse outcomes or defensible proxy labels. Until those labels exist, usefulness scores should be interpreted as prioritization signals rather than validated predictions.

\paragraph{Metadata incompleteness.} The current corpus is large, but structured metadata remains uneven across sources. Task and brain-region coverage are materially weaker than title, URL, description, and license coverage.

\paragraph{Source skew.} High-volume general repositories such as Zenodo and Figshare can dominate candidate pools if source diversity and source reliability are not monitored. Benchmark reports must include source-skew diagnostics computed over qrels-backed run files.

\paragraph{Extraction noise.} Many labels are extracted from titles, descriptions, keywords, and repository metadata rather than independently verified experimental files. Source-specific manual audits are required for modality, species, and task precision.

\paragraph{Unvalidated affordance inference.} Analysis affordances are currently represented and predicted, but content-validated affordance precision/recall is pending. Metadata-only affordance claims should be surfaced with uncertainty.

\paragraph{Graph construction bias.} Calibrated graph proximity now shows a small positive LLM-qrels-backed retrieval gain, but graph edges can also amplify incomplete or noisy metadata. The relationship-edge quality report shows mixed promotion quality, so graph-weight and edge-confidence calibration must be revalidated on adjudicated labels.

\paragraph{Dense embedding domain mismatch.} BGE-large-en-v1.5 provides a strong general dense retrieval representation, and broad embedding benchmarks such as MTEB are useful external references \citep{muennighoff2022mteb}. Scientific dataset usefulness may still require domain-specific calibration and field weighting.

\paragraph{Benchmark label subjectivity.} Scientific usefulness is intent-dependent. Multi-annotator labeling, adjudication, and inter-annotator agreement are required before final performance claims.

\paragraph{Obsidian traceability is not validation.} The Obsidian-compatible review layer improves human auditability by separating generated fields from protected human notes. It is a traceability mechanism, not an epistemic guarantee. Likewise, weak supervision, generated review prompts, or future LLM triage may help prioritize review work but should not be treated as gold labels.

\paragraph{Finding extraction quality is preliminary.} LLM-extracted findings from \texttt{qwen2.5:7b-instruct-q4\_K\_M} have been manually spot-checked on a small smoke set but not formally precision/recall evaluated. Region and species fields rely on the model correctly identifying ontological terms from free text; output quality will vary across abstract styles. A formal precision audit is required before publication claims about finding accuracy. The current 16,150-paper operational snapshot produced 12,050 findings at 74.6\% yield, projecting roughly 190K tier~1 findings if maintained; a later typed-field audit enriched 122K findings with 27 rule-based fields. These counts likely reflect high-citation tier composition and should not be treated as final yield or quality estimates.

\paragraph{Literature search not yet integrated into main search path.} The finding and paper search infrastructure operates as a separate index. Integration into the unified hybrid scorer, ranking pipeline, and frontend is staged for Horizon 3.

\paragraph{No content-level neural signatures yet.} Neural signature extraction is not yet implemented in the active usefulness scorer. The neural-signature dimension is excluded from active weight normalization until content-derived features are available.

\section{Reproducibility Artifacts}
\label{sec:reproducibility-artifacts}

The paper should be regenerated from versioned artifacts rather than hand-written numbers. The expected artifact bundle is:

\begin{itemize}
    \item \texttt{corpus\_manifest.json}
    \item \texttt{source\_distribution.csv}
    \item \texttt{field\_completeness.csv}
    \item \texttt{rejection\_summary.csv}
    \item \texttt{embedding\_manifest.json}
    \item \texttt{graph\_manifest.json}
    \item \texttt{benchmark\_queries.jsonl}
    \item \texttt{qrels.jsonl}
    \item retrieval run files for each variant
    \item \texttt{eval\_report.json}
    \item \texttt{ablation\_report.json}
    \item \texttt{calibration\_report.json}
    \item \texttt{artifacts/frontend/retrieval\_feedback.jsonl} (downstream user feedback — not gold)
    \item \texttt{artifacts/field\_state/dataset\_organization\_views.json}
    \item \texttt{artifacts/field\_state/feedback\_audit\_queue.jsonl}
    \item \texttt{artifacts/field\_state/feedback\_rerank\_priors.json}
\end{itemize}

The current repository includes scripts under \texttt{scripts/eval/} to freeze corpus metadata, build candidate pools, compute IR metrics, initialize ablation reports, audit extraction quality, generate paper tables, compute downstream retrieval success from user feedback, build dataset organisation views, build a feedback audit queue, and derive transparent feedback reranking priors. Missing artifacts should be reported as \texttt{Pending benchmark artifact}, not silently replaced with assumptions.

\subsection{Field-State and Evaluation-Memory Commands}

The lightweight validation-memory layer is regenerated from local artifacts with:

\begin{lstlisting}[language=sh]
python -m neural_search.field_state.cli init
python -m neural_search.field_state.cli report
python -m neural_search.field_state.cli opportunities
python -m neural_search.field_state.cli snapshot
python -m neural_search.field_state.cli export-obsidian --vault <vault>
python -m neural_search.field_state.cli qrels-export \
  --pool reports/eval/benchmark_pool.jsonl \
  --queries artifacts/benchmark_queries.jsonl \
  --corpus data/corpus/normalized/combined_corpus.jsonl \
  --vault <vault> \
  --field neuroscience_dataset_reuse
python -m neural_search.field_state.cli qrels-import --vault <vault>
python -m neural_search.field_state.cli qrels-adjudicate
python -m neural_search.field_state.cli qrels-agreement
python -m neural_search.field_state.cli eval-snapshot
python -m neural_search.field_state.cli claim-evidence-update
python -m neural_search.field_state.cli whitepaper-validation-report
\end{lstlisting}

\subsection{Field-State Memory Graph Commands (v0.9)}

The field-state memory graph is built, validated, and updated with:

\begin{lstlisting}[language=sh]
# Build graph from corpus + evidence + feedback
python scripts/field_state/build_memory_graph.py \
  --corpus data/corpus/normalized/combined_corpus.jsonl

# Validate graph invariants and guardrails
python scripts/field_state/validate_memory_graph.py

# Incremental update (detects changed records by hash)
python scripts/field_state/update_field_state.py \
  --corpus data/corpus/normalized/combined_corpus.jsonl

# Force full rebuild
python scripts/field_state/update_field_state.py \
  --corpus data/corpus/normalized/combined_corpus.jsonl --force

# Compare two snapshots
python scripts/field_state/compare_snapshots.py \
  --old artifacts/field_state/snapshots/<old_ts> \
  --new artifacts/field_state/snapshots/<new_ts>

# Update learning signals
python scripts/field_state/update_learning_signals.py

# Retrieval loading diagnostics
python scripts/eval/diagnose_retrieval_loading.py
\end{lstlisting}

\subsection{Knowledge Graph and Embedding Rebuild Commands (v1.0)}

\begin{lstlisting}[language=sh]
# Rebuild full-corpus knowledge graph (all 7,171 records)
python scripts/build_real_corpus_graph.py

# Rebuild field embeddings for all corpus records (requires GPU ~1h)
python scripts/embed_flat_corpus.py

# Build/load the aggregated NPY embedding cache used by ablation
python scripts/eval/cache_embeddings_npy.py

# Rebuild with explicit corpus path
python scripts/embed_flat_corpus.py \
  --corpus data/corpus/normalized/combined_corpus.jsonl/full_corpus_v09.jsonl

# Generate annotation pool from 100 benchmark queries
python scripts/eval/generate_annotation_pool.py \
  --queries data/eval/benchmark_queries_v2.yaml \
  --out reports/eval/annotation_pool.jsonl

# Annotate candidates (interactive CLI)
python scripts/eval/annotate_candidates.py \
  --pool reports/eval/annotation_pool.jsonl \
  --queries reports/eval/annotation_queries.jsonl \
  --out artifacts/qrels_human.jsonl

# Run neural search baseline on benchmark queries
python scripts/eval/run_neural_search_baseline.py \
  --top-k 20
\end{lstlisting}

These commands produce local JSONL and Markdown artifacts for reviewability. Their existence should be cited as workflow reproducibility, not as evidence of ranking quality unless paired with completed human labels and metric reports.

% ============================================================================
% SECTION 10: FUTURE DIRECTIONS
% ============================================================================
\section{Future Directions}
\label{sec:future}

\subsection{Prioritized Next-Stage Roadmap}
\label{sec:roadmap}

The roadmap below orders work by impact on the three core goals: corpus growth, neural coverage, and search accuracy/relevance. Items are grouped into four horizons.

\paragraph{Horizon 0 --- Literature layer (infrastructure implemented; extraction and validation in progress).}

\begin{enumerate}
    \item \textbf{OpenAlex bulk ingestion} (\textit{implemented}): 255,940 tier~1 papers (neuroscience, $\geq$100 citations) ingested via cursor-based pagination with checkpoint/resume. Staged tiers: tier~2 ($\sim$1.39M open-access), tier~3 ($\sim$4.36M total).

    \item \textbf{Paper-dataset linking} (\textit{linker run complete; audit pending}): The linker was run over all 7,171 corpus datasets. Current method counts are 168 DOI-exact links, 225 title-fuzzy-local links, and 6,778 not-found records. Artifact: \texttt{artifacts/literature/paper\_dataset\_links.jsonl}. Precision and false-negative audits are pending.

    \item \textbf{LLM finding extraction} (\textit{in progress}): \texttt{extract\_batch\_ollama()} using \texttt{qwen2.5:7b-instruct-q4\_K\_M}; \texttt{\_repair\_json()} for markdown fence handling; \texttt{FindingRecord} schema plus 27 typed extension fields; batch-level checkpoint with append-mode output. Current operational estimate: 16,150 papers processed, 12,050 findings, 74.6\% yield, and $\sim$190K projected tier~1 findings if maintained. A 122K-finding typed-field coverage audit places core structural fields above 5\%, rich annotation fields at 2--5\%, and sparse high-specificity fields below 2\%. This is not a validated quality estimate.

    \item \textbf{Literature KG rebuild} (\textit{infrastructure implemented; full rebuild pending}): \texttt{kg\_builder.py} adds \texttt{paper}, \texttt{finding}, and \texttt{venue} node types; new edge types \texttt{paper\_reports\_finding}, \texttt{finding\_involves\_\{region, task, modality, species\}}, \texttt{dataset\_linked\_to\_paper}. The full graph should be rebuilt after extraction and link audits are frozen.

    \item \textbf{Literature BM25 search} (\textit{implemented separately; unified search pending}): \texttt{neural\_search/literature/search.py} provides paper and finding retrieval. Integration into the unified dataset/paper/finding search path remains a separate milestone.

    \item \textbf{Formal audits} (\textit{pending}): Complete the finding extraction audit, paper-dataset link audit, affordance precision audit, and qrels adjudication protocol defined in Section~\ref{sec:audit-protocols}.
\end{enumerate}

\paragraph{Horizon 1 — Immediate (affordance/ontology expansion + LLM annotation, June 2026, in progress).}

\begin{enumerate}
    \setcounter{enumi}{6}
    \item \textbf{Expand affordance library to 21 types} (\textit{complete as of 15~June~2026}).
    Added six new analysis types: speech decoding, seizure detection, sleep stage classification, BCI decoding, latent dynamics modeling, and representational similarity analysis. The \texttt{DatasetFeatures} struct now detects four new signal fields (\texttt{has\_speech\_events}, \texttt{has\_seizure\_annotations}, \texttt{has\_sleep\_stage\_labels}, \texttt{has\_bci\_context}). See Table~\ref{tab:affordances} for the full 21-type registry.

    \item \textbf{Expand brain region ontology from 106 to 130+ entries} (\textit{complete as of 15~June~2026}).
    Added superior temporal gyrus, inferior frontal gyrus (Broca area), olfactory bulb, piriform cortex, medial septum, lateral/medial habenula, claustrum, primary auditory cortex, prelimbic and infralimbic cortex, perirhinal and postrhinal cortex, dorsal/ventral striatum, zona incerta, retina, spinal cord, pedunculopontine nucleus, dorsal raphe, and inferior olive. Ontology tests enforce no duplicate IDs and minimum alias coverage.

    \item \textbf{Fix task ontology: add 9 missing high-frequency task IDs.}
    Nine task IDs in the corpus have no entry in \texttt{behavioral\_task\_ontology.yaml}: \texttt{visual\_stimulation} (492 datasets), \texttt{decision\_making} (196), \texttt{change\_detection} (121), \texttt{passive\_viewing} (72), \texttt{spontaneous\_activity}, \texttt{circuit\_mapping}, \texttt{cell\_type\_mapping}, \texttt{current\_injection}, \texttt{excitability\_analysis}. Adding these entries raises the indexed task space from 87 to 96 IDs.

    \item \textbf{NLP task enrichment second pass.}
    A batch script matches dataset titles and descriptions against the expanded 96-entry task alias map. Target: raise task coverage from 23.4\% to $\geq 35\%$ using confidence=0.60 inferred labels.

    \item \textbf{Promote silver qrels into human-reviewed benchmark strata}.
    The 13,654 non-error LLM-judged (query, dataset) evidence pairs now support silver diagnostic NDCG@10/MRR/Recall@50 metrics across the ablation ladder. The next step is a stratified dual-human review pass over query intents, sources, positive grades, hard negatives, and graph-moved candidates. Publication-grade NDCG/MRR/Recall requires human-adjudicated qrels.

    \item \textbf{Expose knowledge graph similar-datasets API.}
    Surface the 58,521 cross-dataset edges (\texttt{same\_region\_cross\_modality}: 48,212; \texttt{same\_task\_cross\_species}: 10,309) via \texttt{GET /api/datasets/\{id\}/similar}. A Similar Datasets panel on each dataset card will present these neighbors with their relation type and weight.
\end{enumerate}

\paragraph{Horizon 2 — Short term (corpus coverage + finding quality, 2--4 weeks).}

\begin{enumerate}
    \setcounter{enumi}{12}
    \item \textbf{NeuroVault region enrichment.}
    2,000 records at 12\% region coverage. Strategy: NLP on collection descriptions + NeuroVault brain map metadata + linked paper abstracts via CrossRef. Target: $\geq 35\%$ region coverage.

    \item \textbf{OpenNeuro BIDS electrode enrichment.}
    Run \texttt{enrich\_openneuro\_bids\_regions.py} over the 299 OpenNeuro records. iEEG/sEEG/ECoG datasets will have electrode-table anatomical labels mapped to canonical region IDs. Current: 20\% region coverage; target: $\geq 50\%$ for iEEG datasets.

    \item \textbf{Enable SPECTER2 hybrid retrieval.}
    After SPECTER2 corpus embeddings are built ($\sim$2h GPU), enable \texttt{specter2.enabled: true} in the retrieval config. Evaluate whether paper-embedding similarity improves NDCG@10 over the baseline in the frozen benchmark.

    \item \textbf{Expand corpus to 10K+ records.}
    Priority acquisition targets from the coverage gap analysis: retina $\times$ fMRI datasets (0 in corpus), septum $\times$ fMRI (0), barrel cortex $\times$ fMRI (0), spinal cord datasets (3), CLARITY/lightsheet brain datasets (\textless 5). Sources: IBL additional releases, EBRAINS, GIN deep crawl, SPARK expansion.

    \item \textbf{Benchmark first human-backed NDCG report.}
    With $\geq 50$ adjudicated qrels, run the full ablation suite (\texttt{bm25}, \texttt{usefulness}, \texttt{concept\_boost}, \texttt{specter2}, \texttt{diversity}) and compute preliminary NDCG@10, MRR, HN-VIOL with 95\% bootstrap confidence intervals. A publication-grade report should meet the larger benchmark thresholds in Section~\ref{sec:validation-protocol}.
\end{enumerate}

\paragraph{Horizon 3 — Medium term (search quality + spatial ontology, 4--8 weeks).}

\begin{enumerate}
    \setcounter{enumi}{17}
    \item \textbf{Graph-based result explanation.}
    The knowledge graph now encodes \texttt{same\_region\_cross\_modality} and \texttt{same\_task\_cross\_species} edges between 7,171 datasets. Surface these as similarity explanations: ``2 datasets use this region but with different recording modalities'', ``3 datasets share this task across species''.

    \item \textbf{Metapath retrieval over the full corpus graph.}
    Implement metapath-based similarity scoring using the full 81,518-edge graph:
    \begin{equation}
        S_{\text{meta}}(d, q) = \sum_{\mathcal{P}} w_\mathcal{P} \cdot \text{PathSim}(d, q \mid \mathcal{P})
    \end{equation}
    with paths $\mathcal{P}_{\text{task}} = \text{Dataset} \xrightarrow{\texttt{has\_task}} \text{Task} \xleftarrow{\texttt{has\_task}} \text{Dataset}$, etc. The new cross-dataset edges provide direct single-hop similarity.

    \item \textbf{LLM query expansion in production.}
    Enable \texttt{llm\_expansion.enabled: true} after ablation shows it improves recall for ambiguous queries without hurting precision on structured queries. A/B test against a rule-only baseline.

    \item \textbf{Learned-to-rank reranker.}
    After collecting $\geq 300$ adjudicated qrels, train a calibrated reranker $f_\theta(\mathbf{x}_{d,q})$ using the 10-dimensional feature vector (modality, task, species, region, affordance, standards, session scale, trial scale, graph proximity, SPECTER2). LambdaMART or a shallow MLP trained on graded NDCG loss.

    \item \textbf{Neural signature extraction (long-shot, $s_{10}$).}
    Stream representative NWB assets to extract population-level neural signatures (mean firing rate distributions, calcium transient shapes, LFP power spectra). Store as dataset-level fingerprints. Enable $s_{10}$ cosine similarity in the usefulness scorer.

    \item \textbf{Spatial ontology at scale.}
    Integrate Allen Common Coordinate Framework v3 (13,000+ named regions), UBERON cross-species ontology ($\sim$3,000 neuroscience terms for mouse/rat/macaque/human/zebrafish alignment), and HCP Multi-Modal Parcellation (360 human cortical areas). New KG edges: \texttt{region\_is\_part\_of} (parent-child hierarchy), \texttt{region\_overlaps} (cross-atlas), enabling traversal from whole-brain down to individual layers and subregions. Queries like ``layer~5 pyramidal cells'' would resolve through the hierarchy to match findings from multiple anatomical systems.

    \item \textbf{Tier~2 and tier~3 expansion.}
    Ingest $\sim$1.39M open-access neuroscience papers (tier~2) and eventually $\sim$4.36M total (tier~3). Tier~2 finding extraction: $\sim$210h at current rate of 0.54s/paper. Acceleration strategies: pre-filter by abstract quality (skip $<$50 words), cloud GPU burst (A100, $\sim$\$60 for 40h), or parallel Ollama instances. Tier~3 strategy: extract only for papers with DOI-linked corpus datasets.

    \item \textbf{Unified search with finding cards.}
    Merge paper, finding, and dataset results into a single ranked response. Render \texttt{FindingRecord} objects as search cards with finding text, result direction badge ($\uparrow$/$\downarrow$/$\sim$), region/species/modality chips, confidence meter, and ``View paper'' / ``View dataset'' links. Add a KG-driven ``Related findings'' panel from multi-hop traversal.
\end{enumerate}

\paragraph{Horizon 4 — Long term (informative and cross-scale search, 8+ weeks).}

\begin{enumerate}
    \setcounter{enumi}{25}
    \item \textbf{Dataset comparison view.}
    Given two or more datasets, render a structured comparison across all 18 experimental slots: matched dimensions, complementary dimensions, contradicting dimensions, affordance overlap, paper overlap, cross-species homologs. Surfaces the \texttt{same\_region\_cross\_modality} and \texttt{same\_task\_cross\_species} graph edges as structured explanations.

    \item \textbf{Set-coverage scoring in production.}
    Rank result \emph{sets} rather than individual datasets using the multi-dataset set-coverage function (Section~\ref{sec:setcoverage}). Each result card is assigned a role (ANCHOR, REPLICATION, CROSS-SPECIES COMPARATOR, etc.) so the user understands why each dataset was included.

    \item \textbf{Analysis recipe matching.}
    Link datasets to analysis notebooks/recipes based on affordance overlap. A dataset with \texttt{affordance\_spike\_sorting: high} and \texttt{affordance\_neural\_decoding: medium} surfaces relevant NWB notebook templates and starter code.

    \item \textbf{Prospective case studies and longitudinal reuse tracking.}
    Contact researchers who downloaded or cited retrieved datasets. Measure whether the surfaced result was actually reused, what analyses were run, and what would have made discovery faster. These outcomes ground the usefulness score in actual scientific behavior.
\end{enumerate}

\subsection{Sparse + Dense Hybrid Retrieval}

Implement BM25-based candidate generation with RRF fusion:
\begin{equation}
    \text{candidates} = \RRF(\BM_{\text{field}}(\cdot, \query), \text{Dense}(\cdot, \query), \text{Graph}(\cdot, \query))
\end{equation}

\subsection{Metapath-Based Explanations}

Enhance graph reasoning with explicit metapath templates and confidence-weighted path scoring:
\begin{equation}
    S_{\text{metapath}}(d, \query) = \sum_{\mathcal{P} \in \text{templates}(\query)} w_\mathcal{P} \cdot \max_{\text{path} \in \mathcal{P}(d)} \confidence(\text{path})
\end{equation}

\subsection{Learned-to-Rank Integration}

After sufficient human relevance labels, train calibrated rerankers:
\begin{equation}
    \score_{\text{learned}}(d, \query) = f_\theta\left(\mathbf{x}_{d,\query}\right)
\end{equation}
where $\mathbf{x}_{d,\query}$ is the feature vector and $f_\theta$ is learned (e.g., LambdaMART, neural ranker).

\subsection{Graph Proximity}

The $s_9$ graph proximity dimension computes PathSim when a \texttt{KnowledgeGraph} is available:
\begin{equation}
    s_9(C_q, C_d) = \text{PathSim}_{\mathcal{P}}(\text{id}(C_q), \text{id}(C_d) \mid \graph)
\end{equation}

\subsection{Dense Embedding Stack}

The current expanded-corpus artifact has 7,171 normalized rows, and the current BGE-large-en-v1.5 dense field embedding cache covers all 7,171 records at 1,024 dimensions. The canonical fast-load files are \texttt{real\_all.dense.field\_embeddings.agg.ids.json} and \texttt{real\_all.dense.field\_embeddings.agg.mat.npy}; the matrix shape is $7{,}171 \times 1024$. \texttt{scripts/eval/run\_ablation\_ladder.py} now checks for this NPY cache before parsing JSONL, so normal ablation runs avoid reparsing the 82 MB field-embedding JSONL. The older turbovec vector file still has shape $625 \times 1024$ with 4-bit metadata, so ANN recall and latency reports remain engineering validation until that approximate index is rebuilt from the full dense cache. A key implementation note: the \texttt{\_provider\_for\_cache} function in \texttt{field\_semantic.py} must register a \texttt{DenseEmbeddingProvider} for \texttt{provider\_name == "bge-large"} to ensure query-time embeddings are computed in the same 1024-dim vector space as the corpus.

\subsection{Corpus Expansion and Graph-Linked Memory Growth}

The current live corpus artifact contains 7,171 normalized rows and 7,121 unique identifiers. Major sources include NeuroVault (2,000), NeuroMorpho (1,000), DANDI (848), Zenodo (500), Harvard Dataverse (500), Allen (500), GIN (408), OpenNeuro (299), Figshare (200), OSF (200), Brain Image Library (200), IBL (198), CRCNS (153), BlueBrain (100), Buzsaki Lab (35), SPARK (20), and NEMO (10). The Tier-2 ingestion pipeline uses a 4-gate inclusion classifier (raw/processed data, neuroscience signal, open license, persistent identifier), deduplication, and rejection logging. The current report bundle contains a Tier-2 rejection summary of 24,160 candidates, but the local rejection JSONL and corpus-quality report need regeneration before this can be cited as a final quality-gate result.

The expansion machinery separates breadth from evidence depth. High-volume general repositories such as Zenodo, Figshare, and OSF increase recall but are treated as higher-risk heterogeneous sources that require source-skew analysis and off-topic QA. Curated neuroscience sources such as DANDI, OpenNeuro, NeuroVault, NeuroMorpho, IBL, CRCNS, Allen, Brain Image Library, and BlueBrain provide more domain-specific metadata and stronger modality or anatomical priors. Source-specific scripts can append IBL, CRCNS, and NeuroVault records without a full rebuild, while DANDI enrichment scripts refresh rich metadata, study targets, NWB electrode locations, and NWB surgery/experiment-description text for records with missing modality, species, or region fields.

Regional expansion is managed by evidence tier. Structured repository metadata, NWB electrode-table locations, and curated source categories are treated as stronger evidence than title or abstract mentions; CrossRef paper abstracts and free-text NWB fields are retained as silver labels. The regional coverage map exposes verified regions separately from candidate-only mentions and produces a regionless review queue. This prevents the system from inflating anatomical coverage while still giving curators actionable acquisition targets.

Corpus expansion and memory graph growth are now coupled: adding or changing records triggers incremental hash-based update detection, inserts or refreshes dataset/modality/species/region/task/affordance nodes, and connects them into the existing graph through named provenance-carrying edges. The full-corpus knowledge graph rebuild (30~June~2026) produced 7,438 nodes and 81,518 edges from all 7,171 records, with real corpus dataset IDs and 58,521 cross-dataset relationship edges. The separate field-state memory graph remains at 2,200 nodes/3,788 edges for the reviewed 629-record slice and writes versioned snapshots without altering human overlays or adjudicated labels. The expansion plan targets 50K+ records through deeper DANDI and OpenNeuro coverage, EBRAINS, SPARK, UK Biobank (neuroscience subset), and data paper mining from \textit{Scientific Data} and \textit{eLife}. Each new source should be accompanied by a normalizer extension for any new modalities, species, or brain regions it introduces, and validated against the conservative normalizer guardrails before ingestion.

\subsection{Neural Signature Similarity}

The $s_{10}$ dimension is currently a placeholder ($0.0$). Future work should extract neural population signatures directly from NWB/BIDS data:
\begin{equation}
    s_{10}(C_q, C_d) = \cos\!\left(\text{Extractor}(\text{NWB}(C_d)),\, \text{Extractor}(\text{NWB}(C_q))\right)
\end{equation}

\subsection{Human-Labeled Benchmark Calibration}

The current repository now has a canonical LLM-judged qrels snapshot large enough to exercise the retrieval metrics and regression gate, but the paper still needs a multi-annotator human benchmark before making final retrieval claims. The next benchmark step should duplicate-judge a stratified subset of the 317 canonical queries, adjudicate disagreements, and grow expert coverage as failure analysis identifies undersampled intents and sources. The benchmark should enable:
\begin{itemize}
    \item Calibrated usefulness score distributions
    \item Learned reranker training ($f_\theta$) on feature vectors
    \item Statistical significance testing across ablation variants
\end{itemize}

% ============================================================================
% SECTION 11: CONCLUSION
% ============================================================================
\section{Conclusion}
\label{sec:conclusion}

Neural Search represents a principled approach to scientific dataset discovery that goes beyond generic document retrieval. By formalizing datasets as structured experimental objects, grounding search in ontological knowledge, propagating provenance through typed graphs, and maintaining interpretable scoring, the system is designed to help researchers find reusable experimental contexts rather than merely similar text.

The mathematical foundations---spanning information retrieval theory, knowledge graph theory, representation learning, constraint satisfaction, and evaluation methodology---provide a rigorous substrate for continued development. The current live implementation contains a 7,171-row normalized corpus, a full-corpus dense field embedding cache covering all 7,171 records, a full-corpus knowledge graph with 7,438 nodes and 81,518 edges (including 58,521 explicit cross-dataset relationship edges), a DuckDB coverage ledger with 29,561 dimension entries, a coverage gap boost, a source diversity reranker, config-gated SPECTER2 fusion and LLM query expansion, structured dimension-match badges in the frontend, a 317-query canonical benchmark suite with 13,654 LLM-judged qrels, bootstrap CI infrastructure, and a complete annotation pool generator. A literature-scale ingestion layer has ingested 255,940 OpenAlex tier~1 neuroscience papers and run the paper-dataset linker over all 7,171 corpus datasets; current successful links are 168 DOI-exact and 225 title-fuzzy-local, with 6,778 not-found records and a precision audit pending. LLM-powered structured finding extraction is running on local inference and remains in progress, with 190K+ tier~1 findings extracted as of 23~June~2026. A rule-based enrichment pass adds 27 typed fields (frequency band, temporal pattern, negation, spatial frame, and more), enabling compound queries over disease model, brain region, frequency band, molecular marker, pharmacological context, sleep state, and analysis method; these finding counts and coverage rates are operational estimates, not validated finding quality. The June 30 graph rebuild fixes the previous demo-ID mismatch, so real retrieval IDs now receive nonzero graph-context scores. The BGE-large usefulness-score Spearman correlation is 0.3999 over 270 labeled pairs, and graph proximity changes 39\% of ranked pairs in the current ablation artifact.

The next scientific milestone is reliability: dual-judge and human-adjudicate a stratified subset of the canonical qrels so QWK, disagreement analysis, and expert-calibrated NDCG@10 can be reported. The second --- isolating the typed/relationship knowledge-graph layer's contribution from the aggregate \texttt{hybrid\_graph} signal --- is now done: dedicated \texttt{typed\_kg} and \texttt{typed\_kg\_qualified} ablation rungs (2026-06-23) add only the typed finding-relationship score (supports/contradicts edges and a qualified-consensus region bonus) on top of \texttt{hybrid\_rrf}. On a full-corpus replication run, neither rung moved NDCG@10 measurably beyond \texttt{hybrid\_rrf} (\texttt{typed\_kg\_qualified} moved it by +0.0002), consistent with the existing relationship-edge quality analysis's near-coin-flip (47--52\%) helpful rate. This is a second, independent line of evidence that the typed/relationship layer's retrieval contribution is real but small at current calibration --- not yet a reason to invest further in relationship-edge expansion ahead of human-adjudicated qrels, and a reason to re-run the same isolated rungs on the canonical packet-aligned corpus before treating the null result as final. The third is coverage: spatial ontology expansion (Allen CCF v3, 13K+ regions; UBERON cross-species; HCP MMP 360 cortical areas), tier~2 ingest ($\sim$1.39M open-access papers, staged and ready via \texttt{bulk\_ingest\_openalex.py --tier tier2} but not yet started), BIL and NeuroVault re-enrichment, OpenNeuro BIDS electrode extraction, and corpus growth toward 10K+ records will close the dark-pair gaps identified by the coverage ledger. The fourth is accuracy: SPECTER2 fusion, metapath retrieval, additional graph rungs, and a learned-to-rank reranker will be evaluated against the frozen qrels benchmark before any production deployment.

The key insight remains: neuroscience search should retrieve reusable experimental contexts, not just documents. Neural Search now has the retrieval architecture, full-corpus knowledge graph, literature-scale paper-finding linkage, coverage analysis infrastructure, and validation harness needed to test that principle honestly and iteratively---across every brain region, measurement modality, and timescale in the published record.

\appendix
\section{Field-State Validation Artifacts}
\label{app:field-state-artifacts}

The Field-State layer writes compact, reviewable artifacts rather than a large research-assistant database. Core JSONL artifacts are:

\begin{itemize}
    \item \texttt{artifacts/field\_state/claims.jsonl}
    \item \texttt{artifacts/field\_state/benchmark\_gaps.jsonl}
    \item \texttt{artifacts/field\_state/opportunities.jsonl}
    \item \texttt{artifacts/field\_state/qrels\_candidates.jsonl}
    \item \texttt{artifacts/field\_state/qrels\_reviews.jsonl}
    \item \texttt{artifacts/field\_state/adjudicated\_qrels.jsonl}
    \item \texttt{artifacts/field\_state/qrels\_agreement.json}
    \item \texttt{artifacts/field\_state/eval\_snapshot.json}
    \item \texttt{artifacts/field\_state/claim\_evidence\_suggestions.jsonl}
\end{itemize}

Memory graph artifacts (v0.9):

\begin{itemize}
    \item \texttt{artifacts/field\_state/memory\_graph\_nodes.jsonl} --- 2,200 typed nodes in build \texttt{update\_20260613T053153Z}
    \item \texttt{artifacts/field\_state/memory\_graph\_edges.jsonl} --- 3,788 typed edges in build \texttt{update\_20260613T053153Z}
    \item \texttt{artifacts/field\_state/memory\_graph\_manifest.json} --- build summary
    \item \texttt{artifacts/field\_state/current\_manifest.json} --- latest snapshot pointer
    \item \texttt{artifacts/field\_state/snapshots/<timestamp>/} --- versioned snapshot directories
    \item \texttt{artifacts/field\_state/learning\_signals.jsonl} --- provisional learning signals
\end{itemize}

Full-corpus knowledge graph artifacts (rebuilt 30~June~2026):

\begin{itemize}
    \item \texttt{data/graph/neural\_search\_graph.real\_corpus.json} --- 7,438 nodes / 81,518 edges
    \item \texttt{scripts/build\_real\_corpus\_graph.py} --- builder from flat corpus using real corpus dataset IDs and aliases
    \item \texttt{data/embeddings/real\_all.dense.field\_embeddings.agg.ids.json} and \texttt{data/embeddings/real\_all.dense.field\_embeddings.agg.mat.npy} --- 7,171-record aggregated dense embedding cache
    \item \texttt{scripts/eval/cache\_embeddings\_npy.py} --- cache builder for fast full-corpus dense embedding loading
\end{itemize}

Literature layer artifacts (added 17~June~2026):

\begin{itemize}
    \item \texttt{data/corpus/normalized/openalex\_neuro/} --- 255,940 tier~1 OpenAlex neuroscience papers in 26 JSONL shards \citep{priem2022openalex}
    \item \texttt{artifacts/literature/paper\_dataset\_links.jsonl} --- paper-dataset linker output for all 7,171 corpus datasets, with 168 DOI-exact links, 225 title-fuzzy-local links, and 6,778 not-found records
    \item \texttt{artifacts/literature/findings\_tier1\_ollama.jsonl} --- LLM-extracted \texttt{FindingRecord} objects from tier~1 abstracts (extraction running)
    \item \texttt{artifacts/literature/findings\_tier1\_ollama.checkpoint.json} --- batch-level checkpoint enabling resume without reprocessing
    \item \texttt{neural\_search/literature/finding\_extractor.py} --- \texttt{extract\_batch\_ollama()}, \texttt{\_repair\_json()}, \texttt{FindingRecord}, \texttt{LLMProviderConfig}
    \item \texttt{neural\_search/literature/kg\_builder.py} --- literature KG builder (paper/finding/venue nodes and cross-edges)
    \item \texttt{neural\_search/literature/search.py} --- BM25 TF-IDF search across papers and findings
    \item \texttt{neural\_search/ingestion/openalex\_bulk.py} --- cursor-based bulk downloader with tier filtering and checkpoint/resume
    \item \texttt{neural\_search/literature/linking.py} --- DOI exact match + title fuzzy match paper-to-dataset linker
    \item \texttt{scripts/literature/extract\_findings.py} --- CLI for corpus-level extraction with Ollama auto-discovery
\end{itemize}

Coverage ledger and annotation artifacts:

\begin{itemize}
    \item \texttt{data/coverage/ledger.duckdb} --- 7,176 datasets / 29,561 entries / 6 dimensions
    \item \texttt{data/eval/benchmark\_queries\_v2.yaml} --- 100 benchmark queries
    \item \texttt{reports/eval/annotation\_pool.jsonl} --- stratified (query, dataset) pairs for annotation
    \item \texttt{reports/eval/annotation\_queries.jsonl} --- query metadata for annotation tool
\end{itemize}

Generated Markdown reports include weak claims, benchmark gaps, ranked opportunities, latest snapshots, qrels agreement summaries, evaluation snapshots, claim-evidence updates, whitepaper validation reports, memory graph validation, and learning signal summaries under \texttt{reports/field\_state/}.

\subsection{Qrels and Adjudication Format}

Each qrels candidate records a query identifier, dataset identifier, source rank, retrieval score, candidate rationale, and optional hard-negative rationale. Imported human reviews record reviewer identity, a 0--3 relevance label, confidence, notes, evidence links, and review status. Adjudicated qrels are written separately so that raw reviewer judgments remain auditable.

\subsection{Obsidian Roundtrip}

The Obsidian-compatible vault export keeps generated metadata and protected human notes in separate blocks. The current smoke workflow exported 10 qrels review notes, imported unreviewed overlays, and validated the memory roundtrip. This proves the review workflow is reproducible, but it does not prove relevance quality until human labels are completed and adjudicated.

% ============================================================================
% REFERENCES
% ============================================================================
\bibliographystyle{plainnat}
\begin{thebibliography}{10}

\bibitem[Cormack et~al.(2009)]{cormack2009reciprocal}
Cormack, G.~V., Clarke, C.~L., and Buettcher, S.
\newblock Reciprocal rank fusion outperforms condorcet and individual rank learning methods.
\newblock In \emph{Proceedings of SIGIR}, pages 758--759, 2009.

\bibitem[Robertson and Zaragoza(2009)]{robertson2009probabilistic}
Robertson, S. and Zaragoza, H.
\newblock The probabilistic relevance framework: BM25 and beyond.
\newblock \emph{Foundations and Trends in Information Retrieval}, 3(4):333--389, 2009.

\bibitem[Salton et~al.(1975)]{salton1975vector}
Salton, G., Wong, A., and Yang, C.~S.
\newblock A vector space model for automatic indexing.
\newblock \emph{Communications of the ACM}, 18(11):613--620, 1975.

\bibitem[Sun et~al.(2011)]{sun2011pathsim}
Sun, Y., Han, J., Yan, X., Yu, P.~S., and Wu, T.
\newblock PathSim: Meta path-based top-k similarity search in heterogeneous information networks.
\newblock \emph{Proceedings of the VLDB Endowment}, 4(11):992--1003, 2011.

\bibitem[J{\"a}rvelin and Kek{\"a}l{\"a}inen(2002)]{jarvelin2002cumulated}
J{\"a}rvelin, K. and Kek{\"a}l{\"a}inen, J.
\newblock Cumulated gain-based evaluation of IR techniques.
\newblock \emph{ACM Transactions on Information Systems}, 20(4):422--446, 2002.

\bibitem[Guo et~al.(2017)]{guo2017calibration}
Guo, C., Pleiss, G., Sun, Y., and Weinberger, K.~Q.
\newblock On calibration of modern neural networks.
\newblock In \emph{Proceedings of ICML}, pages 1321--1330, 2017.

\bibitem[Wilkinson et~al.(2016)]{wilkinson2016fair}
Wilkinson, M.~D., Dumontier, M., Aalbersberg, I.~J., Appleton, G., Axton, M., Baak, A., et~al.
\newblock The FAIR guiding principles for scientific data management and stewardship.
\newblock \emph{Scientific Data}, 3:160018, 2016.

\bibitem[Gorgolewski et~al.(2016)]{gorgolewski2016bids}
Gorgolewski, K.~J., Auer, T., Calhoun, V.~D., Craddock, R.~C., Das, S., Duff, E.~P., et~al.
\newblock The Brain Imaging Data Structure, a format for organizing and describing outputs of neuroimaging experiments.
\newblock \emph{Scientific Data}, 3:160044, 2016.

\bibitem[Markiewicz et~al.(2021)]{markiewicz2021openneuro}
Markiewicz, C.~J., et~al.
\newblock The OpenNeuro resource for sharing of neuroscience data.
\newblock \emph{eLife}, 10:e71774, 2021.

\bibitem[DANDI Archive(2026)]{dandiArchive2026}
DANDI Archive.
\newblock DANDI Archive documentation and repository portal.
\newblock \url{https://www.dandiarchive.org/}, accessed June 2026.

\bibitem[R{\"u}bel et~al.(2022)]{ruebel2022nwb}
R{\"u}bel, O., Tritt, A., Ly, R., Dichter, B.~K., Ghosh, S., Niu, L., et~al.
\newblock The Neurodata Without Borders ecosystem for neurophysiological data science.
\newblock \emph{eLife}, 11:e78362, 2022.

\bibitem[Thakur et~al.(2021)]{thakur2021beir}
Thakur, N., Reimers, N., R{\"u}ckl{\'e}, A., Srivastava, A., and Gurevych, I.
\newblock BEIR: A heterogeneous benchmark for zero-shot evaluation of information retrieval models.
\newblock In \emph{Proceedings of NeurIPS Datasets and Benchmarks}, 2021.

\bibitem[Muennighoff et~al.(2022)]{muennighoff2022mteb}
Muennighoff, N., Tazi, N., Magne, L., and Reimers, N.
\newblock MTEB: Massive text embedding benchmark.
\newblock \emph{arXiv preprint arXiv:2210.07316}, 2022.

\bibitem[Singh et~al.(2023)]{specter2_2023}
Singh, A., Cohan, A., Feldman, S., Downey, D., and Weld, D.~S.
\newblock SPECTER2: Ultra-light, fast and cost-efficient paper embeddings using highly efficient sentence transformers.
\newblock \emph{arXiv preprint arXiv:2304.05457}, 2023.

\bibitem[Priem et~al.(2022)]{priem2022openalex}
Priem, J., Piwowar, H., and Orr, R.
\newblock OpenAlex: A fully-open index of the world's research works.
\newblock \emph{arXiv preprint arXiv:2205.01833}, 2022.

\bibitem[Qwen Team(2024)]{qwen2024qwen25}
Qwen Team.
\newblock Qwen2.5 Technical Report.
\newblock \emph{arXiv preprint arXiv:2412.15115}, 2024.

\bibitem[Wang et~al.(2024)]{wang2024brainknow}
Wang, X., et~al.
\newblock BrainKnow: Extracting, filtering, and generating neuroscience knowledge using large language models.
\newblock \emph{arXiv preprint arXiv:2403.04346}, 2024.

\end{thebibliography}

% ============================================================================
% APPENDIX
% ============================================================================
\appendix

\section{Algorithm Pseudocode}
\label{app:algorithms}

\begin{algorithm}
\caption{Neural Search Hybrid Retrieval}
\begin{algorithmic}[1]
\Procedure{Search}{$\query, \text{corpus}, \text{config}$}
    \State $\text{intent} \gets \Call{ClassifyIntent}{\query}$
    \State $\mathbf{w} \gets \Call{GetWeights}{\text{intent}, \text{config}}$
    \State $\text{parsed} \gets \Call{ParseQuery}{\query}$
    \State $\text{candidates} \gets \Call{GenerateCandidates}{\text{parsed}, \text{corpus}}$
    \State $\text{candidates} \gets \Call{FilterHardNegatives}{\text{candidates}, \text{parsed}}$
    \For{$d \in \text{candidates}$}
        \State $\text{scores}[d] \gets \Call{ComputeHybridScore}{d, \text{parsed}, \mathbf{w}}$
        \State $\text{explanations}[d] \gets \Call{GenerateExplanation}{d, \text{parsed}}$
    \EndFor
    \State $\text{ranked} \gets \Call{Sort}{\text{candidates}, \text{scores}}$
    \State \Return $\text{ranked}, \text{explanations}$
\EndProcedure
\end{algorithmic}
\end{algorithm}

\begin{algorithm}
\caption{BM25 Field-Weighted Scoring}
\begin{algorithmic}[1]
\Procedure{BM25Field}{$d, \query, \text{fields}, \mathbf{w}_f$}
    \State $\text{score} \gets 0$
    \For{$f \in \text{fields}$}
        \State $\text{score}_f \gets 0$
        \For{$t \in \query$}
            \State $\text{tf} \gets \Call{TermFreq}{t, d_f}$
            \State $\text{idf} \gets \Call{InverseDocFreq}{t}$
            \State $\text{len\_norm} \gets 1 - b + b \cdot |d_f| / \text{avgdl}_f$
            \State $\text{score}_f \gets \text{score}_f + \text{idf} \cdot \frac{\text{tf} \cdot (k_1 + 1)}{\text{tf} + k_1 \cdot \text{len\_norm}}$
        \EndFor
        \State $\text{score} \gets \text{score} + w_f \cdot \text{score}_f$
    \EndFor
    \State \Return $\text{score}$
\EndProcedure
\end{algorithmic}
\end{algorithm}

\begin{algorithm}
\caption{Reciprocal Rank Fusion}
\begin{algorithmic}[1]
\Procedure{RRF}{$\text{ranked\_lists}, k$}
    \State $\text{scores} \gets \{\}$
    \For{$R \in \text{ranked\_lists}$}
        \For{$i, d \in \Call{Enumerate}{R}$}
            \State $\text{scores}[d] \gets \text{scores}[d] + \frac{1}{k + i + 1}$
        \EndFor
    \EndFor
    \State \Return $\Call{SortByValue}{\text{scores}}$
\EndProcedure
\end{algorithmic}
\end{algorithm}

\section{Graph Schema Definition}
\label{app:schema}

\begin{lstlisting}[language=Python, caption=Knowledge Graph Schema (Pydantic)]
class NodeType(str, Enum):
    DATASET = "dataset"
    PAPER = "paper"        # OpenAlex/CrossRef publication record
    FINDING = "finding"    # LLM-extracted structured finding (FindingRecord)
    VENUE = "venue"        # Journal or conference
    TASK = "task"
    MODALITY = "modality"
    SPECIES = "species"
    REGION = "region"
    EVENT = "event"
    ANALYSIS = "analysis"
    STANDARD = "standard"
    REQUIREMENT = "requirement"

class EdgeType(str, Enum):
    HAS_TASK = "has_task"
    HAS_MODALITY = "has_modality"
    RECORDS_REGION = "records_region"
    HAS_SPECIES = "has_species"
    HAS_EVENT = "has_event"
    USES_STANDARD = "uses_standard"
    SUPPORTS_ANALYSIS = "supports_analysis"
    REQUIRES = "requires"
    LINKED_TO_PAPER = "linked_to_paper"
    SIMILAR_TO = "similar_to"
    # Literature layer edges (added 17 June 2026)
    PAPER_REPORTS_FINDING = "paper_reports_finding"
    FINDING_INVOLVES_REGION = "finding_involves_region"
    FINDING_INVOLVES_TASK = "finding_involves_task"
    FINDING_INVOLVES_MODALITY = "finding_involves_modality"
    FINDING_INVOLVES_SPECIES = "finding_involves_species"
    FINDING_INVOLVES_CELL_TYPE = "finding_involves_cell_type"
    FINDING_INVOLVES_MOLECULE = "finding_involves_molecule"
    DATASET_LINKED_TO_PAPER = "dataset_linked_to_paper"
    PAPER_PUBLISHED_IN = "paper_published_in"

class GraphEvidence(BaseModel):
    source_type: str
    source_id: str
    source_field: Optional[str]
    evidence_text: Optional[str]
    extractor: str
    version: str
    confidence: float  # [0, 1]

class KnowledgeGraphNode(BaseModel):
    id: str
    node_type: NodeType
    label: str
    properties: Dict[str, Any]
    evidence: List[GraphEvidence]

class KnowledgeGraphEdge(BaseModel):
    id: str
    source: str
    target: str
    edge_type: EdgeType
    confidence: float
    evidence: List[GraphEvidence]
\end{lstlisting}

\end{document}
