% ============================================================================
% NEURAL SEARCH: ICLR-LEVEL WHITEPAPER
% A comprehensive scientific and mathematical foundation
% ============================================================================
\documentclass{article}

% ICLR 2024 style formatting
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{hyperref}
\usepackage{url}
\usepackage{booktabs}
\usepackage{amsfonts}
\usepackage{nicefrac}
\usepackage{microtype}
\usepackage{xcolor}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{mathtools}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{tikz}
\usetikzlibrary{arrows.meta,shapes,positioning,calc,fit,backgrounds,decorations.pathmorphing,patterns,matrix,chains,scopes,shadows}
\usepackage{pgfplots}
\pgfplotsset{compat=1.17}
\usepgfplotslibrary{groupplots,fillbetween}
\usepackage{tcolorbox}
\tcbuselibrary{skins,breakable}
\usepackage{multirow}
\usepackage{tabularx}
\usepackage{enumitem}
\usepackage{natbib}
\usepackage{float}
\usepackage{wrapfig}

% Page geometry - conference style
\usepackage[margin=1in]{geometry}

% ============================================================================
% THEOREM ENVIRONMENTS
% ============================================================================
\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{remark}{Remark}[section]
\newtheorem{example}{Example}[section]

% ============================================================================
% CUSTOM COMMANDS
% ============================================================================
\newcommand{\dataset}{\mathcal{D}}
\newcommand{\query}{\mathbf{q}}
\newcommand{\score}{\text{score}}
\newcommand{\embed}{\phi}
\newcommand{\graph}{\mathcal{G}}
\newcommand{\nodes}{\mathcal{V}}
\newcommand{\edges}{\mathcal{E}}
\newcommand{\rel}{\text{rel}}
\newcommand{\confidence}{\gamma}
\newcommand{\evidence}{\mathcal{E}}
\newcommand{\fingerprint}{\mathbf{f}}
\newcommand{\compatibility}{\kappa}
\newcommand{\RRF}{\text{RRF}}
\newcommand{\BM}{\text{BM25}}
\newcommand{\NDCG}{\text{NDCG}}
\newcommand{\MRR}{\text{MRR}}
\newcommand{\sigmoid}{\sigma}
\newcommand{\R}{\mathbb{R}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\Prob}{\mathbb{P}}

% Colors for figures
\definecolor{datacolor}{RGB}{66,133,244}
\definecolor{graphcolor}{RGB}{52,168,83}
\definecolor{searchcolor}{RGB}{251,188,5}
\definecolor{evalcolor}{RGB}{234,67,53}
\definecolor{corpuscolor}{RGB}{171,71,188}
\definecolor{embedcolor}{RGB}{0,172,193}
\definecolor{ontologycolor}{RGB}{255,112,67}
\definecolor{affordcolor}{RGB}{124,179,66}
\definecolor{lightgray}{RGB}{245,245,245}
\definecolor{darkgray}{RGB}{97,97,97}

% ============================================================================
% TITLE AND AUTHORS
% ============================================================================
\title{
    \textbf{Neural Search: Experiment-Aware Retrieval for Reusable Neuroscience Datasets via Hybrid Semantic Reasoning and Knowledge Graph Traversal}
}

\author{
    Siddharth Shulyalkar\textsuperscript{1}\\
    \\
    \textsuperscript{1}University of California, San Diego\\
    \texttt{shulyalkar@ucsd.edu}\\
    \\
    \textit{Technical Report v1.0 --- May 2026}
}

\date{}

\begin{document}

\maketitle

% ============================================================================
% ABSTRACT
% ============================================================================
\begin{abstract}
The exponential growth of neuroscience datasets across distributed repositories presents a fundamental challenge for scientific discovery: how can researchers find datasets that are not merely textually similar to their query, but scientifically reusable for their specific experimental and analytical goals? We present \textbf{Neural Search}, a novel retrieval system that treats datasets as structured scientific objects rather than documents, enabling experiment-aware discovery through hybrid semantic reasoning over typed metadata, provenance-weighted knowledge graphs, and multi-aspect embeddings. Our approach formalizes the scientific dataset search problem through: (1) a slot-based experimental context representation that captures organism, task, modality, region, and analysis affordances; (2) a heterogeneous knowledge graph with confidence-weighted edges derived from multiple evidence sources; (3) metapath-based reasoning for discovering implicit dataset relationships; and (4) a hybrid scoring function that combines BM25 lexical matching, dense semantic retrieval, and graph-based signals through reciprocal rank fusion. We introduce the concept of \textit{analysis affordances}---computational requirements that datasets must satisfy for specific scientific analyses---and demonstrate how constraint satisfaction over these affordances enables researchers to find datasets that support their intended methodologies. On an initial benchmark of 30 expert-curated neuroscience queries spanning task search, modality filtering, and cross-species comparison, Neural Search achieves 76.7\% Precision@5, 87.8\% Recall@10, 0.950 MRR, and 0.937 NDCG@10 while maintaining zero hard-negative constraint violations. These results demonstrate a working prototype; larger-scale validation with expanded corpora and independent annotators remains ongoing work. We further present a roadmap for extending Neural Search to bridge neuroscience datasets with computational neuroscience models and artificial neural network research, enabling cross-disciplinary discovery at the intersection of biological and artificial intelligence.
\end{abstract}

% ============================================================================
% SYSTEM OVERVIEW FIGURE
% ============================================================================
\begin{figure*}[t]
\centering
\begin{tikzpicture}[
    scale=0.85,
    transform shape,
    node distance=0.8cm,
    box/.style={rectangle, draw, rounded corners=3pt, minimum width=2.2cm, minimum height=0.7cm, font=\small, align=center},
    bigbox/.style={rectangle, draw, rounded corners=5pt, minimum width=3.5cm, minimum height=1.2cm, font=\small\bfseries, align=center},
    source/.style={box, fill=corpuscolor!30, draw=corpuscolor},
    process/.style={box, fill=datacolor!30, draw=datacolor},
    storage/.style={cylinder, draw=graphcolor, fill=graphcolor!20, shape border rotate=90, minimum width=1.2cm, minimum height=0.8cm, font=\small},
    signal/.style={box, fill=searchcolor!30, draw=searchcolor},
    output/.style={box, fill=evalcolor!30, draw=evalcolor},
    arrow/.style={-{Stealth[length=2mm]}, thick},
    dasharrow/.style={-{Stealth[length=2mm]}, thick, dashed},
    label/.style={font=\scriptsize, align=center}
]
    % === LAYER 1: DATA SOURCES ===
    \node[source] (dandi) at (0,0) {DANDI\\Archive};
    \node[source, right=0.5cm of dandi] (openneuro) {OpenNeuro};
    \node[source, right=0.5cm of openneuro] (allen) {Allen Brain\\Observatory};
    \node[source, right=0.5cm of allen] (nemo) {NeMO\\Archive};
    \node[source, right=0.5cm of nemo] (papers) {OpenAlex\\Papers};

    % Layer label
    \node[left=0.3cm of dandi, font=\footnotesize\itshape] {Data Sources};

    % === LAYER 2: INGESTION ===
    \node[process, below=1.2cm of $(dandi)!0.5!(openneuro)$] (nwb) {NWB/BIDS\\Parser};
    \node[process, below=1.2cm of allen] (normalize) {Metadata\\Normalizer};
    \node[process, below=1.2cm of $(nemo)!0.5!(papers)$] (extract) {Entity\\Extractor};

    \node[left=0.3cm of nwb, font=\footnotesize\itshape] {Ingestion};

    % === LAYER 3: KNOWLEDGE REPRESENTATION ===
    \node[storage, below=1.2cm of nwb] (corpus) {Corpus};
    \node[bigbox, fill=graphcolor!20, draw=graphcolor, below=1.2cm of normalize] (kg) {Knowledge\\Graph};
    \node[storage, below=1.2cm of extract] (embed) {Embeddings};

    \node[left=0.3cm of corpus, font=\footnotesize\itshape] {Knowledge};

    % === LAYER 4: RETRIEVAL SIGNALS ===
    \node[signal, below=1.5cm of corpus] (bm25) {BM25\\Lexical};
    \node[signal, below=1.5cm of $(corpus)!0.4!(kg)$] (onto) {Ontology\\Matching};
    \node[signal, below=1.5cm of kg] (graph) {Graph\\Traversal};
    \node[signal, below=1.5cm of $(kg)!0.6!(embed)$] (dense) {Dense\\Semantic};
    \node[signal, below=1.5cm of embed] (afford) {Affordance\\Scoring};

    \node[left=0.3cm of bm25, font=\footnotesize\itshape] {Signals};

    % === LAYER 5: FUSION ===
    \node[bigbox, fill=searchcolor!40, draw=searchcolor!80, below=1.2cm of graph] (fusion) {Reciprocal Rank\\Fusion (RRF)};

    % === LAYER 6: OUTPUT ===
    \node[output, below=1cm of fusion] (results) {Ranked Results\\+ Explanations};

    % === QUERY INPUT ===
    \node[box, fill=gray!30, draw=gray, left=2.5cm of fusion] (query) {Scientific\\Query};
    \node[box, fill=gray!20, draw=gray, below=0.5cm of query] (intent) {Intent\\Classification};

    % === ARROWS: Data flow ===
    \foreach \src in {dandi,openneuro} {
        \draw[arrow, corpuscolor] (\src) -- (nwb);
    }
    \draw[arrow, corpuscolor] (allen) -- (normalize);
    \draw[arrow, corpuscolor] (nemo) -- (normalize);
    \draw[arrow, corpuscolor] (papers) -- (extract);

    \draw[arrow, datacolor] (nwb) -- (corpus);
    \draw[arrow, datacolor] (nwb) -- (kg);
    \draw[arrow, datacolor] (normalize) -- (corpus);
    \draw[arrow, datacolor] (normalize) -- (kg);
    \draw[arrow, datacolor] (extract) -- (kg);
    \draw[arrow, datacolor] (extract) -- (embed);

    \draw[arrow, graphcolor] (corpus) -- (bm25);
    \draw[arrow, graphcolor] (corpus) -- (onto);
    \draw[arrow, graphcolor] (kg) -- (onto);
    \draw[arrow, graphcolor] (kg) -- (graph);
    \draw[arrow, graphcolor] (embed) -- (dense);
    \draw[arrow, graphcolor] (kg) -- (afford);

    \foreach \sig in {bm25,onto,graph,dense,afford} {
        \draw[arrow, searchcolor] (\sig) -- (fusion);
    }

    \draw[arrow, evalcolor] (fusion) -- (results);

    % Query flow
    \draw[arrow, gray] (query) -- (intent);
    \draw[arrow, gray] (intent) -- (fusion);
    \draw[dasharrow, gray] (query) to[out=0,in=180] (bm25);
    \draw[dasharrow, gray] (query) to[out=0,in=180] (dense);

    % === ANNOTATIONS ===
    \node[below=0.1cm of results, font=\scriptsize\itshape, text=darkgray] {Dataset cards with provenance, affordances, and confidence scores};

\end{tikzpicture}
\caption{\textbf{Neural Search System Architecture.} The system ingests heterogeneous neuroscience datasets from multiple repositories, normalizes metadata into a unified schema, constructs a provenance-weighted knowledge graph, and computes multi-aspect embeddings. At query time, five retrieval signals (BM25 lexical, ontology matching, graph traversal, dense semantic, and affordance scoring) are combined via reciprocal rank fusion with intent-adaptive weighting. Results include ranked datasets with explanations, provenance chains, and analysis affordance annotations.}
\label{fig:system-architecture}
\end{figure*}

% ============================================================================
% SECTION 1: INTRODUCTION
% ============================================================================
\section{Introduction}
\label{sec:introduction}

The past decade has witnessed an unprecedented expansion of publicly available neuroscience datasets. Repositories such as DANDI Archive \citep{rubinen2024dandi}, OpenNeuro \citep{markiewicz2021openneuro}, Allen Brain Observatory \citep{allen2019brain}, NeMO Archive \citep{nemo2023archive}, and the International Brain Laboratory \citep{ibl2021standardized} now collectively host thousands of datasets spanning electrophysiology, calcium imaging, fMRI, behavioral recordings, and multi-modal experiments. This wealth of data presents an enormous opportunity for scientific discovery through data reuse, meta-analysis, and computational modeling \citep{poldrack2014making, yarkoni2019large}.

The scale of this opportunity is staggering. DANDI alone hosts over 500 dandisets containing petabytes of neurophysiology data. OpenNeuro provides access to thousands of neuroimaging datasets representing millions of hours of MRI scanner time. The Allen Brain Observatory offers standardized recordings from thousands of neurons across the mouse visual system. Collectively, these repositories represent an investment of hundreds of millions of dollars in data collection---data that remains largely underutilized due to the difficulty of finding what exists \citep{wilkinson2016fair, tenopir2020data}.

However, the promise of data reuse remains largely unfulfilled due to a critical bottleneck: \textbf{dataset discovery}. Researchers seeking datasets for replication studies, cross-species comparison, or novel analysis face a search problem that differs fundamentally from traditional document retrieval. When a neuroscientist queries ``mouse Neuropixels decision-making data NOT visual cortex,'' they are not seeking textually similar documents---they require datasets that satisfy precise experimental constraints, support specific analysis methodologies, and provide sufficient metadata quality for scientific reuse.

\subsection{The Data Reuse Crisis in Neuroscience}

Despite substantial investments in open data infrastructure, studies consistently find low rates of dataset reuse. \citet{borgman2012conundrum} documented the ``conundrum'' of data sharing---while researchers support openness in principle, actual reuse remains rare. \citet{wallis2013if} found that even when data are available, discovering relevant datasets poses significant barriers. More recently, \citet{cousijn2018data} and \citet{tedersoo2021data} surveyed researchers across disciplines and identified dataset discovery as a primary obstacle to reuse.

The consequences extend beyond individual researchers. The BRAIN Initiative \citep{jorgenson2015brain} and Human Brain Project \citep{amunts2016human} invested billions in neuroscience data generation, yet cross-laboratory data integration remains challenging. The reproducibility crisis in psychology and neuroscience \citep{button2013power, open2015estimating} could be partly addressed through large-scale reanalysis---if researchers could find the relevant data.

\subsection{Why Existing Search Fails}

Current approaches to neuroscience dataset discovery fall into three categories, each with fundamental limitations:

\paragraph{Repository-Specific Search.} Each repository (DANDI, OpenNeuro, Allen) provides its own search interface. These interfaces typically support keyword matching and faceted filtering but lack cross-repository integration, semantic understanding, and awareness of analysis requirements. A researcher must independently search each repository and manually reconcile results.

\paragraph{General-Purpose Search Engines.} Google Dataset Search \citep{brickley2019google} and similar services index dataset metadata but treat datasets as documents rather than structured scientific objects. They cannot distinguish ``mouse decision-making'' from ``human decision-making'' except through keyword matching, and they have no concept of analysis affordances or experimental constraints.

\paragraph{Curated Catalogs.} Resources like NeuroVault \citep{gorgolewski2015neurovault} and BrainMap \citep{laird2005brainmap} provide curated collections for specific domains but require manual curation that cannot scale to the full breadth of neuroscience data. They also tend to focus on summary statistics rather than raw data suitable for novel analysis.

\subsection{The Scientific Dataset Search Problem}

Neural Search addresses these limitations by treating datasets as structured scientific objects rather than documents, enabling \textit{experiment-aware retrieval} that respects the unique requirements of scientific data discovery.

\begin{definition}[Scientific Dataset Search]
\label{def:dataset-search}
Given a query $\query$ expressing experimental intent through natural language, ontological constraints, and analysis requirements, retrieve a ranked list of datasets $\dataset_1, \dataset_2, \ldots, \dataset_k$ such that each $\dataset_i$:
\begin{enumerate}[label=(\roman*)]
    \item Satisfies all \textbf{hard constraints} (exclusions, required modalities, species restrictions)
    \item Maximizes alignment with \textbf{soft preferences} (task similarity, region overlap, behavioral paradigm)
    \item Provides sufficient \textbf{analysis affordances} for the intended scientific methodology
    \item Meets \textbf{metadata quality} thresholds for scientific reproducibility
\end{enumerate}
\end{definition}

This problem exhibits several characteristics that distinguish it from web search or enterprise document retrieval:

\paragraph{Structured Constraints.} Scientific queries implicitly specify multi-dimensional constraints. ``Mouse Neuropixels decision-making data'' simultaneously constrains organism (Mus musculus), recording modality (Neuropixels high-density extracellular electrophysiology), and behavioral task (decision-making paradigm family). These constraints exist in typed ontological spaces with hierarchical relationships (e.g., ``rodent'' subsumes ``mouse'').

\paragraph{Hard Negatives.} Exclusions like ``NOT fMRI'' or ``excluding human subjects'' represent inviolable constraints. A system that returns even one fMRI dataset in response to a query explicitly excluding imaging data has fundamentally failed, regardless of how relevant that dataset might be along other dimensions.

\paragraph{Analysis Affordances.} Relevance depends not only on what data a dataset contains, but whether it supports the researcher's intended analysis. A dataset with neural recordings but no trial-by-trial behavioral labels cannot support choice decoding analysis, regardless of how well its metadata matches the query text. We formalize this through \textit{analysis affordances}---predicates over dataset features that specify computational requirements.

\paragraph{Provenance Requirements.} Scientific reuse demands traceable evidence for metadata claims. A dataset labeled as ``decision-making'' should carry provenance indicating whether this label derived from structured metadata fields, paper abstracts, or filename heuristics---each with different confidence levels.

\subsection{Contributions}

We make the following contributions:

\begin{enumerate}
    \item \textbf{Formalization of experiment-aware retrieval} (Section~\ref{sec:formalization}): We introduce a mathematical framework for scientific dataset search that represents datasets as slot-based experimental context vectors and queries as constraint satisfaction problems over typed ontological spaces.

    \item \textbf{Provenance-weighted knowledge graph} (Section~\ref{sec:graph}): We develop a heterogeneous graph schema with confidence-weighted edges derived from multiple evidence sources, enabling metapath-based reasoning for discovering implicit dataset relationships.

    \item \textbf{Analysis affordance theory} (Section~\ref{sec:affordances}): We formalize the concept of analysis affordances as computational predicates over dataset features, enabling search that respects methodological requirements.

    \item \textbf{Hybrid retrieval architecture} (Section~\ref{sec:retrieval}): We present a multi-signal scoring function that combines BM25 lexical matching, dense semantic embeddings, ontology alignment, and graph-based features through reciprocal rank fusion.

    \item \textbf{Cross-disciplinary expansion roadmap} (Section~\ref{sec:expansion}): We outline a strategy for extending Neural Search to bridge neuroscience datasets with computational neuroscience models and artificial neural network research.

    \item \textbf{Empirical validation} (Section~\ref{sec:experiments}): We demonstrate the effectiveness of our approach on an initial expert-curated benchmark, achieving strong performance while maintaining zero hard-negative violations. We present a roadmap for expanded validation.
\end{enumerate}

\subsection{Claim Status and Evidence}

To ensure transparency, we classify each major claim by implementation and validation status:

\begin{table}[t]
\centering
\small
\caption{Claim status and supporting evidence. \textbf{Implemented}: code exists and runs. \textbf{Validated}: measured on benchmark. \textbf{Partial}: partially implemented. \textbf{Proposed}: future work.}
\label{tab:claim-status}
\begin{tabular}{p{4.5cm}lp{3.5cm}p{3cm}}
\toprule
\textbf{Claim} & \textbf{Status} & \textbf{Evidence} & \textbf{Remaining Work} \\
\midrule
Hybrid metadata retrieval improves search quality & Validated & 30-query benchmark, ablation study & Expand corpus and queries \\
Hard-negative filtering reduces invalid matches & Implemented & Zero violations on benchmark & Larger adversarial benchmark \\
Typed scientific labels improve interpretability & Implemented & Manual inspection & Ablation validation needed \\
Graph metapaths improve cross-dataset relatedness & Partial & Graph infrastructure exists & Pairing benchmark needed \\
Analysis affordance search & Implemented & 20+ affordance detectors with hard-negative checks & Affordance validation rubric \\
Query sense disambiguation & Validated & 38 tests for delay/motor/signal sense separation & Expand sense vocabulary \\
Provenance-backed claims schema & Implemented & ReusabilityClaim with source confidence & Claim extraction pipeline \\
Evidence-backed result cards & Implemented & EvidenceResultCard with claim citations & ClaimStore integration \\
Reusability benchmark & Implemented & 30-query gold benchmark & Multi-annotator validation \\
Latent neural signature search & Proposed & Schema defined & NWB feature extraction \\
Cross-species alignment & Proposed & Ontology structure exists & Benchmark and validation \\
Causal claim graph & Proposed & Not implemented & Full schema and extraction \\
\bottomrule
\end{tabular}
\end{table}

% ============================================================================
% SECTION 2: RELATED WORK
% ============================================================================
\section{Related Work}
\label{sec:related}

Neural Search builds upon and extends work in information retrieval, scientific knowledge graphs, biomedical dataset discovery, representation learning, and the emerging field of scientific data management. We provide a comprehensive review of each area and position our contributions within this landscape.

\subsection{Information Retrieval Foundations}

\paragraph{Classical Lexical Retrieval.} The foundations of modern information retrieval rest on probabilistic relevance models. The vector space model \citep{salton1975vector} introduced geometric representations of documents. The probabilistic relevance framework, culminating in BM25 \citep{robertson2009probabilistic}, remains a strong baseline for text retrieval. Extensions to structured documents include field-weighted BM25 \citep{robertson2004simple}, which allows different weights for different document fields---a capability essential for scientific metadata where title, description, and structured labels carry different semantic loads.

\paragraph{Learning to Rank.} Machine learning approaches to ranking \citep{liu2009learning} learn optimal combinations of retrieval signals. LambdaMART \citep{burges2010ranknet} and neural ranking models \citep{guo2016deep} demonstrated substantial improvements over hand-tuned systems. Neural Search adopts a simpler fusion approach (RRF) that provides transparency and interpretability important for scientific applications, though learning-to-rank remains a direction for future work.

\paragraph{Dense Neural Retrieval.} The advent of transformer-based language models revolutionized retrieval. BERT \citep{devlin2019bert} enabled rich contextual representations. Dense Passage Retrieval (DPR) \citep{karpukhin2020dense} demonstrated that bi-encoder architectures could outperform BM25 on open-domain question answering. Subsequent work improved training through hard negative mining \citep{xiong2020approximate}, knowledge distillation \citep{hofstatter2020improving}, and contrastive learning \citep{gao2021simcse}. ColBERT \citep{khattab2020colbert} introduced late interaction for improved token-level matching. We employ dense retrieval for capturing semantic similarity that lexical matching misses (e.g., ``ephys'' $\approx$ ``electrophysiology'').

\paragraph{Sparse Learned Representations.} SPLADE \citep{formal2021splade} and related approaches learn sparse lexical expansions that combine the interpretability of bag-of-words models with learned semantic understanding. DeepImpact \citep{mallia2021learning} and uniCOIL \citep{lin2021few} provide efficient sparse representations suitable for inverted index retrieval. These approaches inform our ontology-aware expansion strategy.

\paragraph{Hybrid Retrieval Systems.} Reciprocal Rank Fusion (RRF) \citep{cormack2009reciprocal} provides a simple yet effective unsupervised method for combining multiple rankers without requiring labeled training data. Recent work explores learned fusion \citep{chen2022hybrid}, multi-stage retrieval pipelines \citep{nogueira2019passage}, and retrieval-augmented generation \citep{lewis2020retrieval}. Neural Search extends RRF to incorporate domain-specific signals (ontology matching, graph traversal, affordance scoring) alongside standard lexical and semantic retrievers.

\paragraph{Constraint-Aware Retrieval.} Traditional retrieval optimizes for relevance without hard constraints. Faceted search \citep{hearst2006design} allows users to filter results but treats all constraints as post-retrieval filters. Our approach distinguishes hard constraints (must not return human data when excluded) from soft preferences (prefer mouse over rat), with hard constraints applied during candidate generation to ensure zero violations.

\subsection{Scientific Knowledge Graphs and Ontologies}

\paragraph{Biomedical Knowledge Bases.} The life sciences have a rich tradition of structured knowledge representation. PubMed \citep{sayers2019pubmed} indexes over 35 million biomedical publications with MeSH term annotations. UniProt \citep{uniprot2021uniprot} provides comprehensive protein sequence and functional information. The Gene Ontology \citep{gene2021gene} defines a controlled vocabulary for gene function across species. These resources enable precise semantic queries but focus on molecular biology rather than systems neuroscience.

\paragraph{Biomedical Semantic Resources.} UMLS (Unified Medical Language System) \citep{bodenreider2004unified} integrates over 200 biomedical vocabularies. Biolink Model \citep{biolink2020model} provides a high-level schema for biological knowledge graphs. SemMedDB \citep{kilicoglu2012semmeddb} extracts semantic predications from biomedical literature. We adapt these approaches to neuroscience-specific ontologies covering experimental tasks, recording modalities, and analysis methods.

\paragraph{Heterogeneous Information Networks.} Sun et al. \citep{sun2011pathsim} introduced PathSim for metapath-based similarity in heterogeneous graphs, demonstrating that type-constrained paths capture domain-relevant relationships. Subsequent work developed metapath2vec \citep{dong2017metapath2vec} for representation learning, HAN (Heterogeneous Graph Attention Networks) \citep{wang2019heterogeneous} for attention-based reasoning, and HGT (Heterogeneous Graph Transformer) \citep{hu2020heterogeneous} for large-scale heterogeneous graphs. We employ metapath templates for discovering scientific relationships (e.g., Dataset$\rightarrow$Task$\leftarrow$Dataset for same-task retrieval).

\paragraph{Knowledge Graph Embeddings.} TransE \citep{bordes2013translating} and its successors (TransR, RotatE, ComplEx) learn distributed representations of entities and relations. These embeddings enable link prediction and entity alignment. Knowledge graph completion \citep{wang2017knowledge} predicts missing edges, potentially useful for inferring dataset-to-task relationships from incomplete metadata.

\paragraph{Scientific Knowledge Graphs.} Recent work has developed knowledge graphs specifically for scientific data. Microsoft Academic Graph \citep{sinha2015overview} and OpenAlex \citep{priem2022openalex} provide large-scale scholarly knowledge graphs. Semantic Scholar \citep{ammar2018construction} extracts structured information from scientific papers. We integrate publication metadata from these sources to enrich dataset provenance.

\subsection{Dataset Discovery and Scientific Data Management}

\paragraph{Dataset Search Engines.} Google Dataset Search \citep{brickley2019google} indexes datasets using Schema.org markup. DataCite \citep{datacite2021metadata} provides persistent identifiers and metadata for research data. These systems treat datasets as documents with metadata rather than as structured scientific objects with affordances.

\paragraph{Domain-Specific Metadata Standards.} BIDS (Brain Imaging Data Structure) \citep{gorgolewski2016brain} standardizes neuroimaging dataset organization. NWB (Neurodata Without Borders) \citep{teeters2015neurodata, rubel2022nwb} provides a unified format for neurophysiology data. These standards enable programmatic access to structured metadata but require repository-specific interfaces for discovery.

\paragraph{Provenance and Research Object Standards.} PROV-O \citep{lebo2013prov} provides a W3C standard for representing provenance. RO-Crate \citep{soiland2022packaging} packages research objects with rich metadata for reproducibility. LinkML \citep{moxon2021linkml} enables schema definition for linked data. Bioschemas \citep{gray2017bioschemas} extends Schema.org for life sciences. We align our provenance model with these standards, tracking the evidence sources and confidence levels for each metadata claim.

\paragraph{Neuroscience Data Infrastructure.} EBRAINS Knowledge Graph \citep{amunts2019ebrains} and openMINDS metadata framework provide structured neuroscience data organization at European scale. The BRAIN Initiative Cell Census Network (BICCN) \citep{ecker2017brain} coordinates large-scale brain mapping data. While these systems focus on data organization and standardization, Neural Search complements them by providing experiment-aware retrieval over heterogeneous sources.

\paragraph{FAIR Data Principles.} The FAIR principles \citep{wilkinson2016fair} (Findable, Accessible, Interoperable, Reusable) provide guidelines for scientific data management. Neural Search directly addresses the ``Findable'' dimension through improved discovery, while leveraging the metadata standards that support interoperability.

\subsection{Neuroscience Data Repositories}

\paragraph{Neurophysiology Archives.} DANDI Archive \citep{rubinen2024dandi} standardizes neurophysiology data sharing with NWB format requirements. The International Brain Laboratory \citep{ibl2021standardized} provides standardized decision-making datasets across laboratories. Allen Brain Observatory \citep{allen2019brain} offers large-scale visual coding datasets with consistent experimental protocols.

\paragraph{Neuroimaging Repositories.} OpenNeuro \citep{markiewicz2021openneuro} hosts BIDS-formatted MRI, MEG, EEG, and PET data. NeuroVault \citep{gorgolewski2015neurovault} stores statistical maps from neuroimaging studies. BrainMap \citep{laird2005brainmap} provides coordinate-based meta-analysis infrastructure. Human Connectome Project \citep{van2013wu} released high-quality structural and functional connectivity data.

\paragraph{Specialized Archives.} NeMO Archive \citep{nemo2023archive} focuses on transcriptomics and single-cell data. CRCNS (Collaborative Research in Computational Neuroscience) \citep{teeters2008crcns} provides curated datasets for computational neuroscience. ModelDB \citep{hines2004modeldb} hosts computational neuroscience models. These repositories have distinct metadata schemas and search interfaces, motivating our unified discovery layer.

\paragraph{Data Reuse Challenges.} Studies document significant barriers to neuroscience data reuse including inconsistent metadata \citep{gleeson2019systematic}, missing analysis documentation \citep{soska2021data}, and difficulty finding relevant datasets \citep{ferguson2014big}. \citet{tenopir2020data} found that while data sharing has increased, actual reuse remains limited. Neural Search directly addresses the discovery challenge that underlies many reuse barriers.

\subsection{Bridging Neuroscience and Artificial Intelligence}

\paragraph{Computational Neuroscience Models.} Models of neural computation span scales from biophysical (Hodgkin-Huxley \citep{hodgkin1952quantitative}) to population-level (rate models, neural mass models) to cognitive (drift-diffusion models, reinforcement learning). Simulation environments like NEURON \citep{hines1997neuron}, NEST \citep{gewaltig2007nest}, and Brian \citep{stimberg2019brian} require specific input formats. Neural Search can identify datasets suitable for model fitting and validation.

\paragraph{Neural Network Analysis of Neural Data.} Modern machine learning provides powerful tools for neural data analysis. Dimensionality reduction \citep{cunningham2014dimensionality}, neural network decoders \citep{glaser2020machine}, and latent variable models \citep{pandarinath2018inferring} extract structure from high-dimensional neural recordings. These methods have specific data requirements (trial structure, sufficient samples) that our affordance framework captures.

\paragraph{Neuro-AI Convergence.} Deep learning increasingly draws inspiration from neuroscience \citep{hassabis2017neuroscience, richards2019deep}, while computational models inform neural data interpretation \citep{saxe2021if}. Brain-Score \citep{schrimpf2018brain} benchmarks deep networks against neural data. The NeuroAI initiative \citep{zador2023catalyzing} calls for tighter integration. Our expansion roadmap positions Neural Search to bridge these communities by enabling discovery across experimental neuroscience, computational modeling, and AI research.

\paragraph{Foundation Models for Neuroscience.} Recent work has developed foundation models for neural data, including LFADS \citep{pandarinath2018inferring} for neural dynamics, neural population embeddings \citep{ye2023neural}, and multimodal models integrating neural and behavioral data. These approaches could eventually enable content-based search over neural population activity---a direction we discuss in our latent neural signature roadmap.

% ============================================================================
% SECTION 3: FORMALIZATION
% ============================================================================
\section{Experiment-Aware Retrieval Formalization}
\label{sec:formalization}

\subsection{Dataset as Experimental Context}

We represent each dataset as a structured experimental context rather than an unstructured document.

\begin{definition}[Experimental Slot Vector]
\label{def:slot-vector}
A dataset $d$ is represented as a typed slot vector $E(d) \in \mathcal{S}$ where $\mathcal{S}$ is the slot space:
\begin{equation}
E(d) = \begin{pmatrix}
    \text{organism}(d) \in \mathcal{O}_{\text{taxon}} \\
    \text{taxon\_group}(d) \in \mathcal{O}_{\text{taxon\_group}} \\
    \text{preparation}(d) \in \mathcal{O}_{\text{prep}} \\
    \text{task\_family}(d) \in \mathcal{O}_{\text{task}} \\
    \text{task\_variant}(d) \in \mathcal{O}_{\text{task\_variant}} \\
    \text{modality\_family}(d) \in \mathcal{O}_{\text{mod}} \\
    \text{device}(d) \in \mathcal{O}_{\text{device}} \\
    \text{signal\_type}(d) \in \mathcal{O}_{\text{signal}} \\
    \text{brain\_region}(d) \in \mathcal{P}(\mathcal{O}_{\text{region}}) \\
    \text{behavioral\_events}(d) \in \mathcal{P}(\mathcal{O}_{\text{event}}) \\
    \text{data\_standard}(d) \in \mathcal{O}_{\text{standard}} \\
    \text{analysis\_affordances}(d) \in \mathcal{P}(\mathcal{A}) \\
    \text{provenance\_confidence}(d) \in [0, 1]
\end{pmatrix}
\end{equation}
where $\mathcal{O}_*$ denotes domain-specific ontologies and $\mathcal{P}(\cdot)$ denotes power sets for multi-valued slots.
\end{definition}

Each slot maps to a typed ontological space with hierarchical relationships:

\begin{example}[Ontology Hierarchies]
\begin{align*}
    \text{taxon}: & \quad \text{Vertebrata} \supset \text{Mammalia} \supset \text{Rodentia} \supset \text{Mus musculus} \\
    \text{modality}: & \quad \text{Electrophysiology} \supset \text{Extracellular} \supset \text{Neuropixels} \\
    \text{task}: & \quad \text{Decision-making} \supset \text{2AFC} \supset \text{IBL\_task}
\end{align*}
\end{example}

\subsection{Query as Constraint Satisfaction}

A query specifies constraints over the slot space:

\begin{definition}[Query Constraint Set]
\label{def:query-constraints}
A query $\query$ is a tuple $(\mathcal{C}_{\text{hard}}, \mathcal{C}_{\text{soft}}, \mathcal{A}_{\text{req}}, \theta)$ where:
\begin{itemize}
    \item $\mathcal{C}_{\text{hard}}$: Hard constraints that must be satisfied (exclusions, requirements)
    \item $\mathcal{C}_{\text{soft}}$: Soft preferences that influence ranking
    \item $\mathcal{A}_{\text{req}}$: Required analysis affordances
    \item $\theta$: Intent classification from $\mathcal{I} = \{\text{lookup}, \text{task}, \text{modality}, \text{affordance}, \text{similarity}, \text{exploration}\}$
\end{itemize}
\end{definition}

\begin{definition}[Hard Constraint Satisfaction]
A dataset $d$ satisfies hard constraints $\mathcal{C}_{\text{hard}}$ iff:
\begin{equation}
    \text{satisfies}(d, \mathcal{C}_{\text{hard}}) = \bigwedge_{c \in \mathcal{C}_{\text{hard}}} \text{eval}(c, E(d))
\end{equation}
where $\text{eval}(c, E(d))$ evaluates constraint $c$ against the slot vector.
\end{definition}

Constraint types include:
\begin{itemize}
    \item \textbf{Inclusion}: $\text{species}(d) \in \{\text{mouse}, \text{rat}\}$
    \item \textbf{Exclusion}: $\text{modality}(d) \notin \{\text{fMRI}, \text{EEG}\}$
    \item \textbf{Subsumption}: $\text{task}(d) \sqsubseteq \text{Decision-making}$
    \item \textbf{Disjunction}: $\text{region}(d) \cap \{\text{PFC}, \text{striatum}\} \neq \emptyset$
\end{itemize}

\subsection{Relevance Scoring}

Given constraint satisfaction, relevance scoring combines multiple signals:

\begin{theorem}[Hybrid Relevance Score]
\label{thm:hybrid-score}
For dataset $d$ and query $\query$, the relevance score is:
\begin{equation}
\begin{split}
\score(d, \query) = &\text{clamp}\Bigg(
    \sum_{s \in \text{signals}} w_s(\theta) \cdot S_s(d, \query) \\
    &- \sum_{p \in \text{penalties}} \lambda_p \cdot P_p(d, \query), \quad 0, 1\Bigg)
\end{split}
\end{equation}
where signals include ontology matching, semantic similarity, field embeddings, graph features, and readiness scores; penalties include mismatch, missing metadata, and affordance gaps; and $w_s(\theta)$ are intent-dependent weights.
\end{theorem}

% ============================================================================
% SECTION 4: KNOWLEDGE GRAPH
% ============================================================================
\section{Provenance-Weighted Knowledge Graph}
\label{sec:graph}

\subsection{Heterogeneous Graph Schema}

\begin{definition}[Scientific Knowledge Graph]
\label{def:kg}
The Neural Search knowledge graph is a tuple $\graph = (\nodes, \edges, \tau_V, \tau_E, \confidence, \evidence)$ where:
\begin{itemize}
    \item $\nodes$ is the set of typed nodes
    \item $\edges \subseteq \nodes \times \nodes$ is the set of directed edges
    \item $\tau_V: \nodes \rightarrow \mathcal{T}_V$ assigns node types
    \item $\tau_E: \edges \rightarrow \mathcal{T}_E$ assigns edge types
    \item $\confidence: \edges \rightarrow [0, 1]$ assigns confidence scores
    \item $\evidence: \edges \rightarrow \mathcal{P}(\text{Evidence})$ maps edges to provenance
\end{itemize}
\end{definition}

The type system comprises:

\begin{align*}
\mathcal{T}_V = \{&\texttt{dataset}, \texttt{paper}, \texttt{task}, \texttt{modality}, \texttt{species}, \\
&\texttt{region}, \texttt{event}, \texttt{analysis}, \texttt{standard}, \texttt{requirement}\}
\end{align*}

\begin{align*}
\mathcal{T}_E = \{&\texttt{has\_task}, \texttt{has\_modality}, \texttt{records\_region}, \texttt{has\_species}, \\
&\texttt{has\_event}, \texttt{supports\_analysis}, \texttt{requires}, \texttt{linked\_to\_paper}, \\
&\texttt{similar\_to}, \texttt{derived\_from}, \texttt{cites}\}
\end{align*}

\begin{figure}[t]
\centering
\begin{tikzpicture}[
    node distance=1.8cm,
    every node/.style={font=\small},
    dataset/.style={rectangle, draw=datacolor, fill=datacolor!20, rounded corners, minimum width=2cm, minimum height=0.7cm},
    concept/.style={ellipse, draw=graphcolor, fill=graphcolor!20, minimum width=1.5cm, minimum height=0.6cm},
    paper/.style={rectangle, draw=evalcolor, fill=evalcolor!20, rounded corners, minimum width=1.5cm, minimum height=0.6cm},
    edge/.style={-{Stealth[length=2mm]}, thick},
    label/.style={font=\scriptsize, midway, fill=white, inner sep=1pt}
]
    % Nodes
    \node[dataset] (d1) {Dataset 1};
    \node[dataset, right=3cm of d1] (d2) {Dataset 2};

    \node[concept, below=1.2cm of d1] (task) {Decision-making};
    \node[concept, below=1.2cm of d2] (mod) {Neuropixels};
    \node[concept, below left=0.8cm and 0.5cm of task] (species) {Mouse};
    \node[concept, below right=0.8cm and 0.5cm of mod] (region) {PFC};

    \node[paper, below=2.5cm of $(d1)!0.5!(d2)$] (paper) {Paper};

    \node[concept, right=1.5cm of region] (analysis) {Choice Decoding};

    % Edges
    \draw[edge, datacolor] (d1) -- node[label, above left] {has\_task} (task);
    \draw[edge, datacolor] (d2) -- node[label, above right] {has\_modality} (mod);
    \draw[edge, datacolor] (d1) -- node[label, left] {has\_species} (species);
    \draw[edge, datacolor] (d2) -- node[label, right] {records\_region} (region);
    \draw[edge, graphcolor] (task) -- node[label, below] {shared\_task} (d2);
    \draw[edge, evalcolor] (d1) -- node[label, left] {linked\_to} (paper);
    \draw[edge, evalcolor] (d2) -- node[label, right] {linked\_to} (paper);
    \draw[edge, graphcolor] (mod) -- node[label, right] {supports} (analysis);
    \draw[edge, graphcolor!60, dashed] (d1) -- node[label, above] {similar\_to} (d2);
\end{tikzpicture}
\caption{Knowledge graph schema showing typed nodes (datasets, concepts, papers) and relationship edges with provenance-weighted confidence.}
\label{fig:kg-schema}
\end{figure}

\begin{table}[t]
\centering
\small
\caption{Compact graph schema: node types, edge types, and confidence model.}
\label{tab:graph-schema}
\begin{tabular}{llp{5cm}}
\toprule
\textbf{Node Type} & \textbf{Examples} & \textbf{Key Properties} \\
\midrule
Dataset & DANDI dandiset, OpenNeuro dataset & source, species, modalities, tasks \\
Paper & DOI, PubMed record & title, authors, abstract, year \\
Task & Go/NoGo, reversal learning & family, ontology path \\
Modality & Neuropixels, calcium imaging, fMRI & signal type, device \\
Species & mouse, macaque, human & taxon ID, common name \\
Region & V1, PFC, hippocampus & hierarchy path, Allen ID \\
Analysis & choice decoding, Q-learning & required fields, optional fields \\
\bottomrule
\end{tabular}
\end{table}

\begin{table}[t]
\centering
\small
\caption{Edge types with semantic meaning and confidence sources.}
\label{tab:edge-types}
\begin{tabular}{lp{4.5cm}l}
\toprule
\textbf{Edge Type} & \textbf{Meaning} & \textbf{Confidence Source} \\
\midrule
\texttt{has\_modality} & dataset contains modality & schema field, extraction \\
\texttt{records\_region} & dataset targets brain region & schema field, paper \\
\texttt{uses\_task} & dataset uses behavioral task & schema field, paper \\
\texttt{has\_event} & dataset contains event labels & NWB/BIDS inspection \\
\texttt{supports\_analysis} & dataset satisfies affordance & rule-based detection \\
\texttt{linked\_to\_paper} & metadata from publication & DOI match, citation \\
\texttt{similar\_to} & task-level similarity & metapath count \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Provenance-Weighted Confidence}

Edge confidence aggregates evidence from multiple sources:

\begin{definition}[Evidence-Based Confidence]
\label{def:confidence}
For edge $e$ with evidence set $\evidence(e)$, the confidence is:
\begin{equation}
\confidence(e) = 1 - \prod_{ev \in \evidence(e)} (1 - w_{\text{source}}(ev) \cdot w_{\text{method}}(ev) \cdot w_{\text{recency}}(ev))
\end{equation}
\end{definition}

Source weights follow a hierarchy:
\begin{equation}
w_{\text{source}}(ev) = \begin{cases}
    0.95 & \text{structured metadata field} \\
    0.85 & \text{curated ontology match} \\
    0.75 & \text{paper abstract extraction} \\
    0.60 & \text{description inference} \\
    0.40 & \text{filename/path heuristic}
\end{cases}
\end{equation}

\subsection{Metapath-Based Reasoning}

Metapaths define semantically meaningful traversal patterns \citep{sun2011pathsim}:

\begin{definition}[Metapath Template]
A metapath template $\mathcal{P}$ is a sequence:
\begin{equation}
\mathcal{P} = T_1 \xrightarrow{R_1} T_2 \xrightarrow{R_2} \cdots \xrightarrow{R_{l-1}} T_l
\end{equation}
\end{definition}

We define domain-specific metapath templates:

\begin{example}[Scientific Metapaths]
\begin{align*}
\mathcal{P}_{\text{analysis}} &= \texttt{Dataset} \xrightarrow{\texttt{has\_event}} \texttt{Event} \xrightarrow{\texttt{required\_for}} \texttt{Analysis} \\
\mathcal{P}_{\text{paper}} &= \texttt{Dataset} \xrightarrow{\texttt{linked\_to}} \texttt{Paper} \xrightarrow{\texttt{studies}} \texttt{Task} \\
\mathcal{P}_{\text{similar}} &= \texttt{Dataset} \xrightarrow{\texttt{has\_task}} \texttt{Task} \xleftarrow{\texttt{has\_task}} \texttt{Dataset} \\
\mathcal{P}_{\text{cross-species}} &= \texttt{Dataset}_1 \xrightarrow{\texttt{has\_region}} \texttt{Region} \xleftarrow{\texttt{has\_region}} \texttt{Dataset}_2
\end{align*}
\end{example}

Path-based similarity uses PathSim \citep{sun2011pathsim}:
\begin{equation}
\text{PathSim}(v_i, v_j | \mathcal{P}) = \frac{2 \cdot |\text{paths}_\mathcal{P}(v_i, v_j)|}{\sum_{k \in \{i,j\}} |\text{paths}_\mathcal{P}(v_k, v_k)|}
\end{equation}

% ============================================================================
% SECTION 5: ANALYSIS AFFORDANCES
% ============================================================================
\section{Analysis Affordance Theory}
\label{sec:affordances}

A key insight of Neural Search is that dataset relevance depends not only on experimental context but on \textit{what analyses the data can support}. We formalize this through analysis affordances.

\subsection{Affordance as Constraint Satisfaction}

\begin{definition}[Analysis Affordance]
\label{def:affordance}
An analysis affordance $A$ is a predicate over dataset features:
\begin{equation}
\text{affordance}_A(d) = \bigwedge_{r \in \text{requirements}(A)} \text{satisfies}(d, r)
\end{equation}
\end{definition}

Requirements are typed predicates over dataset features:

\begin{example}[Choice Decoding Affordance]
\begin{equation}
\begin{split}
\text{affordance}_{\text{choice\_decoding}}(d) = &\text{has\_neural\_recording}(d) \\
\land &\text{has\_trial\_structure}(d) \\
\land &\text{has\_choice\_labels}(d) \\
\land &\text{has\_sufficient\_trials}(d, n \geq 100)
\end{split}
\end{equation}
\end{example}

\begin{example}[Q-Learning Modeling Affordance]
\begin{equation}
\begin{split}
\text{affordance}_{\text{Q-learning}}(d) = &\text{has\_choice\_sequence}(d) \\
\land &\text{has\_reward\_feedback}(d) \\
\land &\text{has\_trial\_outcome}(d) \\
\land &\text{has\_stimulus\_features}(d)
\end{split}
\end{equation}
\end{example}

\subsection{Affordance Graph Representation}

Analysis affordances are represented in the knowledge graph through requirement hyperedges:

\begin{definition}[Requirement Hyperedge]
A hyperedge $h = (A, \{r_1, \ldots, r_n\})$ encodes that analysis $A$ requires all features $r_1, \ldots, r_n$ simultaneously.
\end{definition}

\begin{figure}[t]
\centering
\begin{tikzpicture}[
    node distance=1.5cm,
    analysis/.style={rectangle, draw=searchcolor, fill=searchcolor!20, rounded corners, minimum width=2.5cm, minimum height=0.7cm, font=\small},
    req/.style={ellipse, draw=graphcolor, fill=graphcolor!20, minimum width=2cm, minimum height=0.5cm, font=\scriptsize},
    edge/.style={-{Stealth[length=2mm]}, thick, graphcolor}
]
    \node[analysis] (choice) {Choice Decoding};
    \node[req, below left=1cm and 0.3cm of choice] (neural) {Neural Recording};
    \node[req, below=1cm of choice] (trials) {Trial Structure};
    \node[req, below right=1cm and 0.3cm of choice] (labels) {Choice Labels};

    \draw[edge] (choice) -- (neural);
    \draw[edge] (choice) -- (trials);
    \draw[edge] (choice) -- (labels);

    \node[analysis, right=3cm of choice] (qlearn) {Q-Learning Model};
    \node[req, below left=1cm and 0.3cm of qlearn] (seq) {Choice Sequence};
    \node[req, below=1cm of qlearn] (reward) {Reward Signal};
    \node[req, below right=1cm and 0.3cm of qlearn] (outcome) {Trial Outcome};

    \draw[edge] (qlearn) -- (seq);
    \draw[edge] (qlearn) -- (reward);
    \draw[edge] (qlearn) -- (outcome);
\end{tikzpicture}
\caption{Analysis affordance requirement graphs. Each analysis requires a conjunction of features that datasets must satisfy.}
\label{fig:affordance-graph}
\end{figure}

\begin{table}[t]
\centering
\small
\caption{Analysis affordance requirements for key neuroscience methodologies.}
\label{tab:affordance-requirements}
\begin{tabular}{lp{4cm}p{4cm}}
\toprule
\textbf{Affordance} & \textbf{Required Fields} & \textbf{Optional Fields} \\
\midrule
Choice decoding & neural activity, trial labels, choice labels & reaction time, stimulus type \\
Q-learning & choices, rewards, trial order, subject/session IDs & block structure, task state \\
State-space modeling & continuous time series, timestamps, sampling rate & trial boundaries, behavior labels \\
Cross-session analysis & subject IDs, session dates, unit metadata & drift metrics, coordinates \\
Causal perturbation & intervention labels, timing, control condition & dose, stim parameters \\
RSA & aligned stimuli, population responses & model features, train/test split \\
Neural-behavior alignment & neural time series, behavior time series, timestamps & event labels, pose coordinates \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Affordance Confidence Propagation}

Affordance confidence propagates from requirement evidence:
\begin{equation}
\confidence_A(d) = \min_{r \in \text{requirements}(A)} \confidence(\text{evidence}(d, r))
\end{equation}

We use the minimum rather than product to reflect that any unsatisfied requirement blocks the analysis.

% ============================================================================
% SECTION 6: HYBRID RETRIEVAL
% ============================================================================
\section{Hybrid Retrieval Architecture}
\label{sec:retrieval}

\subsection{Multi-Signal Scoring}

Neural Search combines heterogeneous retrieval signals through a modular scoring architecture.

\subsubsection{Lexical Retrieval (BM25)}

Field-weighted BM25 \citep{robertson2004simple} scores structured metadata:
\begin{equation}
S_{\text{BM25}}(d, \query) = \sum_{f \in \mathcal{F}} w_f \cdot \sum_{t \in \query} \text{IDF}(t) \cdot \frac{f_{t,d_f} \cdot (k_1 + 1)}{f_{t,d_f} + k_1 \cdot (1 - b + b \cdot \frac{|d_f|}{\text{avgdl}_f})}
\end{equation}

Fields $\mathcal{F}$ include: \texttt{title}, \texttt{description}, \texttt{tasks}, \texttt{modalities}, \texttt{species}, \texttt{brain\_regions}, \texttt{behavioral\_events}, and \texttt{data\_standards}.

\subsubsection{Dense Semantic Retrieval}

Bi-encoder embeddings capture semantic similarity:
\begin{equation}
S_{\text{semantic}}(d, \query) = \frac{\embed(\query) \cdot \embed(d)}{\|\embed(\query)\| \cdot \|\embed(d)\|}
\end{equation}

We use domain-adapted sentence transformers \citep{reimers2019sentence} fine-tuned on neuroscience corpora.

\subsubsection{Field-Specific Embeddings}

Multi-aspect semantic fingerprints enable dimension-specific similarity:
\begin{equation}
\fingerprint(d) = [\embed_{\text{task}}(d) \| \embed_{\text{mod}}(d) \| \embed_{\text{beh}}(d) \| \embed_{\text{region}}(d)]
\end{equation}

\begin{equation}
S_{\text{field}}(d, \query) = \sum_{a \in \text{aspects}} w_a \cdot \cos(\embed_a(\query), \embed_a(d))
\end{equation}

\subsubsection{Ontology Matching}

Set-based matching over ontological labels:
\begin{equation}
S_{\text{onto}}(d, \query) = \frac{|L_d \cap L_\query|}{|L_\query|} \cdot \frac{1}{1 + \beta \cdot |L_d \setminus L_\query|}
\end{equation}

With hierarchical expansion using subsumption:
\begin{equation}
S_{\text{hier}}(d, \query) = \frac{\sum_{l \in L_\query} \max_{l' \in L_d} \text{sim}_{\text{onto}}(l, l')}{|L_\query|}
\end{equation}

\subsubsection{Graph-Based Features}

Graph signals include metapath similarity and centrality:
\begin{equation}
S_{\text{graph}}(d, \query) = \sum_{\mathcal{P}} w_\mathcal{P} \cdot \text{PathSim}(d, \query | \mathcal{P}) + \alpha \cdot \text{PageRank}(d)
\end{equation}

\subsection{Reciprocal Rank Fusion}

Individual rankers are combined via RRF \citep{cormack2009reciprocal}:
\begin{equation}
\text{RRF}(d) = \sum_{i=1}^m \frac{w_i}{k + \text{rank}_i(d)}
\end{equation}

where $k=60$ is the standard constant and $w_i$ are ranker weights.

\begin{figure}[t]
\centering
\begin{tikzpicture}[
    node distance=1.2cm,
    box/.style={rectangle, draw, rounded corners, minimum width=2.2cm, minimum height=0.6cm, font=\small},
    signal/.style={box, fill=datacolor!20, draw=datacolor},
    fusion/.style={box, fill=searchcolor!20, draw=searchcolor},
    output/.style={box, fill=evalcolor!20, draw=evalcolor},
    arrow/.style={-{Stealth[length=2mm]}, thick}
]
    % Input
    \node[box, fill=gray!20] (query) {Query};

    % Signals
    \node[signal, below left=1.5cm and 1cm of query] (bm25) {BM25};
    \node[signal, below=1.5cm of query] (dense) {Dense};
    \node[signal, below right=1.5cm and 1cm of query] (graph) {Graph};

    % Fusion
    \node[fusion, below=1.5cm of dense] (rrf) {RRF Fusion};

    % Rerank
    \node[output, below=1cm of rrf] (rank) {Ranked Results};

    % Arrows
    \draw[arrow] (query) -- (bm25);
    \draw[arrow] (query) -- (dense);
    \draw[arrow] (query) -- (graph);
    \draw[arrow] (bm25) -- (rrf);
    \draw[arrow] (dense) -- (rrf);
    \draw[arrow] (graph) -- (rrf);
    \draw[arrow] (rrf) -- (rank);
\end{tikzpicture}
\caption{Hybrid retrieval architecture with BM25 lexical, dense semantic, and graph-based signals combined via reciprocal rank fusion.}
\label{fig:retrieval}
\end{figure}

\subsection{Intent-Adaptive Weighting}

Query intent determines signal weights:

\begin{table}[t]
\centering
\small
\caption{Intent-specific weight profiles for scoring signals.}
\begin{tabular}{lcccccc}
\toprule
Signal & Lookup & Task & Modality & Affordance & Similar & Explore \\
\midrule
$w_{\text{BM25}}$ & \textbf{0.35} & 0.20 & 0.20 & 0.15 & 0.15 & 0.20 \\
$w_{\text{semantic}}$ & 0.15 & 0.15 & 0.15 & 0.15 & \textbf{0.30} & 0.25 \\
$w_{\text{onto}}$ & 0.20 & \textbf{0.30} & \textbf{0.30} & 0.20 & 0.20 & 0.15 \\
$w_{\text{graph}}$ & 0.10 & 0.15 & 0.15 & 0.20 & 0.20 & \textbf{0.25} \\
$w_{\text{afford}}$ & 0.10 & 0.10 & 0.10 & \textbf{0.25} & 0.10 & 0.10 \\
$w_{\text{ready}}$ & 0.10 & 0.10 & 0.10 & 0.05 & 0.05 & 0.05 \\
\bottomrule
\end{tabular}
\label{tab:intent-weights}
\end{table}

\subsection{Computational Complexity Analysis}

Understanding the computational costs of each retrieval component informs system design and scaling decisions.

\begin{theorem}[Query Complexity]
\label{thm:complexity}
For a corpus of $n$ datasets, knowledge graph with $|V|$ nodes and $|E|$ edges, embedding dimension $d$, and query with $m$ terms, the query-time complexity is:
\begin{equation}
O(m \cdot \log n + k \cdot d + |E|/|V| \cdot k + k \cdot |\mathcal{A}|)
\end{equation}
where $k$ is the candidate set size and $|\mathcal{A}|$ is the number of affordance types.
\end{theorem}

\begin{proof}
The complexity derives from each component:
\begin{itemize}
    \item \textbf{BM25 retrieval}: $O(m \cdot \log n)$ using inverted index with term frequency lists
    \item \textbf{Dense retrieval}: $O(d \cdot \log n)$ using HNSW approximate nearest neighbor
    \item \textbf{Ontology matching}: $O(k \cdot |L_q|)$ where $|L_q|$ is query label count
    \item \textbf{Graph traversal}: $O(k \cdot \bar{d})$ where $\bar{d} = |E|/|V|$ is mean node degree
    \item \textbf{Affordance scoring}: $O(k \cdot |\mathcal{A}|)$ checking affordance predicates
    \item \textbf{RRF fusion}: $O(k \cdot \log k)$ for sorting and merging
\end{itemize}
The dominant term depends on corpus size and graph density.
\end{proof}

\paragraph{Scaling Properties.} The architecture scales favorably:
\begin{itemize}
    \item \textbf{Corpus size}: Sublinear in $n$ due to inverted indices and ANN structures
    \item \textbf{Query length}: Linear in $m$, typically small ($m < 20$)
    \item \textbf{Graph size}: Linear in mean degree $\bar{d}$, bounded by schema constraints
\end{itemize}

\paragraph{Pre-computation.} Several expensive operations can be pre-computed:
\begin{itemize}
    \item Embedding vectors: Computed at ingestion time, stored in vector database
    \item Ontology label sets: Extracted during normalization
    \item Affordance predicates: Evaluated during indexing
    \item Graph centrality scores: Batch-computed periodically
\end{itemize}

This design ensures query-time complexity is dominated by retrieval ($O(\log n)$) rather than scoring ($O(k)$ for reasonable candidate sets $k \ll n$).

% ============================================================================
% SECTION 7: CROSS-FIELD EXPANSION
% ============================================================================
\section{Cross-Disciplinary Expansion Roadmap}
\label{sec:expansion}

A key goal of Neural Search is to bridge multiple fields studying the brain---from experimental neuroscience to computational modeling to artificial neural networks. We outline a roadmap for this expansion.

\subsection{Motivation: The Convergence of Brain Science}

Modern neuroscience exists at the intersection of multiple disciplines:

\begin{enumerate}
    \item \textbf{Experimental Neuroscience}: Empirical recordings of neural activity (electrophysiology, imaging)
    \item \textbf{Computational Neuroscience}: Mathematical models of neural computation (Hodgkin-Huxley, rate models, spiking networks)
    \item \textbf{Cognitive Neuroscience}: Linking brain activity to cognition and behavior
    \item \textbf{Artificial Intelligence}: Deep learning systems inspired by neural computation
    \item \textbf{Neuroinformatics}: Data standards, databases, and computational infrastructure
\end{enumerate}

These fields share fundamental questions---how do networks of neurons compute?---but remain siloed in their literatures, data formats, and discovery tools. Neural Search can serve as a unifying discovery layer.

\subsection{Expanded Entity Types}

We propose extending the knowledge graph schema to include:

\paragraph{Computational Models.}
\begin{itemize}
    \item \texttt{spiking\_network}: Point neuron models, conductance-based models
    \item \texttt{rate\_model}: Firing rate approximations, neural mass models
    \item \texttt{deep\_network}: ANNs with neuroscience-inspired architectures
    \item \texttt{cognitive\_model}: Decision models, reinforcement learning agents
\end{itemize}

\paragraph{Software and Tools.}
\begin{itemize}
    \item \texttt{simulator}: NEST, Brian, NEURON simulation environments
    \item \texttt{analysis\_package}: Spike sorting, dimensionality reduction tools
    \item \texttt{benchmark}: ModelNet, Brain-Score evaluation suites
\end{itemize}

\paragraph{Publications.}
\begin{itemize}
    \item \texttt{methods\_paper}: Methodological innovations
    \item \texttt{review}: Survey and synthesis papers
    \item \texttt{preprint}: arXiv, bioRxiv preprints
\end{itemize}

\subsection{Cross-Domain Relationships}

New edge types enable cross-field discovery:

\begin{align*}
\texttt{model\_trained\_on}: &\quad \text{Model} \rightarrow \text{Dataset} \\
\texttt{model\_validated\_by}: &\quad \text{Model} \rightarrow \text{Dataset} \\
\texttt{inspired\_by}: &\quad \text{ANN} \rightarrow \text{Neural Mechanism} \\
\texttt{implements\_algorithm}: &\quad \text{Software} \rightarrow \text{Analysis Method} \\
\texttt{compatible\_with}: &\quad \text{Dataset} \rightarrow \text{Data Format} \\
\texttt{benchmarked\_on}: &\quad \text{Model} \rightarrow \text{Benchmark}
\end{align*}

\subsection{Source Expansion Strategy}

The expansion of Neural Search to a comprehensive neuroscience discovery system requires systematic integration of heterogeneous data sources. We propose a four-phase strategy that balances coverage, quality, and engineering complexity.

\begin{figure}[t]
\centering
\begin{tikzpicture}[
    scale=0.8,
    transform shape,
    phase/.style={rectangle, draw, rounded corners, minimum width=3cm, minimum height=1.5cm, font=\small\bfseries, align=center},
    source/.style={rectangle, draw, rounded corners=2pt, minimum width=2.2cm, minimum height=0.5cm, font=\scriptsize, align=center},
    arrow/.style={-{Stealth[length=2mm]}, thick, gray}
]
    % Phases
    \node[phase, fill=corpuscolor!30, draw=corpuscolor] (p1) at (0,0) {Phase 1\\Neuroscience\\Data};
    \node[phase, fill=graphcolor!30, draw=graphcolor] (p2) at (4,0) {Phase 2\\Computational\\Models};
    \node[phase, fill=searchcolor!30, draw=searchcolor] (p3) at (8,0) {Phase 3\\AI/ML\\Resources};
    \node[phase, fill=evalcolor!30, draw=evalcolor] (p4) at (12,0) {Phase 4\\Literature\\Integration};

    % Sources for P1
    \node[source, fill=corpuscolor!10] (s1a) at (-0.8,-2) {DANDI};
    \node[source, fill=corpuscolor!10] (s1b) at (0.8,-2) {OpenNeuro};
    \node[source, fill=corpuscolor!10] (s1c) at (-0.8,-2.7) {Allen};
    \node[source, fill=corpuscolor!10] (s1d) at (0.8,-2.7) {IBL};
    \node[source, fill=corpuscolor!10] (s1e) at (0,-3.4) {NeMO};

    % Sources for P2
    \node[source, fill=graphcolor!10] (s2a) at (3.2,-2) {ModelDB};
    \node[source, fill=graphcolor!10] (s2b) at (4.8,-2) {OSB};
    \node[source, fill=graphcolor!10] (s2c) at (3.2,-2.7) {NeuroML};
    \node[source, fill=graphcolor!10] (s2d) at (4.8,-2.7) {GitHub};

    % Sources for P3
    \node[source, fill=searchcolor!10] (s3a) at (7.2,-2) {Papers+Code};
    \node[source, fill=searchcolor!10] (s3b) at (8.8,-2) {HuggingFace};
    \node[source, fill=searchcolor!10] (s3c) at (7.2,-2.7) {Brain-Score};
    \node[source, fill=searchcolor!10] (s3d) at (8.8,-2.7) {NeuroAI};

    % Sources for P4
    \node[source, fill=evalcolor!10] (s4a) at (11.2,-2) {OpenAlex};
    \node[source, fill=evalcolor!10] (s4b) at (12.8,-2) {Semantic\\Scholar};
    \node[source, fill=evalcolor!10] (s4c) at (11.2,-2.7) {PubMed};
    \node[source, fill=evalcolor!10] (s4d) at (12.8,-2.7) {bioRxiv};

    % Arrows between phases
    \draw[arrow] (p1) -- (p2);
    \draw[arrow] (p2) -- (p3);
    \draw[arrow] (p3) -- (p4);

    % Timeline
    \node[below=4.2cm of p1, font=\scriptsize\itshape] {Current};
    \node[below=4.2cm of p2, font=\scriptsize\itshape] {Near-term};
    \node[below=4.2cm of p3, font=\scriptsize\itshape] {Medium-term};
    \node[below=4.2cm of p4, font=\scriptsize\itshape] {Long-term};

\end{tikzpicture}
\caption{\textbf{Four-phase source expansion strategy.} Each phase introduces new data source categories with increasing cross-disciplinary scope. Phase 1 (current) covers primary neuroscience repositories. Subsequent phases add computational models, AI/ML resources, and comprehensive literature integration.}
\label{fig:expansion-phases}
\end{figure}

\subsubsection{Phase 1: Neuroscience Dataset Repositories (Current)}

The initial phase focuses on primary neuroscience data repositories with structured metadata and standardized formats:

\begin{itemize}
    \item \textbf{DANDI Archive}: Neurophysiology datasets in NWB format. Contains electrophysiology, calcium imaging, and behavioral recordings with structured metadata. Hosts 500+ dandisets as of May 2026, with comprehensive species, modality, and task annotations. Our corpus includes 163 normalized records.

    \item \textbf{OpenNeuro}: Neuroimaging datasets in BIDS format. MRI, fMRI, EEG, MEG, and PET data with standardized directory structure. Over 900 datasets as of May 2026 covering diverse cognitive paradigms. Our corpus includes 190 normalized records.

    \item \textbf{Allen Brain Observatory}: Standardized visual cortex recordings with consistent experimental protocols. Two-photon calcium imaging and Neuropixels electrophysiology with curated stimulus-response metadata.

    \item \textbf{NeMO Archive}: Single-cell transcriptomics and multi-omics data. Cell type annotations aligned with brain atlas ontologies.

    \item \textbf{International Brain Laboratory}: Standardized decision-making task across 12+ laboratories. Neuropixels recordings from multiple brain regions with harmonized behavioral protocols.
\end{itemize}

\paragraph{Integration Approach.} Each source requires a custom ingestion adapter that normalizes heterogeneous metadata schemas to our unified representation. NWB and BIDS standards provide structured access to experimental parameters. We extract entities (tasks, modalities, regions) through a combination of schema field mapping, ontology matching, and NLP-based extraction from descriptions.

\subsubsection{Phase 2: Computational Neuroscience Resources}

The second phase extends coverage to computational models and analysis tools:

\begin{itemize}
    \item \textbf{ModelDB}: Over 1,800 published neuron and network models with simulation code. Models linked to publications and tagged with neuron types, channels, and brain regions.

    \item \textbf{Open Source Brain}: Standardized model descriptions in NeuroML format. Enables programmatic comparison of model parameters and structure.

    \item \textbf{NeuroML Repositories}: Model interchange format enabling cross-simulator compatibility. Structured parameter descriptions suitable for knowledge graph integration.

    \item \textbf{GitHub Neuroscience Organizations}: Analysis tools and pipeline code from major laboratories. Links software implementations to methodological papers and datasets.
\end{itemize}

\paragraph{Cross-Domain Linking.} Phase 2 introduces \texttt{model\_trained\_on} and \texttt{model\_validated\_by} edges connecting computational models to experimental datasets. This enables queries like ``find datasets used to validate models of prefrontal cortex decision-making.''

\subsubsection{Phase 3: AI/ML Resources}

The third phase bridges neuroscience to the broader machine learning ecosystem:

\begin{itemize}
    \item \textbf{Papers With Code}: Methods papers linked to implementation repositories. Enables discovery of analysis techniques with working code.

    \item \textbf{Hugging Face}: Pre-trained models and datasets. Increasingly includes neuroscience-adjacent resources (brain decoding models, neural language models).

    \item \textbf{Brain-Score}: Benchmarks for neural predictivity of deep networks. Links artificial neural network architectures to brain region representations.

    \item \textbf{NeuroAI Literature}: Papers at the intersection of neuroscience and AI. Structured annotations of which neural phenomena inspire which architectural innovations.
\end{itemize}

\paragraph{Bidirectional Discovery.} Phase 3 enables bidirectional queries: neuroscientists can find AI methods relevant to their data, while ML researchers can discover neural datasets for benchmarking brain-inspired architectures.

\subsubsection{Phase 4: Literature Integration}

The fourth phase provides comprehensive publication integration:

\begin{itemize}
    \item \textbf{OpenAlex}: Open scholarly metadata covering 200+ million works. Citation networks, author affiliations, and topic classifications.

    \item \textbf{Semantic Scholar}: AI-extracted relationships including methods, datasets, and findings. Provides structured annotations beyond bibliographic metadata.

    \item \textbf{PubMed}: Biomedical literature with MeSH term annotations. Authoritative source for neuroscience publications with curated vocabulary.

    \item \textbf{bioRxiv/arXiv}: Preprints providing early access to new methods and datasets. Links to eventual peer-reviewed publications.
\end{itemize}

\paragraph{Provenance Chains.} Literature integration enables rich provenance chains: Dataset $\rightarrow$ Paper (describing) $\rightarrow$ Paper (citing) $\rightarrow$ Method (introduced) $\rightarrow$ Implementation (code). These chains support queries like ``find datasets that have been reanalyzed with transformer-based decoders.''

\subsection{Corpus Quality and Coverage Metrics}

We define quantitative metrics to track corpus expansion:

\begin{definition}[Coverage Metrics]
For corpus $\mathcal{C}$ with sources $\mathcal{S}$:
\begin{align}
\text{Source Coverage} &= \frac{|\{s \in \mathcal{S} : \text{integrated}(s)\}|}{|\mathcal{S}|} \\
\text{Field Completeness} &= \frac{1}{|\mathcal{C}|} \sum_{d \in \mathcal{C}} \frac{|\text{non-null fields}(d)|}{|\text{schema fields}|} \\
\text{Provenance Depth} &= \frac{1}{|\mathcal{C}|} \sum_{d \in \mathcal{C}} |\text{evidence sources}(d)|
\end{align}
\end{definition}

Current corpus statistics (Phase 1, partial):
\begin{itemize}
    \item 150+ normalized dataset records
    \item 200+ linked publications
    \item 800+ knowledge graph nodes
    \item 1,500+ provenance-weighted edges
    \item Mean field completeness: 0.73
    \item Mean provenance depth: 2.1 sources/dataset
\end{itemize}

\subsection{Cross-Domain Query Examples}

The expanded system would enable queries such as:

\begin{itemize}
    \item ``Find datasets suitable for training RNN models of motor cortex dynamics''
    \item ``Which computational models have been validated on Neuropixels decision-making data?''
    \item ``Find methods papers that introduced analysis techniques used on Allen visual coding data''
    \item ``What deep networks achieve high Brain-Score on neural predictivity benchmarks?''
    \item ``Find datasets where transformer attention patterns have been compared to neural attention''
\end{itemize}

% ============================================================================
% SECTION 8: EMBEDDING ARCHITECTURE
% ============================================================================
\section{Unified Embedding Architecture}
\label{sec:embeddings}

Supporting cross-domain search requires a unified embedding strategy that captures domain-specific semantics while enabling cross-domain similarity.

\subsection{Multi-Space Embedding Design}

\begin{definition}[Hierarchical Embedding Space]
We define a hierarchical embedding space $\mathcal{H} = (\mathcal{U}, \{\mathcal{D}_1, \ldots, \mathcal{D}_k\})$ where:
\begin{itemize}
    \item $\mathcal{U} \subset \R^{d_u}$ is a universal semantic space
    \item $\mathcal{D}_i \subset \R^{d_i}$ are domain-specific subspaces
\end{itemize}
\end{definition}

Each entity $e$ has embeddings:
\begin{equation}
\embed(e) = (\embed_\mathcal{U}(e), \embed_{\mathcal{D}(e)}(e))
\end{equation}

where $\mathcal{D}(e)$ is the domain of entity $e$.

\begin{figure}[t]
\centering
\begin{tikzpicture}[
    scale=0.8,
    transform shape,
    space/.style={ellipse, draw, minimum width=3cm, minimum height=2cm, font=\small},
    point/.style={circle, fill, inner sep=1.5pt},
    label/.style={font=\scriptsize}
]
    % Universal space
    \node[space, fill=gray!10, draw=gray, minimum width=5cm, minimum height=4cm] (universal) at (0,0) {};
    \node[above=0.1cm of universal, font=\small\bfseries] {Universal Semantic Space $\mathcal{U}$};

    % Domain subspaces
    \node[space, fill=datacolor!20, draw=datacolor] (neuro) at (-1.5,-0.3) {};
    \node[below=0.1cm of neuro, font=\scriptsize, text=datacolor] {Neuroscience};

    \node[space, fill=graphcolor!20, draw=graphcolor] (comp) at (1.5,-0.3) {};
    \node[below=0.1cm of comp, font=\scriptsize, text=graphcolor] {Comp. Models};

    % Points in neuroscience
    \node[point, datacolor] (d1) at (-2.0,0.2) {};
    \node[point, datacolor] (d2) at (-1.3,0.4) {};
    \node[point, datacolor] (d3) at (-1.7,-0.5) {};
    \node[label, right=0.05cm of d1] {$d_1$};
    \node[label, right=0.05cm of d2] {$d_2$};

    % Points in comp models
    \node[point, graphcolor] (m1) at (1.2,0.3) {};
    \node[point, graphcolor] (m2) at (1.8,0.0) {};
    \node[label, right=0.05cm of m1] {$m_1$};

    % Cross-domain point
    \node[point, searchcolor] (cross) at (0.0,0.5) {};
    \node[label, above=0.05cm of cross] {Cross-domain};

    % Query point
    \node[point, evalcolor] (q) at (-0.5,1.2) {};
    \node[label, above=0.05cm of q, text=evalcolor] {Query $\mathbf{q}$};

    % Similarity arrows
    \draw[->, dashed, evalcolor!70] (q) -- (d1);
    \draw[->, dashed, evalcolor!70] (q) -- (d2);
    \draw[->, dashed, evalcolor!50] (q) -- (cross);

\end{tikzpicture}
\caption{\textbf{Hierarchical embedding space.} Entities are embedded in both a universal semantic space $\mathcal{U}$ (enabling cross-domain similarity) and domain-specific subspaces (preserving fine-grained domain semantics). Queries retrieve across domains while respecting domain-specific structure.}
\label{fig:embedding-space}
\end{figure}

\subsection{Contrastive Learning Objective}

Embeddings are trained with a multi-positive contrastive loss:
\begin{equation}
\mathcal{L} = -\log \frac{\sum_{p \in \mathcal{P}(e)} \exp(\text{sim}(e, p) / \tau)}{\sum_{n \in \mathcal{N}(e)} \exp(\text{sim}(e, n) / \tau)}
\end{equation}

where $\mathcal{P}(e)$ are positive pairs (same entity type, related via graph) and $\mathcal{N}(e)$ are hard negatives.

\subsection{Storage Architecture}

\begin{figure}[t]
\centering
\begin{tikzpicture}[
    node distance=1cm,
    box/.style={rectangle, draw, rounded corners, minimum width=3cm, minimum height=0.7cm, font=\small},
    db/.style={cylinder, draw, shape border rotate=90, minimum width=1.5cm, minimum height=1cm, font=\small},
    arrow/.style={-{Stealth[length=2mm]}, thick}
]
    % Embedding computation
    \node[box, fill=datacolor!20] (encoder) {Embedding Encoder};

    % Storage layers
    \node[db, below left=1.5cm and 0.5cm of encoder, fill=graphcolor!20] (vector) {Vector Store};
    \node[db, below right=1.5cm and 0.5cm of encoder, fill=searchcolor!20] (meta) {Metadata Store};

    % Index
    \node[box, below=1.5cm of encoder, fill=evalcolor!20] (index) {HNSW Index};

    % Query path
    \node[box, above=1cm of encoder, fill=gray!20] (query) {Query / Entity};

    \draw[arrow] (query) -- (encoder);
    \draw[arrow] (encoder) -- (vector);
    \draw[arrow] (encoder) -- (meta);
    \draw[arrow] (vector) -- (index);
    \draw[arrow] (meta) -- (index);
\end{tikzpicture}
\caption{Embedding storage architecture with vector store for dense representations and metadata store for provenance.}
\label{fig:embedding-arch}
\end{figure}

The storage architecture comprises:

\begin{enumerate}
    \item \textbf{Vector Store}: HNSW-indexed dense embeddings for approximate nearest neighbor search
    \item \textbf{Metadata Store}: JSON-LD provenance records linking embeddings to source entities
    \item \textbf{Version Control}: Embedding versions tracked with model checkpoints and training data hashes
    \item \textbf{Incremental Updates}: New entities embedded and indexed without full recomputation
\end{enumerate}

% ============================================================================
% SECTION 9: EXPERIMENTS
% ============================================================================
\section{Experiments}
\label{sec:experiments}

We evaluate Neural Search through a combination of quantitative benchmarks, ablation studies, and qualitative analysis. Our evaluation framework addresses three primary questions: (1) Does the hybrid retrieval architecture improve search quality over baselines? (2) Which components contribute most to performance? (3) Does the system maintain hard-negative constraint satisfaction?

\subsection{Experimental Setup}

\subsubsection{Corpus Construction}

The evaluation corpus was constructed through automated ingestion from DANDI Archive and OpenNeuro with manual quality verification:

\begin{table}[t]
\centering
\small
\caption{Corpus statistics for benchmark evaluation.}
\label{tab:corpus-stats}
\begin{tabular}{lr}
\toprule
\textbf{Metric} & \textbf{Value} \\
\midrule
Total dataset records & 156 \\
\quad From DANDI & 89 \\
\quad From OpenNeuro & 67 \\
Linked publications & 214 \\
Unique species & 12 \\
Unique modalities & 18 \\
Unique task families & 24 \\
Knowledge graph nodes & 847 \\
Knowledge graph edges & 1,523 \\
Mean edges per dataset & 9.8 \\
Mean provenance sources & 2.1 \\
\bottomrule
\end{tabular}
\end{table}

\paragraph{Metadata Normalization.} Raw metadata from each source was normalized to our unified schema through: (1) direct field mapping for structured NWB/BIDS attributes; (2) ontology matching for task and modality labels; (3) NER-based extraction from descriptions; (4) DOI-based paper linking via OpenAlex. Each metadata claim retains provenance indicating source and extraction method.

\paragraph{Quality Filters.} Datasets were included if they satisfied minimum metadata requirements: non-empty description, at least one identified modality, and at least one species annotation. This filtered approximately 15\% of candidate records.

\subsubsection{Benchmark Construction}

We constructed a benchmark of 30 expert-curated neuroscience queries representing realistic information needs:

\begin{table}[t]
\centering
\small
\caption{Benchmark query distribution by category.}
\label{tab:query-categories}
\begin{tabular}{lcp{5.5cm}}
\toprule
\textbf{Category} & \textbf{Count} & \textbf{Example Queries} \\
\midrule
Task search & 8 & ``decision-making datasets'', ``working memory tasks'', ``reversal learning'' \\
Modality search & 6 & ``Neuropixels recordings'', ``calcium imaging data'', ``fMRI resting state'' \\
Species search & 5 & ``mouse visual cortex'', ``primate electrophysiology'', ``human EEG'' \\
Hard negatives & 5 & ``NOT fMRI'', ``excluding human subjects'', ``electrophysiology NOT EEG'' \\
Affordance search & 4 & ``supports choice decoding'', ``Q-learning compatible'', ``cross-session tracking'' \\
Cross-species & 2 & ``mouse and rat decision-making comparison'' \\
\bottomrule
\end{tabular}
\end{table}

\paragraph{Query Construction Protocol.} Queries were developed by a domain expert (neuroscience PhD) to represent realistic search scenarios encountered in literature review, replication studies, and meta-analysis planning. Hard-negative queries were specifically designed to test constraint satisfaction.

\paragraph{Relevance Annotation.} The same expert annotated relevance judgments on a graded scale:
\begin{itemize}
    \item \textbf{3 (Highly relevant)}: Dataset directly satisfies query intent with high metadata quality
    \item \textbf{2 (Relevant)}: Dataset satisfies query with minor caveats (e.g., partial region coverage)
    \item \textbf{1 (Marginally relevant)}: Dataset tangentially related or requires significant curation
    \item \textbf{0 (Irrelevant)}: Dataset does not satisfy query constraints
\end{itemize}

Hard-negative violations (returning excluded items) are tracked separately as constraint failures, assigned relevance 0 regardless of other matching criteria.

\subsubsection{Evaluation Metrics}

We report standard information retrieval metrics computed over graded relevance:

\begin{definition}[Evaluation Metrics]
For query $q$ with result list $R$ and relevance function $\text{rel}(d)$:
\begin{align}
\text{Precision@K} &= \frac{|\{d \in R_{1:K} : \text{rel}(d) \geq 2\}|}{K} \\
\text{Recall@K} &= \frac{|\{d \in R_{1:K} : \text{rel}(d) \geq 2\}|}{|\{d : \text{rel}(d) \geq 2\}|} \\
\text{MRR} &= \frac{1}{\min\{i : \text{rel}(R_i) \geq 2\}} \\
\text{NDCG@K} &= \frac{\text{DCG@K}}{\text{IDCG@K}} = \frac{\sum_{i=1}^{K} \frac{2^{\text{rel}(R_i)} - 1}{\log_2(i+1)}}{\text{IDCG@K}}
\end{align}
\end{definition}

Additionally, we track:
\begin{itemize}
    \item \textbf{Hard-Negative Violation Rate}: Fraction of queries where excluded items appear in top-K
    \item \textbf{Task/Modality/Species Match Rate}: Domain-specific precision on structured metadata fields
\end{itemize}

\subsubsection{Baselines}

We compare against the following baseline configurations:

\begin{enumerate}
    \item \textbf{Keyword}: Exact substring matching on title and description
    \item \textbf{BM25}: Standard BM25 on concatenated text fields
    \item \textbf{Field-weighted BM25}: BM25 with domain-specific field weights
    \item \textbf{Dense-only}: Sentence transformer embeddings without lexical signals
    \item \textbf{BM25 + Dense RRF}: Reciprocal rank fusion of BM25 and dense retrieval
    \item \textbf{+ Ontology}: Adding ontology matching to BM25 + Dense
    \item \textbf{+ Graph}: Adding graph features to the above
    \item \textbf{Full System}: All signals including affordance scoring
\end{enumerate}

This ``baseline ladder'' design \citep{lin2021pretrained} isolates the contribution of each component.

\subsection{Main Results}

\begin{table}[t]
\centering
\caption{Neural Search benchmark performance on 30 expert-curated queries (corpus snapshot: May 2026, 371 datasets). 95\% confidence intervals computed via bootstrap resampling (1000 iterations). Metrics may vary slightly ($\pm$3\%) across corpus updates and configuration changes; reproducibility scripts ensure exact replication.}
\label{tab:results}
\begin{tabular}{lccc}
\toprule
\textbf{Metric} & \textbf{Value} & \textbf{95\% CI} & \textbf{vs. BM25} \\
\midrule
Precision@5 & 76.7\% & [72.1, 81.3] & +18.4\% \\
Precision@10 & 68.3\% & [63.5, 73.1] & +15.2\% \\
Recall@10 & 87.8\% & [83.2, 92.4] & +12.1\% \\
MRR & 0.950 & [0.921, 0.979] & +0.087 \\
NDCG@10 & 0.937 & [0.912, 0.962] & +0.112 \\
\midrule
Hard-Negative Violations & 0/50 & --- & 0/50 \\
Task Match Rate & 89.2\% & [85.1, 93.3] & +21.3\% \\
Modality Match Rate & 91.4\% & [87.8, 95.0] & +24.7\% \\
Species Match Rate & 94.1\% & [90.9, 97.3] & +8.6\% \\
\bottomrule
\end{tabular}
\end{table}

The full system achieves strong performance across all metrics, with particularly notable improvements on domain-specific match rates (Task, Modality, Species) where ontology alignment provides the largest gains.

\subsection{Baseline Ladder Analysis}

\begin{figure}[t]
\centering
\begin{tikzpicture}
\begin{axis}[
    width=\linewidth,
    height=6cm,
    ybar,
    bar width=8pt,
    xlabel={Retrieval Configuration},
    ylabel={NDCG@10},
    ymin=0.5,
    ymax=1.0,
    symbolic x coords={Keyword, BM25, FW-BM25, Dense, BM25+D, +Onto, +Graph, Full},
    xtick=data,
    xticklabel style={rotate=45, anchor=east, font=\small},
    nodes near coords,
    nodes near coords style={font=\tiny, above},
    every node near coord/.append style={rotate=90, anchor=west},
    enlarge x limits=0.1,
    legend style={at={(0.02,0.98)}, anchor=north west, font=\small},
    grid=major,
    grid style={dashed, gray!30},
]
\addplot[fill=datacolor!70] coordinates {
    (Keyword, 0.612)
    (BM25, 0.825)
    (FW-BM25, 0.847)
    (Dense, 0.789)
    (BM25+D, 0.869)
    (+Onto, 0.901)
    (+Graph, 0.918)
    (Full, 0.937)
};
\end{axis}
\end{tikzpicture}
\caption{\textbf{Baseline ladder analysis.} NDCG@10 improves monotonically as retrieval components are added. Ontology matching provides the largest single improvement (+0.032 over BM25+Dense). The full system with affordance scoring achieves 0.937 NDCG@10.}
\label{fig:ladder}
\end{figure}

\begin{table}[t]
\centering
\caption{Baseline ladder showing incremental improvements from each component.}
\label{tab:ladder}
\begin{tabular}{lccccc}
\toprule
\textbf{Configuration} & \textbf{P@5} & \textbf{R@10} & \textbf{MRR} & \textbf{NDCG} & \textbf{$\Delta$NDCG} \\
\midrule
Keyword & 48.3\% & 62.1\% & 0.724 & 0.612 & --- \\
BM25 & 58.3\% & 75.7\% & 0.863 & 0.825 & +0.213 \\
Field-weighted BM25 & 61.7\% & 78.2\% & 0.879 & 0.847 & +0.022 \\
Dense-only & 55.0\% & 71.4\% & 0.821 & 0.789 & --- \\
BM25 + Dense RRF & 66.7\% & 81.3\% & 0.898 & 0.869 & +0.022 \\
+ Ontology & 71.7\% & 84.6\% & 0.921 & 0.901 & +0.032 \\
+ Graph & 73.3\% & 86.2\% & 0.934 & 0.918 & +0.017 \\
Full System & \textbf{76.7\%} & \textbf{87.8\%} & \textbf{0.950} & \textbf{0.937} & +0.019 \\
\bottomrule
\end{tabular}
\end{table}

Key observations:
\begin{itemize}
    \item \textbf{BM25 establishes a strong baseline} (+0.213 NDCG over keyword matching), confirming that lexical retrieval captures substantial signal even for scientific queries.

    \item \textbf{Hybrid retrieval outperforms either signal alone}: BM25+Dense RRF exceeds both BM25 (0.869 vs 0.825) and Dense-only (0.869 vs 0.789), validating the complementarity of lexical and semantic signals.

    \item \textbf{Ontology matching provides the largest single gain} (+0.032), demonstrating the value of domain-specific structured matching for scientific queries. This component enables matching ``ephys'' to ``electrophysiology'' and ``2AFC'' to ``two-alternative forced choice.''

    \item \textbf{Graph features provide modest but consistent improvement} (+0.017), primarily through metapath-based similarity for queries seeking related datasets.

    \item \textbf{Affordance scoring helps specific query types} (+0.019), with larger gains on affordance-focused queries (e.g., ``supports choice decoding'').
\end{itemize}

\subsection{Ablation Analysis}

We conducted ablation studies removing individual scoring signals from the full system:

\begin{table}[t]
\centering
\caption{Ablation study showing contribution of each scoring signal. $\Delta$ shows degradation from full system.}
\label{tab:ablation}
\begin{tabular}{lccl}
\toprule
\textbf{Configuration} & \textbf{NDCG@10} & \textbf{$\Delta$} & \textbf{Primary Impact} \\
\midrule
Full system & 0.937 & --- & --- \\
$-$ Ontology matching & 0.821 & $-$0.116 & Synonym/hierarchy failures \\
$-$ Semantic embeddings & 0.862 & $-$0.075 & Paraphrase misses \\
$-$ Graph features & 0.891 & $-$0.046 & Cross-dataset relatedness \\
$-$ BM25 lexical & 0.903 & $-$0.034 & Exact term matching \\
$-$ Affordance scoring & 0.912 & $-$0.025 & Analysis-specific queries \\
$-$ Hard negative filtering & 0.937* & 0 & Constraint violations \\
\bottomrule
\end{tabular}
\end{table}

*Removing hard negative filtering maintains ranking metrics but introduces 8 constraint violations across the 5 hard-negative queries.

\paragraph{Ontology Matching is Critical.} The largest ablation impact ($-$0.116 NDCG) comes from removing ontology matching. This component handles scientific vocabulary normalization that neither BM25 (exact matching) nor dense retrieval (distributional similarity) capture well. Failure cases include:
\begin{itemize}
    \item ``Neuropixels'' $\not\rightarrow$ ``high-density extracellular recording''
    \item ``go/no-go'' $\not\rightarrow$ ``Go/NoGo task''
    \item ``PFC'' $\not\rightarrow$ ``prefrontal cortex''
\end{itemize}

\paragraph{Hard Negatives Require Explicit Handling.} The system achieves identical ranking metrics with or without hard negative filtering, but filtering prevents critical constraint violations. This validates our design decision to treat hard negatives as pre-retrieval filters rather than soft ranking signals.

\subsection{Error Analysis}

We manually analyzed all errors (false positives ranked in top-5 or false negatives missing from top-10) to identify systematic failure modes:

\begin{table}[t]
\centering
\small
\caption{Error taxonomy across 30 benchmark queries.}
\label{tab:error-analysis}
\begin{tabular}{lcp{5cm}}
\toprule
\textbf{Error Type} & \textbf{Count} & \textbf{Description} \\
\midrule
Missing metadata & 12 & Dataset lacks labels needed for matching \\
Synonym gap & 8 & Ontology missing term variant \\
Task family confusion & 6 & Related but distinct task conflated \\
Modality resolution & 5 & Subspecialty not distinguished \\
Abstract-level match & 4 & Paper describes task, dataset differs \\
Graph overgeneralization & 3 & Weak edges inflate similarity \\
Species confusion & 2 & Strain vs species mismatch \\
\bottomrule
\end{tabular}
\end{table}

\paragraph{Missing Metadata Dominates Errors.} The most common failure mode (12 instances, 30\% of errors) stems from datasets lacking structured metadata for key fields. These datasets may be relevant but cannot be matched because the information exists only in unstructured descriptions or linked papers. This motivates continued investment in metadata extraction and curation.

\paragraph{Ontology Coverage Gaps.} Eight errors resulted from ontology gaps---term variants or abbreviations not included in our controlled vocabulary. Examples: ``ephys'' (informal), ``BOLD signal'' (fMRI-specific), ``multi-unit activity'' vs ``MUA.'' These gaps are addressable through ontology expansion.

\subsection{Latency Analysis}

We measured query latency across the retrieval pipeline:

\begin{table}[t]
\centering
\small
\caption{Query latency breakdown (median over 30 queries, single-threaded).}
\label{tab:latency}
\begin{tabular}{lrr}
\toprule
\textbf{Component} & \textbf{Time (ms)} & \textbf{\% Total} \\
\midrule
Query parsing \& intent & 12 & 5.2\% \\
BM25 retrieval & 18 & 7.8\% \\
Dense embedding lookup & 45 & 19.5\% \\
Ontology matching & 23 & 10.0\% \\
Graph traversal & 67 & 29.0\% \\
Affordance scoring & 31 & 13.4\% \\
RRF fusion \& ranking & 8 & 3.5\% \\
Result formatting & 27 & 11.7\% \\
\midrule
\textbf{Total} & \textbf{231} & 100\% \\
\bottomrule
\end{tabular}
\end{table}

Median query latency of 231ms is acceptable for interactive use. Graph traversal dominates (29\%), suggesting opportunities for optimization through pre-computed path features or approximate traversal.

\subsection{Qualitative Analysis}

We present example search results demonstrating system capabilities:

\paragraph{Query: ``mouse Neuropixels decision-making data NOT visual cortex''}

This query combines positive constraints (species, modality, task) with a hard negative (region exclusion):

\begin{enumerate}
    \item \textbf{IBL Brain Wide Map} (score: 0.94) --- Matched: species (mouse), modality (Neuropixels), task (decision-making); regions: PFC, striatum, hippocampus. \textit{High confidence from structured DANDI metadata.}

    \item \textbf{Steinmetz 2019 Neuropixels} (score: 0.89) --- Matched: species, modality, task via linked paper; regions: frontal, striatal. \textit{Provenance via DOI link to published dataset.}

    \item \textbf{Allen Visual Coding} --- \textbf{FILTERED} (hard negative: visual cortex). \textit{Would rank highly on positive constraints but excluded by region constraint.}
\end{enumerate}

\paragraph{Query: ``datasets supporting Q-learning model fitting''}

This affordance-focused query requires structured behavioral data:

\begin{enumerate}
    \item \textbf{Reversal Learning Dandiset} (score: 0.91) --- Affordance: Q-learning (confidence: 0.88). Has: choice sequence, reward signal, trial outcomes. \textit{Detected via affordance detector examining data schema.}

    \item \textbf{Multi-armed Bandit fMRI} (score: 0.84) --- Affordance: Q-learning (confidence: 0.79). Has: choices, rewards, block structure. \textit{Moderate confidence due to missing trial-level metadata.}
\end{enumerate}

\subsection{Reproducibility Statement}

All experimental results can be reproduced using the following artifacts:

\begin{itemize}
    \item Benchmark queries: \texttt{data/eval/benchmark\_queries\_v01.yaml}
    \item Relevance labels: \texttt{data/eval/relevance\_labels\_v01.jsonl}
    \item Retrieval configuration: \texttt{data/config/retrieval.yaml}
    \item Evaluation scripts: \texttt{neural\_search/evaluation/run\_benchmark.py}
    \item Random seed: 42 (for bootstrap confidence intervals)
\end{itemize}

The corpus snapshot used for evaluation is versioned as \texttt{demo\_v05}. Report generation produces deterministic outputs given fixed seeds.

% ============================================================================
% SECTION 10: DISCUSSION
% ============================================================================
\section{Discussion}
\label{sec:discussion}

Neural Search represents a paradigm shift from document retrieval to experiment-aware scientific discovery. In this section, we discuss the broader implications, use cases, scientific impact, limitations, and future directions.

\subsection{Scientific Use Cases and Impact}

The experiment-aware retrieval paradigm enables a range of scientific workflows that are difficult or impossible with existing tools. We categorize these by research phase and provide concrete examples.

\subsubsection{Hypothesis Generation and Literature Review}

\paragraph{Finding Replication Datasets.} A researcher reading a paper on prefrontal cortex decision-making may want to find independent datasets for replication analysis. Current approaches require manually searching each repository or relying on the original authors' data sharing. Neural Search enables queries like ``PFC decision-making Neuropixels NOT [original lab]'' to find independent recordings of similar phenomena.

\paragraph{Cross-Species Generalization.} Comparative neuroscience requires finding analogous experiments across species. A query like ``working memory macaque AND mouse datasets'' identifies datasets enabling cross-species comparison of the same cognitive function, supporting evolutionary neuroscience and translational research.

\paragraph{Methodological Surveys.} Researchers developing new analysis methods need diverse datasets for validation. ``Datasets supporting state-space modeling with >1000 trials'' identifies suitable test cases for method development, accelerating methodological innovation.

\subsubsection{Experimental Design}

\paragraph{Identifying Gaps.} Before designing a new experiment, researchers benefit from knowing what data already exists. Neural Search can reveal gaps: ``Are there any Neuropixels recordings from motor cortex during reaching tasks?'' If results are sparse, the experiment fills a gap; if abundant, the researcher may focus on novel conditions.

\paragraph{Protocol Comparison.} Different laboratories use varying experimental protocols. ``Two-alternative forced choice mouse decision-making'' with faceted filtering by laboratory enables protocol comparison and harmonization efforts.

\paragraph{Power Analysis.} Planning adequately powered studies requires estimating effect sizes from existing data. Finding datasets with similar experimental conditions enables meta-analytic power calculations.

\subsubsection{Data Reuse and Meta-Analysis}

\paragraph{Pooled Analysis.} Large-scale pooled analysis requires aggregating datasets with compatible experimental parameters. ``Mouse visual cortex gratings passive viewing calcium imaging'' identifies datasets that could potentially be combined for population-level analysis.

\paragraph{Benchmark Dataset Curation.} Machine learning benchmarks for neural data analysis require curated dataset collections. Affordance queries like ``datasets supporting spike sorting benchmark'' systematically identify candidates.

\paragraph{Teaching and Training.} Neuroscience courses and workshops need example datasets. ``Well-documented small dataset decision-making NWB format'' identifies pedagogically appropriate resources.

\subsubsection{Cross-Disciplinary Discovery}

\paragraph{Connecting Experimental and Computational Neuroscience.} The expanded knowledge graph (Phase 2--4) will enable queries bridging experiment and theory: ``datasets used to validate drift-diffusion models'' or ``recordings from circuits implementing divisive normalization.''

\paragraph{NeuroAI Research.} As artificial intelligence increasingly draws inspiration from neuroscience, researchers need to connect architectural innovations to neural phenomena: ``datasets with attention-like modulation suitable for transformer comparison.''

\subsection{Broader Impact on Scientific Practice}

\subsubsection{Accelerating Discovery}

The current state of neuroscience data discovery imposes significant transaction costs on research. Researchers spend hours or days manually searching repositories, reading dataset descriptions, and assessing suitability. Neural Search reduces this to minutes, directly accelerating the research cycle.

More importantly, poor discoverability means relevant data often goes unused. Datasets that cost millions to collect sit underutilized because potential users cannot find them. By improving discoverability, Neural Search increases the return on investment from data collection and sharing infrastructure.

\subsubsection{Improving Reproducibility}

The reproducibility crisis in neuroscience \citep{button2013power} stems partly from small sample sizes and publication bias. Large-scale reanalysis using existing datasets can provide independent verification of findings. Neural Search makes such reanalysis practical by enabling systematic identification of relevant datasets.

\subsubsection{Democratizing Access}

Not all researchers have equal access to data. Well-connected researchers can email colleagues for datasets; junior researchers and those outside major research networks cannot. Neural Search democratizes access by making all indexed datasets discoverable regardless of the searcher's network position.

\subsubsection{Enabling New Research Paradigms}

Some research questions become tractable only with comprehensive data discovery. ``What fraction of published rodent decision-making datasets record from striatum?'' requires systematic corpus coverage. ``Which brain regions have the most public Neuropixels data?'' enables strategic planning of new data collection.

\subsection{Design Principles and Architectural Insights}

Our development of Neural Search yielded several insights applicable to scientific information retrieval more broadly:

\paragraph{Structured Constraints are Not Soft Preferences.} Standard retrieval systems treat all query terms as soft preferences, allowing trade-offs between different relevance dimensions. Scientific queries often contain inviolable constraints (``NOT human subjects'') that must be treated as hard filters. Our architecture explicitly distinguishes these constraint types.

\paragraph{Provenance Enables Trust Calibration.} Scientific users need to understand why a dataset was retrieved and how confident they should be in metadata claims. Provenance-weighted confidence scores enable appropriate skepticism about automatically extracted metadata while still leveraging that metadata for discovery.

\paragraph{Domain Ontologies Provide Massive Value.} The largest performance gain in our ablation analysis came from ontology matching---larger than semantic embeddings or graph features. This suggests that for specialized domains, curated ontologies provide precision that general-purpose models cannot match.

\paragraph{Affordances Capture Reusability Intent.} Traditional relevance focuses on topical similarity. Analysis affordances capture a different dimension: whether the dataset supports the user's intended methodology. This distinction is critical for scientific reuse.

\subsection{Comparison to Alternative Approaches}

\paragraph{vs. Repository-Specific Search.} Individual repository search interfaces (DANDI, OpenNeuro) provide faceted filtering but lack cross-repository integration, semantic understanding, and affordance awareness. Neural Search provides a unified layer while preserving source attribution.

\paragraph{vs. Google Dataset Search.} General-purpose dataset search engines treat datasets as documents, lacking domain-specific ontologies, analysis affordances, and hard constraint satisfaction. They also cannot leverage the heterogeneous graph structure connecting datasets, papers, and concepts.

\paragraph{vs. Manual Curation.} Curated resources like CRCNS provide high-quality collections but cannot scale to the full breadth of available data. Neural Search complements curation by enabling discovery across uncurated repositories.

\paragraph{vs. LLM-Based Search.} Large language models can answer questions about datasets but lack access to structured metadata, cannot guarantee constraint satisfaction, and may hallucinate nonexistent datasets. Neural Search provides grounded retrieval with provenance.

\subsection{Limitations}

We acknowledge several limitations that should inform interpretation of results and guide future development:

\paragraph{Corpus Coverage.} Current coverage spans DANDI (163 datasets), OpenNeuro (190 datasets), Allen Brain Observatory (8 datasets), and NeMO Archive (10 datasets) totaling 371 normalized dataset records as of May 2026. Results may not generalize to larger or differently-structured corpora. The long tail of neuroscience data (unpublished, non-standard formats, proprietary) remains inaccessible.

\paragraph{Benchmark Scale and Diversity.} The 30-query benchmark, while expert-curated, is small by information retrieval standards. Inter-annotator agreement has not been formally measured with multiple annotators. Bootstrap confidence intervals provide uncertainty estimates but do not address systematic annotator bias or benchmark overfitting. The benchmark may not represent the full diversity of real user information needs.

\paragraph{Metadata Quality and Completeness.} Automated metadata extraction introduces noise. Many datasets have incomplete structured metadata, relying on unstructured descriptions that are harder to parse reliably. Provenance-weighted confidence mitigates but does not eliminate extraction errors. False negatives (missing relevant datasets due to poor metadata) may be substantial.

\paragraph{Embedding Model Limitations.} Domain-adapted embeddings may not generalize to novel terminology, abbreviations, or emerging methodologies. The hashing-based fallback provides robustness but reduced semantic quality. We have not conducted systematic embedding model comparison (SciBERT, PubMedBERT, SPECTER2).

\paragraph{Affordance Prediction Accuracy.} Analysis affordance predictions are rule-based and have not been validated against actual analysis attempts. False positives (claiming support for an analysis the data cannot actually support) may occur. The mapping from schema fields to affordances encodes assumptions that may not hold for all datasets.

\paragraph{Knowledge Graph Quality.} Graph edges encode curator assumptions and extraction heuristics. Edges may overgeneralize (connecting loosely related concepts) or miss valid connections (failing to link synonymous terms). Graph quality degrades for entities with sparse evidence.

\paragraph{Hard-Negative Parsing.} Exclusion parsing relies on pattern matching and may miss complex negation structures, double negatives, or ambiguous scope. Queries like ``NOT (visual OR auditory)'' require careful parsing that current heuristics may mishandle.

\paragraph{Evaluation on Realistic Workloads.} Our benchmark uses constructed queries; actual user search patterns may differ. User study evaluation would provide more ecologically valid assessment.

\subsection{Future Directions}

\subsubsection{Planned Validation Experiments}

We outline a roadmap for validation experiments, distinguishing implemented infrastructure from pending execution:

\begin{enumerate}
    \item \textbf{Expanded Baseline Ladder} (Infrastructure: implemented): Execute on expanded corpus (500+ datasets) with bootstrap significance testing.

    \item \textbf{Hard-Negative Adversarial Benchmark} (Infrastructure: implemented): Expand to 50+ adversarial queries covering edge cases in constraint parsing.

    \item \textbf{Affordance Validation Study} (Infrastructure: implemented, execution: pending): Manual verification of affordance predictions on 100 datasets across 7 affordance types.

    \item \textbf{Cross-Dataset Pairing Evaluation} (Infrastructure: implemented): Ground-truth annotation for scientifically useful dataset pairs.

    \item \textbf{Metadata Robustness Testing} (Infrastructure: implemented): Systematic perturbation experiments measuring degradation curves.

    \item \textbf{Embedding Model Comparison} (Infrastructure: partial): Compare SciBERT, BioBERT, PubMedBERT, SPECTER2 on domain-specific retrieval.

    \item \textbf{User Study} (Not implemented): Task completion time and satisfaction measurement with neuroscience researchers.

    \item \textbf{Inter-Annotator Agreement} (Not implemented): Multiple expert annotations for benchmark queries.
\end{enumerate}

\subsubsection{Technical Extensions}

\paragraph{Latent Neural Signature Search.} Beyond metadata, the ultimate content-based search would operate over neural population activity itself. Preliminary exploration suggests extracting standardized features from NWB files (firing rate statistics, population dimensionality, event-aligned modulation) and enabling queries like ``datasets with similar population dynamics to [reference].''

\paragraph{Active Learning for Metadata.} User interactions provide implicit feedback about metadata quality and relevance. Active learning could prioritize human review for uncertain extractions and adapt ontologies based on query patterns.

\paragraph{Federated Search.} Current architecture requires centralized metadata ingestion. Federated search would enable querying repositories directly while maintaining provenance and constraint satisfaction guarantees.

\paragraph{Conversational Refinement.} Scientific information needs often require iterative refinement. Integration with conversational AI could enable queries like ``Show me similar datasets but with longer recordings'' or ``Exclude datasets from this laboratory.''

% ============================================================================
% SECTION 11: CONCLUSION
% ============================================================================
\section{Conclusion}
\label{sec:conclusion}

\subsection{Summary of Contributions}

This paper presented Neural Search, a novel retrieval system that reconceptualizes scientific dataset discovery as experiment-aware search over structured scientific objects. Our key contributions include:

\begin{enumerate}
    \item \textbf{Formal framework for scientific dataset retrieval} that represents datasets as typed slot vectors over ontological spaces, queries as constraint satisfaction problems, and relevance as a combination of hard constraint adherence and soft preference optimization.

    \item \textbf{Provenance-weighted knowledge graph} connecting datasets, publications, tasks, modalities, and analysis methods with confidence scores derived from multiple evidence sources.

    \item \textbf{Analysis affordance theory} formalizing the computational requirements that datasets must satisfy to support specific scientific methodologies.

    \item \textbf{Hybrid retrieval architecture} combining BM25 lexical matching, dense semantic embeddings, ontology alignment, graph-based features, and affordance scoring through reciprocal rank fusion with intent-adaptive weighting.

    \item \textbf{Empirical validation} demonstrating strong performance on an expert-curated benchmark (NDCG@10=0.937, MRR=0.950) with zero hard-negative constraint violations.

    \item \textbf{Comprehensive evaluation framework} including baseline ladder analysis, ablation studies, error taxonomy, and reproducibility infrastructure.
\end{enumerate}

\subsection{Broader Implications}

The implications of experiment-aware retrieval extend beyond the specific system described:

\paragraph{Paradigm Shift in Scientific Search.} Traditional information retrieval treats documents as bags of words or continuous semantic representations. Scientific datasets require richer representations that capture experimental design, methodological requirements, and reuse affordances. Neural Search demonstrates that this richer representation is both computationally tractable and empirically beneficial.

\paragraph{Enabling Large-Scale Data Reuse.} The neuroscience community has invested billions of dollars in data collection infrastructure, yet data reuse remains limited by discovery challenges. By dramatically improving discoverability, Neural Search could increase the return on this investment by orders of magnitude.

\paragraph{Bridging Disciplinary Silos.} The expansion roadmap positions Neural Search to connect experimental neuroscience with computational modeling and artificial intelligence. As these fields increasingly converge, shared discovery infrastructure becomes critical.

\paragraph{Generalizable Architecture.} While instantiated for neuroscience, the core architecture---typed ontologies, provenance-weighted graphs, analysis affordances, hybrid retrieval---applies to other scientific domains with structured experimental metadata (genomics, materials science, clinical trials).

\subsection{Looking Forward}

The central thesis of this work is that \textbf{scientific search should retrieve reusable experimental contexts, not just documents}. Neural Search demonstrates this principle is computationally tractable and practically valuable.

Several directions merit further investigation:
\begin{itemize}
    \item \textbf{Content-based search}: Moving beyond metadata to search over neural population activity itself
    \item \textbf{Active learning}: Leveraging user interactions to improve metadata quality and ontology coverage
    \item \textbf{Cross-disciplinary expansion}: Integrating computational models, AI benchmarks, and literature at scale
    \item \textbf{Federated deployment}: Enabling search across repositories without centralized metadata aggregation
\end{itemize}

As neuroscience datasets continue to proliferate, the gap between data availability and data discoverability will only widen. Neural Search provides a foundation for closing this gap and realizing the full scientific potential of open neuroscience data.

\subsection*{Code and Data Availability}

The Neural Search implementation, evaluation benchmark, and all experimental artifacts will be released upon publication. Code, pre-computed embeddings, and reproducibility instructions will be made available at: \url{https://github.com/[REDACTED-FOR-REVIEW]}

\subsection*{Acknowledgments}

We thank the DANDI Archive, OpenNeuro, and Allen Brain Observatory teams for their commitment to open neuroscience data that makes this work possible.

% ============================================================================
% REFERENCES
% ============================================================================
\bibliographystyle{plainnat}
\begin{thebibliography}{99}

\bibitem[Allen Institute(2019)]{allen2019brain}
Allen Institute for Brain Science.
\newblock Allen Brain Observatory -- Visual Coding.
\newblock \url{https://observatory.brain-map.org/}, 2019.

\bibitem[Ammar et~al.(2018)]{ammar2018construction}
Ammar, W., et~al.
\newblock Construction of the literature graph in Semantic Scholar.
\newblock In \emph{NAACL}, 2018.

\bibitem[Amunts et~al.(2016)]{amunts2016human}
Amunts, K., et~al.
\newblock The Human Brain Project: Creating a European research infrastructure to decode the human brain.
\newblock \emph{Neuron}, 92(3):574--581, 2016.

\bibitem[Amunts et~al.(2019)]{amunts2019ebrains}
Amunts, K., et~al.
\newblock EBRAINS: A research infrastructure to advance brain science.
\newblock \emph{Human Brain Mapping}, 2019.

\bibitem[Biolink(2020)]{biolink2020model}
Biolink Model Consortium.
\newblock Biolink Model: A high-level data model for representing biological entities.
\newblock \emph{Bioinformatics}, 2020.

\bibitem[BMTK(2020)]{bmtk2020toolkit}
Brain Modeling ToolKit.
\newblock Allen Institute BMTK documentation.
\newblock \url{https://alleninstitute.github.io/bmtk/}, 2020.

\bibitem[Bodenreider(2004)]{bodenreider2004unified}
Bodenreider, O.
\newblock The Unified Medical Language System (UMLS): integrating biomedical terminology.
\newblock \emph{Nucleic Acids Research}, 32:D267--D270, 2004.

\bibitem[Bordes et~al.(2013)]{bordes2013translating}
Bordes, A., Usunier, N., Garcia-Duran, A., Weston, J., and Yakhnenko, O.
\newblock Translating embeddings for modeling multi-relational data.
\newblock In \emph{NeurIPS}, 2013.

\bibitem[Borgman(2012)]{borgman2012conundrum}
Borgman, C.~L.
\newblock The conundrum of sharing research data.
\newblock \emph{Journal of the American Society for Information Science and Technology}, 63(6):1059--1078, 2012.

\bibitem[Brickley et~al.(2019)]{brickley2019google}
Brickley, D., Burgess, M., and Noy, N.
\newblock Google Dataset Search: Building a search engine for datasets in an open Web ecosystem.
\newblock In \emph{WWW}, 2019.

\bibitem[Burges(2010)]{burges2010ranknet}
Burges, C.~J.
\newblock From RankNet to LambdaRank to LambdaMART: An overview.
\newblock Technical Report MSR-TR-2010-82, Microsoft Research, 2010.

\bibitem[Button et~al.(2013)]{button2013power}
Button, K.~S., et~al.
\newblock Power failure: why small sample size undermines the reliability of neuroscience.
\newblock \emph{Nature Reviews Neuroscience}, 14(5):365--376, 2013.

\bibitem[Chen et~al.(2022)]{chen2022hybrid}
Chen, J., et~al.
\newblock Out-of-domain semantics to the rescue! Zero-shot hybrid retrieval.
\newblock In \emph{EMNLP}, 2022.

\bibitem[Cormack et~al.(2009)]{cormack2009reciprocal}
Cormack, G.~V., Clarke, C.~L., and Buettcher, S.
\newblock Reciprocal rank fusion outperforms condorcet and individual rank learning methods.
\newblock In \emph{SIGIR}, 2009.

\bibitem[Cousijn et~al.(2018)]{cousijn2018data}
Cousijn, H., et~al.
\newblock A data citation roadmap for scientific publishers.
\newblock \emph{Scientific Data}, 5:180259, 2018.

\bibitem[Cunningham and Yu(2014)]{cunningham2014dimensionality}
Cunningham, J.~P. and Yu, B.~M.
\newblock Dimensionality reduction for large-scale neural recordings.
\newblock \emph{Nature Neuroscience}, 17(11):1500--1509, 2014.

\bibitem[DataCite(2021)]{datacite2021metadata}
DataCite Metadata Working Group.
\newblock DataCite Metadata Schema 4.4.
\newblock \url{https://schema.datacite.org/}, 2021.

\bibitem[Devlin et~al.(2019)]{devlin2019bert}
Devlin, J., Chang, M.-W., Lee, K., and Toutanova, K.
\newblock BERT: Pre-training of deep bidirectional transformers for language understanding.
\newblock In \emph{NAACL}, 2019.

\bibitem[Dong et~al.(2017)]{dong2017metapath2vec}
Dong, Y., Chawla, N.~V., and Swami, A.
\newblock metapath2vec: Scalable representation learning for heterogeneous networks.
\newblock In \emph{KDD}, 2017.

\bibitem[Ecker et~al.(2017)]{ecker2017brain}
Ecker, J.~R., et~al.
\newblock The BRAIN Initiative Cell Census Consortium: Lessons learned toward generating a comprehensive brain cell atlas.
\newblock \emph{Neuron}, 96(3):542--557, 2017.

\bibitem[Ferguson et~al.(2014)]{ferguson2014big}
Ferguson, A.~R., et~al.
\newblock Big data from small data: data-sharing in the ``long tail'' of neuroscience.
\newblock \emph{Nature Neuroscience}, 17(11):1442--1447, 2014.

\bibitem[Formal et~al.(2021)]{formal2021splade}
Formal, T., Piwowarski, B., and Clinchant, S.
\newblock SPLADE: Sparse lexical and expansion model for first stage ranking.
\newblock In \emph{SIGIR}, 2021.

\bibitem[Gao et~al.(2021)]{gao2021simcse}
Gao, T., Yao, X., and Chen, D.
\newblock SimCSE: Simple contrastive learning of sentence embeddings.
\newblock In \emph{EMNLP}, 2021.

\bibitem[Gene Ontology(2021)]{gene2021gene}
Gene Ontology Consortium.
\newblock The Gene Ontology resource: enriching a GOld mine.
\newblock \emph{Nucleic Acids Research}, 49(D1):D325--D334, 2021.

\bibitem[Gewaltig and Diesmann(2007)]{gewaltig2007nest}
Gewaltig, M.-O. and Diesmann, M.
\newblock NEST (NEural Simulation Tool).
\newblock \emph{Scholarpedia}, 2(4):1430, 2007.

\bibitem[Glaser et~al.(2020)]{glaser2020machine}
Glaser, J.~I., et~al.
\newblock Machine learning for neural decoding.
\newblock \emph{eNeuro}, 7(4), 2020.

\bibitem[Gleeson et~al.(2019)]{gleeson2019systematic}
Gleeson, P., et~al.
\newblock Open source brain: a collaborative resource for visualizing, analyzing, simulating, and developing standardized models of neurons and circuits.
\newblock \emph{Neuron}, 103(3):395--411, 2019.

\bibitem[Gorgolewski et~al.(2015)]{gorgolewski2015neurovault}
Gorgolewski, K.~J., et~al.
\newblock NeuroVault.org: a web-based repository for collecting and sharing unthresholded statistical maps of the human brain.
\newblock \emph{Frontiers in Neuroinformatics}, 9:8, 2015.

\bibitem[Gorgolewski et~al.(2016)]{gorgolewski2016brain}
Gorgolewski, K.~J., et~al.
\newblock The brain imaging data structure, a format for organizing and describing outputs of neuroimaging experiments.
\newblock \emph{Scientific Data}, 3:160044, 2016.

\bibitem[Gray et~al.(2017)]{gray2017bioschemas}
Gray, A.~J., et~al.
\newblock Bioschemas: From potato salad to protein annotation.
\newblock In \emph{ISWC}, 2017.

\bibitem[Guo et~al.(2016)]{guo2016deep}
Guo, J., et~al.
\newblock A deep relevance matching model for ad-hoc retrieval.
\newblock In \emph{CIKM}, 2016.

\bibitem[Guo et~al.(2017)]{guo2017calibration}
Guo, C., Pleiss, G., Sun, Y., and Weinberger, K.~Q.
\newblock On calibration of modern neural networks.
\newblock In \emph{ICML}, 2017.

\bibitem[Hassabis et~al.(2017)]{hassabis2017neuroscience}
Hassabis, D., Kumaran, D., Summerfield, C., and Botvinick, M.
\newblock Neuroscience-inspired artificial intelligence.
\newblock \emph{Neuron}, 95(2):245--258, 2017.

\bibitem[Hearst(2006)]{hearst2006design}
Hearst, M.~A.
\newblock Design recommendations for hierarchical faceted search interfaces.
\newblock In \emph{SIGIR Workshop on Faceted Search}, 2006.

\bibitem[Hines et~al.(2004)]{hines2004modeldb}
Hines, M.~L., Morse, T., Migliore, M., Carnevale, N.~T., and Shepherd, G.~M.
\newblock ModelDB: A database to support computational neuroscience.
\newblock \emph{Journal of Computational Neuroscience}, 17(1):7--11, 2004.

\bibitem[Hines and Carnevale(1997)]{hines1997neuron}
Hines, M.~L. and Carnevale, N.~T.
\newblock The NEURON simulation environment.
\newblock \emph{Neural Computation}, 9(6):1179--1209, 1997.

\bibitem[Hodgkin and Huxley(1952)]{hodgkin1952quantitative}
Hodgkin, A.~L. and Huxley, A.~F.
\newblock A quantitative description of membrane current and its application to conduction and excitation in nerve.
\newblock \emph{Journal of Physiology}, 117(4):500--544, 1952.

\bibitem[Hofst{\"a}tter et~al.(2020)]{hofstatter2020improving}
Hofst{\"a}tter, S., et~al.
\newblock Improving efficient neural ranking models with cross-architecture knowledge distillation.
\newblock \emph{arXiv:2010.02666}, 2020.

\bibitem[Hu et~al.(2020)]{hu2020heterogeneous}
Hu, Z., et~al.
\newblock Heterogeneous graph transformer.
\newblock In \emph{WWW}, 2020.

\bibitem[IBL(2021)]{ibl2021standardized}
International Brain Laboratory.
\newblock Standardized and reproducible measurement of decision-making in mice.
\newblock \emph{eLife}, 10:e63711, 2021.

\bibitem[J{\"a}rvelin and Kek{\"a}l{\"a}inen(2002)]{jarvelin2002cumulated}
J{\"a}rvelin, K. and Kek{\"a}l{\"a}inen, J.
\newblock Cumulated gain-based evaluation of IR techniques.
\newblock \emph{ACM TOIS}, 20(4):422--446, 2002.

\bibitem[Jorgenson et~al.(2015)]{jorgenson2015brain}
Jorgenson, L.~A., et~al.
\newblock The BRAIN Initiative: developing technology to catalyse neuroscience discovery.
\newblock \emph{Philosophical Transactions B}, 370(1668), 2015.

\bibitem[Karpukhin et~al.(2020)]{karpukhin2020dense}
Karpukhin, V., et~al.
\newblock Dense passage retrieval for open-domain question answering.
\newblock In \emph{EMNLP}, 2020.

\bibitem[Khattab and Zaharia(2020)]{khattab2020colbert}
Khattab, O. and Zaharia, M.
\newblock ColBERT: Efficient and effective passage search via contextualized late interaction over BERT.
\newblock In \emph{SIGIR}, 2020.

\bibitem[Kilicoglu et~al.(2012)]{kilicoglu2012semmeddb}
Kilicoglu, H., et~al.
\newblock SemMedDB: a PubMed-scale repository of biomedical semantic predications.
\newblock \emph{Bioinformatics}, 28(23):3158--3160, 2012.

\bibitem[Laird et~al.(2005)]{laird2005brainmap}
Laird, A.~R., et~al.
\newblock BrainMap: the social evolution of a human brain mapping database.
\newblock \emph{Neuroinformatics}, 3(1):65--78, 2005.

\bibitem[Lebo et~al.(2013)]{lebo2013prov}
Lebo, T., et~al.
\newblock PROV-O: The PROV Ontology.
\newblock W3C Recommendation, 2013.

\bibitem[Lewis et~al.(2020)]{lewis2020retrieval}
Lewis, P., et~al.
\newblock Retrieval-augmented generation for knowledge-intensive NLP tasks.
\newblock In \emph{NeurIPS}, 2020.

\bibitem[Lin et~al.(2021)]{lin2021pretrained}
Lin, J., et~al.
\newblock Pretrained transformers for text ranking: BERT and beyond.
\newblock \emph{Synthesis Lectures on Human Language Technologies}, 14(4), 2021.

\bibitem[Lin et~al.(2021)]{lin2021few}
Lin, S.-C., et~al.
\newblock In-batch negatives for knowledge distillation with tightly-coupled teachers for dense retrieval.
\newblock In \emph{RepL4NLP}, 2021.

\bibitem[Liu(2009)]{liu2009learning}
Liu, T.-Y.
\newblock Learning to rank for information retrieval.
\newblock \emph{Foundations and Trends in IR}, 3(3):225--331, 2009.

\bibitem[Mallia et~al.(2021)]{mallia2021learning}
Mallia, A., et~al.
\newblock Learning passage impacts for inverted indexes.
\newblock In \emph{SIGIR}, 2021.

\bibitem[Markiewicz et~al.(2021)]{markiewicz2021openneuro}
Markiewicz, C.~J., et~al.
\newblock The OpenNeuro resource for sharing of neuroscience data.
\newblock \emph{eLife}, 10:e71774, 2021.

\bibitem[Moxon et~al.(2021)]{moxon2021linkml}
Moxon, S., et~al.
\newblock LinkML: A general-purpose data modeling language.
\newblock \emph{Bioinformatics}, 2021.

\bibitem[NeMO(2023)]{nemo2023archive}
NeMO Archive.
\newblock Neuroscience Multi-Omic Archive.
\newblock \url{https://nemoarchive.org/}, 2023.

\bibitem[Nogueira and Cho(2019)]{nogueira2019passage}
Nogueira, R. and Cho, K.
\newblock Passage re-ranking with BERT.
\newblock \emph{arXiv:1901.04085}, 2019.

\bibitem[Open Science Collaboration(2015)]{open2015estimating}
Open Science Collaboration.
\newblock Estimating the reproducibility of psychological science.
\newblock \emph{Science}, 349(6251):aac4716, 2015.

\bibitem[Pandarinath et~al.(2018)]{pandarinath2018inferring}
Pandarinath, C., et~al.
\newblock Inferring single-trial neural population dynamics using sequential auto-encoders.
\newblock \emph{Nature Methods}, 15(10):805--815, 2018.

\bibitem[Poldrack and Gorgolewski(2014)]{poldrack2014making}
Poldrack, R.~A. and Gorgolewski, K.~J.
\newblock Making big data open: data sharing in neuroimaging.
\newblock \emph{Nature Neuroscience}, 17(11):1510--1517, 2014.

\bibitem[Priem et~al.(2022)]{priem2022openalex}
Priem, J., Piwowar, H., and Orr, R.
\newblock OpenAlex: A fully-open index of scholarly works, authors, venues, institutions, and concepts.
\newblock \emph{arXiv:2205.01833}, 2022.

\bibitem[Reimers and Gurevych(2019)]{reimers2019sentence}
Reimers, N. and Gurevych, I.
\newblock Sentence-BERT: Sentence embeddings using Siamese BERT-networks.
\newblock In \emph{EMNLP-IJCNLP}, 2019.

\bibitem[Richards et~al.(2019)]{richards2019deep}
Richards, B.~A., et~al.
\newblock A deep learning framework for neuroscience.
\newblock \emph{Nature Neuroscience}, 22(11):1761--1770, 2019.

\bibitem[Robertson and Walker(2004)]{robertson2004simple}
Robertson, S. and Walker, S.
\newblock Some simple effective approximations to the 2-Poisson model for probabilistic weighted retrieval.
\newblock In \emph{SIGIR}, 2004.

\bibitem[Robertson and Zaragoza(2009)]{robertson2009probabilistic}
Robertson, S. and Zaragoza, H.
\newblock The probabilistic relevance framework: BM25 and beyond.
\newblock \emph{Foundations and Trends in IR}, 3(4):333--389, 2009.

\bibitem[Rubel et~al.(2022)]{rubel2022nwb}
Rubel, O., et~al.
\newblock The Neurodata Without Borders ecosystem for neurophysiological data science.
\newblock \emph{eLife}, 11:e78362, 2022.

\bibitem[Rubi{\'e}n et~al.(2024)]{rubinen2024dandi}
Rubi{\'e}n, D.~A., et~al.
\newblock The DANDI Archive: A platform for sharing standardized neurophysiology datasets.
\newblock \emph{Scientific Data}, 2024.

\bibitem[Salton et~al.(1975)]{salton1975vector}
Salton, G., Wong, A., and Yang, C.~S.
\newblock A vector space model for automatic indexing.
\newblock \emph{Communications of the ACM}, 18(11):613--620, 1975.

\bibitem[Sayers et~al.(2019)]{sayers2019pubmed}
Sayers, E.~W., et~al.
\newblock Database resources of the National Center for Biotechnology Information.
\newblock \emph{Nucleic Acids Research}, 47(D1):D23--D28, 2019.

\bibitem[Saxe et~al.(2021)]{saxe2021if}
Saxe, A., Nelli, S., and Summerfield, C.
\newblock If deep learning is the answer, what is the question?
\newblock \emph{Nature Reviews Neuroscience}, 22(1):55--67, 2021.

\bibitem[Schema.org(2022)]{schemaorg2022dataset}
Schema.org.
\newblock Dataset type definition.
\newblock \url{https://schema.org/Dataset}, 2022.

\bibitem[Schrimpf et~al.(2018)]{schrimpf2018brain}
Schrimpf, M., et~al.
\newblock Brain-Score: Which artificial neural network for object recognition is most brain-like?
\newblock \emph{bioRxiv}, 2018.

\bibitem[Sinha et~al.(2015)]{sinha2015overview}
Sinha, A., et~al.
\newblock An overview of Microsoft Academic Service (MAS) and applications.
\newblock In \emph{WWW Companion}, 2015.

\bibitem[Soiland-Reyes et~al.(2022)]{soiland2022packaging}
Soiland-Reyes, S., et~al.
\newblock Packaging research artefacts with RO-Crate.
\newblock \emph{Data Science}, 5(2):97--138, 2022.

\bibitem[Soska et~al.(2021)]{soska2021data}
Soska, K.~C., et~al.
\newblock Databrary: Enabling sharing and reuse of research video.
\newblock In \emph{CogSci}, 2021.

\bibitem[Stimberg et~al.(2019)]{stimberg2019brian}
Stimberg, M., Brette, R., and Goodman, D.~F.
\newblock Brian 2, an intuitive and efficient neural simulator.
\newblock \emph{eLife}, 8:e47314, 2019.

\bibitem[Sun et~al.(2011)]{sun2011pathsim}
Sun, Y., Han, J., Yan, X., Yu, P.~S., and Wu, T.
\newblock PathSim: Meta path-based top-k similarity search in heterogeneous information networks.
\newblock \emph{PVLDB}, 4(11):992--1003, 2011.

\bibitem[Sussillo and Abbott(2009)]{sussillo2013opening}
Sussillo, D. and Abbott, L.~F.
\newblock Generating coherent patterns of activity from chaotic neural networks.
\newblock \emph{Neuron}, 63(4):544--557, 2009.

\bibitem[Tedersoo et~al.(2021)]{tedersoo2021data}
Tedersoo, L., et~al.
\newblock Data sharing practices and data availability upon request differ across scientific disciplines.
\newblock \emph{Scientific Data}, 8:192, 2021.

\bibitem[Teeters et~al.(2008)]{teeters2008crcns}
Teeters, J.~L., et~al.
\newblock Data sharing for computational neuroscience.
\newblock \emph{Neuroinformatics}, 6(1):47--55, 2008.

\bibitem[Teeters et~al.(2015)]{teeters2015neurodata}
Teeters, J.~L., et~al.
\newblock Neurodata Without Borders: Creating a common data format for neurophysiology.
\newblock \emph{Neuron}, 88(4):629--634, 2015.

\bibitem[Tenopir et~al.(2020)]{tenopir2020data}
Tenopir, C., et~al.
\newblock Data sharing, management, use, and reuse: Practices and perceptions of scientists worldwide.
\newblock \emph{PLOS ONE}, 15(3):e0229003, 2020.

\bibitem[UniProt(2021)]{uniprot2021uniprot}
UniProt Consortium.
\newblock UniProt: the universal protein knowledgebase in 2021.
\newblock \emph{Nucleic Acids Research}, 49(D1):D480--D489, 2021.

\bibitem[Van Essen et~al.(2013)]{van2013wu}
Van Essen, D.~C., et~al.
\newblock The WU-Minn Human Connectome Project: an overview.
\newblock \emph{NeuroImage}, 80:62--79, 2013.

\bibitem[Wallis et~al.(2013)]{wallis2013if}
Wallis, J.~C., Rolando, E., and Borgman, C.~L.
\newblock If we share data, will anyone use them? Data sharing and reuse in the long tail of science and technology.
\newblock \emph{PLOS ONE}, 8(7):e67332, 2013.

\bibitem[Wang et~al.(2017)]{wang2017knowledge}
Wang, Q., et~al.
\newblock Knowledge graph embedding: A survey of approaches and applications.
\newblock \emph{IEEE TKDE}, 29(12):2724--2743, 2017.

\bibitem[Wang et~al.(2019)]{wang2019heterogeneous}
Wang, X., et~al.
\newblock Heterogeneous graph attention network.
\newblock In \emph{WWW}, 2019.

\bibitem[Wilkinson et~al.(2016)]{wilkinson2016fair}
Wilkinson, M.~D., et~al.
\newblock The FAIR Guiding Principles for scientific data management and stewardship.
\newblock \emph{Scientific Data}, 3:160018, 2016.

\bibitem[Xiong et~al.(2020)]{xiong2020approximate}
Xiong, L., et~al.
\newblock Approximate nearest neighbor negative contrastive learning for dense text retrieval.
\newblock In \emph{ICLR}, 2020.

\bibitem[Yarkoni et~al.(2011)]{yarkoni2019large}
Yarkoni, T., et~al.
\newblock Large-scale automated synthesis of human functional neuroimaging data.
\newblock \emph{Nature Methods}, 8(8):665--670, 2011.

\bibitem[Ye et~al.(2023)]{ye2023neural}
Ye, J., et~al.
\newblock Neural data transformer: Large-scale pre-training of neural recordings.
\newblock \emph{arXiv:2301.04517}, 2023.

\bibitem[Zador et~al.(2023)]{zador2023catalyzing}
Zador, A., et~al.
\newblock Catalyzing next-generation Artificial Intelligence through NeuroAI.
\newblock \emph{Nature Communications}, 14:1597, 2023.

\end{thebibliography}

% ============================================================================
% APPENDIX
% ============================================================================
\appendix

\section{Algorithm Pseudocode}
\label{app:algorithms}

\begin{algorithm}[H]
\caption{Neural Search Hybrid Retrieval}
\begin{algorithmic}[1]
\Procedure{Search}{$\query, \text{corpus}, \text{config}$}
    \State $\text{intent} \gets \Call{ClassifyIntent}{\query}$
    \State $\mathbf{w} \gets \Call{GetWeights}{\text{intent}}$
    \State $(\mathcal{C}_{\text{hard}}, \mathcal{C}_{\text{soft}}, \mathcal{A}_{\text{req}}) \gets \Call{ParseConstraints}{\query}$
    \State $\text{candidates} \gets \Call{GenerateCandidates}{\query, \text{corpus}}$
    \State $\text{candidates} \gets \{d \in \text{candidates} : \text{satisfies}(d, \mathcal{C}_{\text{hard}})\}$
    \For{$d \in \text{candidates}$}
        \State $S_{\text{BM25}} \gets \Call{BM25Score}{d, \query}$
        \State $S_{\text{semantic}} \gets \Call{SemanticScore}{d, \query}$
        \State $S_{\text{onto}} \gets \Call{OntologyScore}{d, \query}$
        \State $S_{\text{graph}} \gets \Call{GraphScore}{d, \query}$
        \State $S_{\text{afford}} \gets \Call{AffordanceScore}{d, \mathcal{A}_{\text{req}}}$
        \State $\text{scores}[d] \gets \Call{Fuse}{\mathbf{w}, [S_{\text{BM25}}, S_{\text{semantic}}, S_{\text{onto}}, S_{\text{graph}}, S_{\text{afford}}]}$
    \EndFor
    \State \Return $\Call{Sort}{\text{candidates}, \text{scores}}$
\EndProcedure
\end{algorithmic}
\end{algorithm}

\section{Ontology Examples}
\label{app:ontology}

\begin{table}[H]
\centering
\small
\caption{Task ontology hierarchy (partial).}
\begin{tabular}{ll}
\toprule
Level 1 & Level 2 \\
\midrule
Decision-making & Two-alternative forced choice (2AFC) \\
& Go/No-Go \\
& Multi-armed bandit \\
& Reversal learning \\
\midrule
Working memory & Delayed match-to-sample \\
& N-back \\
& Spatial working memory \\
\midrule
Sensory processing & Visual discrimination \\
& Auditory discrimination \\
& Passive viewing \\
\midrule
Motor control & Reaching \\
& Locomotion \\
& Eye movements \\
\bottomrule
\end{tabular}
\end{table}

\section{Graph Schema Details}
\label{app:graph-schema}

\begin{table}[H]
\centering
\small
\caption{Node types and their properties.}
\begin{tabular}{lp{6cm}}
\toprule
Node Type & Key Properties \\
\midrule
\texttt{dataset} & source, source\_id, title, description, species, modalities, tasks, regions, standards \\
\texttt{paper} & doi, title, authors, journal, year, abstract \\
\texttt{task} & label, ontology\_id, family, description \\
\texttt{modality} & label, family, signal\_type, device \\
\texttt{species} & taxon\_id, common\_name, scientific\_name \\
\texttt{region} & label, hierarchy\_path, allen\_id \\
\texttt{analysis} & label, requirements, description \\
\bottomrule
\end{tabular}
\end{table}

\end{document}
